---
title: "Fig1E_F"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## Fig 1E

Boxplot showing expression level of HIST1H1B, a nonpolyadenylated histone subunit gene.

Boxplot showing expression of average median ratio gene, calculated as the median expression level in PolyA samples divided by the median expression level in RiboD samples

```{r load libraries}
library(tidyverse)
library(patchwork)
library(cowplot)
```

```{r load in files}
# expression files
SS_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/SS_polyA_log2tpm1.tsv")
SS_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/SS_riboD_log2tpm1.tsv")

aRMS_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/aRMS_polyA_log2tpm1.tsv")
aRMS_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/aRMS_riboD_log2tpm1.tsv")

WT_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/WT_polyA_log2tpm1.tsv")
WT_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/WT_riboD_log2tpm1.tsv")

NB_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/NB_polyA_log2tpm1.tsv")
NB_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/NB_riboD_log2tpm1.tsv")

ALL_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/ALL_polyA_log2tpm1.tsv")
ALL_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/ALL_riboD_log2tpm1.tsv")

AML_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/AML_polyA_log2tpm1.tsv")
AML_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/AML_riboD_log2tpm1.tsv")

# sample ID files, from correlation_analysis
SS_polyA_list <- read_tsv("../../input_data/sample_selection/SS_polyA_list.tsv")
SS_riboD_list <- read_tsv("../../input_data/sample_selection/SS_riboD_list.tsv")

aRMS_polyA_list <- read_tsv("../../input_data/sample_selection/aRMS_polyA_list.tsv")
aRMS_riboD_list <- read_tsv("../../input_data/sample_selection/aRMS_riboD_list.tsv")

WT_polyA_list <- read_tsv("../../input_data/sample_selection/WT_polyA_list.tsv")
WT_riboD_list <- read_tsv("../../input_data/sample_selection/WT_riboD_list.tsv")

NB_polyA_list <- read_tsv("../../input_data/sample_selection/NB_polyA_list.tsv")
NB_riboD_list <- read_tsv("../../input_data/sample_selection/NB_riboD_list.tsv")

ALL_polyA_list <- read_tsv("../../input_data/sample_selection/ALL_polyA_list.tsv")
ALL_riboD_list <- read_tsv("../../input_data/sample_selection/ALL_riboD_list.tsv")

AML_polyA_list <- read_tsv("../../input_data/sample_selection/AML_polyA_list.tsv")
AML_riboD_list <- read_tsv("../../input_data/sample_selection/AML_riboD_list.tsv")

# to convert EnsemblIDs to HugoIDs
gene_names <- read.table("../../input_data/EnsGeneID_Hugo_Observed_Conversions.txt",
header = TRUE, sep = "\t", stringsAsFactors = FALSE )
```

Convert EnsemblIDs to HugoIDs

```{r convert to Hugo function}
convert_to_hugo <- function(df) {
  df %>%
    left_join(gene_names, by = c("Gene" = "EnsGeneID")) %>%
    mutate(Gene = ifelse(!is.na(HugoID), HugoID, Gene)) %>%
    select(-HugoID)
}
```

Finding median expression genes

```{r median expression gene - SS}
# Find gene medians in polyA SS
SS_polyA_medians <- SS_polyA_log2tpm1 %>%
  rowwise() %>%
  mutate(SS_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, SS_polyA_median)

# Find gene medians in riboD SS
SS_riboD_medians <- SS_riboD_log2tpm1 %>%
  rowwise() %>%
  mutate(SS_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, SS_riboD_median)

# Join and compute ratio (polyA median/riboD median)
SS_gene_medians <- SS_polyA_medians %>%
  inner_join(SS_riboD_medians, by = "Gene") %>%
  mutate(SS_median_ratio = SS_polyA_median / SS_riboD_median)

# Filter out genes with no expression
SS_gene_medians_filtered <- SS_gene_medians %>%
  filter(SS_polyA_median > 0 & SS_riboD_median > 0)

# Recompute median ratio value
SS_median_ratio_val <- median(SS_gene_medians_filtered$SS_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
SS_most_average_gene <- SS_gene_medians_filtered %>%
  mutate(dist = abs(SS_median_ratio - SS_median_ratio_val)) %>%
  arrange(dist) %>%
  slice(1)  

SS_most_average_gene_hugo <- SS_most_average_gene %>%
  convert_to_hugo()
SS_most_average_gene_hugo
write_tsv(SS_most_average_gene_hugo, "../../output_data/Fig1E_F/SS_most_average_gene_hugo.tsv.gz")
```

```{r median expression gene - aRMS}
# Find gene medians in polyA aRMS
aRMS_polyA_medians <- aRMS_polyA_log2tpm1 %>%
  rowwise() %>%
  mutate(aRMS_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, aRMS_polyA_median)

# Find gene medians in riboD aRMS
aRMS_riboD_medians <- aRMS_riboD_log2tpm1 %>%
  rowwise() %>%
  mutate(aRMS_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, aRMS_riboD_median)

# Join and compute ratio (polyA median/riboD median)
aRMS_gene_medians <- aRMS_polyA_medians %>%
  inner_join(aRMS_riboD_medians, by = "Gene") %>%
  mutate(aRMS_median_ratio = aRMS_polyA_median / aRMS_riboD_median)

# Filter out genes with no or near-zero expression
aRMS_gene_medians_filtered <- aRMS_gene_medians %>%
  filter(aRMS_polyA_median > 0 & aRMS_riboD_median > 0)

# Recompute median ratio value
aRMS_median_ratio_val <- median(aRMS_gene_medians_filtered$aRMS_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
aRMS_most_average_gene <- aRMS_gene_medians_filtered %>%
  mutate(dist = abs(aRMS_median_ratio - aRMS_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)   # Isolates gene closest to ave median ration

aRMS_most_average_gene_hugo <- aRMS_most_average_gene %>%
  convert_to_hugo()
aRMS_most_average_gene_hugo
```

```{r median expression gene - WT}
# Find gene medians in polyA WT
WT_polyA_medians <- WT_polyA_log2tpm1 %>%
  rowwise() %>%
  mutate(WT_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, WT_polyA_median)

# Find gene medians in riboD WT
WT_riboD_medians <- WT_riboD_log2tpm1 %>%
  rowwise() %>%
  mutate(WT_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, WT_riboD_median)

# Join and compute ratio (polyA median/riboD median)
WT_gene_medians <- WT_polyA_medians %>%
  inner_join(WT_riboD_medians, by = "Gene") %>%
  mutate(WT_median_ratio = WT_polyA_median / WT_riboD_median)

# Filter out genes with no or near-zero expression
WT_gene_medians_filtered <- WT_gene_medians %>%
  filter(WT_polyA_median > 0 & WT_riboD_median > 0)

# Recompute median ratio value
WT_median_ratio_val <- median(WT_gene_medians_filtered$WT_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
WT_most_average_gene <- WT_gene_medians_filtered %>%
  mutate(dist = abs(WT_median_ratio - WT_median_ratio_val)) %>%
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

WT_most_average_gene_hugo <- WT_most_average_gene %>%
  convert_to_hugo()
WT_most_average_gene_hugo
```

```{r median expression gene - NB}
# Find gene medians in polyA NB
NB_polyA_medians <- NB_polyA_log2tpm1 %>%
  rowwise() %>%
  mutate(NB_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, NB_polyA_median)

# Find gene medians in riboD NB
NB_riboD_medians <- NB_riboD_log2tpm1 %>%
  rowwise() %>%
  mutate(NB_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, NB_riboD_median)

# Join and compute ratio (polyA median/riboD median)
NB_gene_medians <- NB_polyA_medians %>%
  inner_join(NB_riboD_medians, by = "Gene") %>%
  mutate(NB_median_ratio = NB_polyA_median / NB_riboD_median)

# Filter out genes with no or near-zero expression
NB_gene_medians_filtered <- NB_gene_medians %>%
  filter(NB_polyA_median > 0 & NB_riboD_median > 0)

# Recompute median ratio value
NB_median_ratio_val <- median(NB_gene_medians_filtered$NB_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
NB_most_average_gene <- NB_gene_medians_filtered %>%
  mutate(dist = abs(NB_median_ratio - NB_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

NB_most_average_gene_hugo <- NB_most_average_gene %>%
  convert_to_hugo()
NB_most_average_gene_hugo
```

```{r median expression gene - ALL}
# Find gene medians in polyA ALL
ALL_polyA_medians <- ALL_polyA_log2tpm1 %>%
  rowwise() %>%
  mutate(ALL_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, ALL_polyA_median)

# Find gene medians in riboD ALL
ALL_riboD_medians <- ALL_riboD_log2tpm1 %>%
  rowwise() %>%
  mutate(ALL_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, ALL_riboD_median)

# Join and compute ratio (polyA median/riboD median)
ALL_gene_medians <- ALL_polyA_medians %>%
  inner_join(ALL_riboD_medians, by = "Gene") %>%
  mutate(ALL_median_ratio = ALL_polyA_median / ALL_riboD_median)

# Filter out genes with no or near-zero expression
ALL_gene_medians_filtered <- ALL_gene_medians %>%
  filter(ALL_polyA_median > 0 & ALL_riboD_median > 0)

# Recompute median ratio value
ALL_median_ratio_val <- median(ALL_gene_medians_filtered$ALL_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
ALL_most_average_gene <- ALL_gene_medians_filtered %>%
  mutate(dist = abs(ALL_median_ratio - ALL_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

ALL_most_average_gene_hugo <- ALL_most_average_gene %>%
  convert_to_hugo()
ALL_most_average_gene_hugo
```

```{r median expression gene - AML}
# Find gene medians in polyA AML
AML_polyA_medians <- AML_polyA_log2tpm1 %>%
  rowwise() %>%
  mutate(AML_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, AML_polyA_median)

# Find gene medians in riboD AML
AML_riboD_medians <- AML_riboD_log2tpm1 %>%
  rowwise() %>%
  mutate(AML_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, AML_riboD_median)

# Join and compute ratio (polyA median/riboD median)
AML_gene_medians <- AML_polyA_medians %>%
  inner_join(AML_riboD_medians, by = "Gene") %>%
  mutate(AML_median_ratio = AML_polyA_median / AML_riboD_median)

# Filter out genes with no or near-zero expression
AML_gene_medians_filtered <- AML_gene_medians %>%
  filter(AML_polyA_median > 1 & AML_riboD_median > 0.9)

# Recompute median ratio value
AML_median_ratio_val <- median(AML_gene_medians_filtered$AML_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
AML_most_average_gene <- AML_gene_medians_filtered %>%
  mutate(dist = abs(AML_median_ratio - AML_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

AML_most_average_gene_hugo <- AML_most_average_gene %>%
  convert_to_hugo()
AML_most_average_gene_hugo
```

Converting expression to hugoIDs

```{r convert to HugoIDs}
# SS
SS_polyA_expression_hugo <- SS_polyA_log2tpm1 %>%
  convert_to_hugo()

SS_riboD_expression_hugo <- SS_riboD_log2tpm1 %>%
  convert_to_hugo()

SS_gene_medians_hugo <- SS_gene_medians %>%
  convert_to_hugo()

# aRMS
aRMS_polyA_expression_hugo <- aRMS_polyA_log2tpm1 %>%
  convert_to_hugo() 

aRMS_riboD_expression_hugo <- aRMS_riboD_log2tpm1 %>%
  convert_to_hugo()

aRMS_gene_median_hugo <- aRMS_gene_medians %>%
  convert_to_hugo()

# WT
WT_polyA_expression_hugo <- WT_polyA_log2tpm1 %>%
  convert_to_hugo()

WT_riboD_expression_hugo <- WT_riboD_log2tpm1 %>%
  convert_to_hugo()

WT_gene_medians_hugo <- WT_gene_medians %>%
  convert_to_hugo()

# NB
NB_polyA_expression_hugo <- NB_polyA_log2tpm1 %>%
  convert_to_hugo()

NB_riboD_expression_hugo <- NB_riboD_log2tpm1 %>%
  convert_to_hugo()

NB_gene_medians_hugo <- NB_gene_medians %>%
  convert_to_hugo()

# ALL
ALL_polyA_expression_hugo <- ALL_polyA_log2tpm1 %>%
  convert_to_hugo()

ALL_riboD_expression_hugo <- ALL_riboD_log2tpm1 %>%
  convert_to_hugo()

ALL_gene_medians_hugo <- ALL_gene_medians %>%
  convert_to_hugo()

# AML
AML_polyA_expression_hugo <- AML_polyA_log2tpm1 %>%
  convert_to_hugo()

AML_riboD_expression_hugo <- AML_riboD_log2tpm1 %>%
  convert_to_hugo()

AML_gene_medians_hugo <- AML_gene_medians %>%
  convert_to_hugo()
```

Filtering for HIST1H1B, a non-polyadenylated gene

```{r HIST1H1B filter}
H1B_SS_polyA_expression <- SS_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_SS_riboD_expression <- SS_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_aRMS_polyA_expression <- aRMS_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_aRMS_riboD_expression <- aRMS_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_WT_polyA_expression <- WT_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_WT_riboD_expression <- WT_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_NB_polyA_expression <- NB_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_NB_riboD_expression <- NB_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_ALL_polyA_expression <- ALL_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_ALL_riboD_expression <- ALL_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_AML_polyA_expression <- AML_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_AML_riboD_expression <- AML_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 
```

Filtering for most averagely expressed gene

```{r avg expression gene filter}
#SS Most Average Expressed gene
avg_SS_polyA_expression <- SS_polyA_expression_hugo %>%
   filter(Gene == SS_most_average_gene_hugo$Gene) 

avg_SS_riboD_expression <- SS_riboD_expression_hugo %>%
   filter(Gene == SS_most_average_gene_hugo$Gene)

#aRMS Most Average Expressed gene
avg_aRMS_riboD_expression <- aRMS_riboD_expression_hugo %>%
   filter(Gene == aRMS_most_average_gene_hugo$Gene)

avg_aRMS_polyA_expression <- aRMS_polyA_expression_hugo %>%
   filter(Gene == aRMS_most_average_gene_hugo$Gene)

#WT Most Average Expressed gene 
avg_WT_riboD_expression <- WT_riboD_expression_hugo %>%
   filter(Gene == WT_most_average_gene_hugo$Gene)

avg_WT_polyA_expression <- WT_polyA_expression_hugo %>%
   filter(Gene == WT_most_average_gene_hugo$Gene)

#NB Most Average Expressed gene
avg_NB_riboD_expression <- NB_riboD_expression_hugo %>%
   filter(Gene == NB_most_average_gene_hugo$Gene)

avg_NB_polyA_expression <- NB_polyA_expression_hugo %>%
   filter(Gene == NB_most_average_gene_hugo$Gene)

#ALL Most Average Expressed gene
avg_ALL_riboD_expression <- ALL_riboD_expression_hugo %>%
   filter(Gene == ALL_most_average_gene_hugo$Gene)

avg_ALL_polyA_expression <- ALL_polyA_expression_hugo %>%
   filter(Gene == ALL_most_average_gene_hugo$Gene)

#AML Most Average Expressed gene
avg_AML_riboD_expression <- AML_riboD_expression_hugo %>%
   filter(Gene == AML_most_average_gene_hugo$Gene)

avg_AML_polyA_expression <- AML_polyA_expression_hugo %>%
   filter(Gene == AML_most_average_gene_hugo$Gene)
```

Reformatting HIST1H1B data

```{r reformat HIST1H1B}
H1B_SS_polyA_long <- H1B_SS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "polyA", Gene = "HIST1H1B")

H1B_SS_riboD_long <- H1B_SS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "riboD", Gene = "HIST1H1B")

H1B_aRMS_polyA_long <- H1B_aRMS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "polyA", Gene = "HIST1H1B")

H1B_aRMS_riboD_long <- H1B_aRMS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "riboD", Gene = "HIST1H1B")

H1B_WT_polyA_long <- H1B_WT_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "polyA", Gene = "HIST1H1B")

H1B_WT_riboD_long <- H1B_WT_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "riboD", Gene = "HIST1H1B")

H1B_NB_polyA_long <- H1B_NB_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "polyA", Gene = "HIST1H1B")

H1B_NB_riboD_long <- H1B_NB_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "riboD", Gene = "HIST1H1B")

H1B_ALL_polyA_long <- H1B_ALL_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "polyA", Gene = "HIST1H1B")

H1B_ALL_riboD_long <- H1B_ALL_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "riboD", Gene = "HIST1H1B")

H1B_AML_polyA_long <- H1B_AML_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "polyA", Gene = "HIST1H1B")

H1B_AML_riboD_long <- H1B_AML_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "riboD", Gene = "HIST1H1B")
```

Reformatting avg gene

```{r reformat avg gene}
# SS polyA 
avg_SS_polyA_long <- avg_SS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "polyA", Gene = Gene)

# SS riboD
avg_SS_riboD_long <- avg_SS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "riboD", Gene = Gene)

# aRMS polyA
avg_aRMS_polyA_long <- avg_aRMS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "polyA", Gene = Gene)

# aRMS riboD
avg_aRMS_riboD_long <- avg_aRMS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "riboD", Gene = Gene)

#NB polyA
avg_NB_polyA_long <- avg_NB_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "polyA", Gene = Gene)

#NB riboD
avg_NB_riboD_long <- avg_NB_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "riboD", Gene = Gene)

#WT polyA
avg_WT_polyA_long <- avg_WT_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "polyA", Gene = Gene)

#WT riboD
avg_WT_riboD_long <- avg_WT_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "riboD", Gene = Gene)

#AML polyA
avg_AML_polyA_long <- avg_AML_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "polyA", Gene = Gene)

#AML riboD
avg_AML_riboD_long <- avg_AML_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "riboD", Gene = Gene)

#ALL polyA
avg_ALL_polyA_long <- avg_ALL_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "polyA", Gene = Gene)

#ALL riboD
avg_ALL_riboD_long <- avg_ALL_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "riboD", Gene = Gene)
```

Combining dataframes

```{r combining}
combined_long <- bind_rows(

  # -------------------------
  # HIST1H1B — SS, aRMS, WT, NB, ALL, AML
  # -------------------------
  H1B_SS_polyA_long %>%
    mutate(Disease = "SS"),
  H1B_SS_riboD_long %>%
    mutate(Disease = "SS"),
  
  H1B_aRMS_polyA_long %>%
    mutate(Disease = "aRMS"),
  H1B_aRMS_riboD_long %>%
    mutate(Disease = "aRMS"),
  
  H1B_WT_polyA_long %>%
    mutate(Disease = "WT"),
  H1B_WT_riboD_long %>%
    mutate(Disease = "WT"),
  
  H1B_NB_polyA_long %>%
    mutate(Disease = "NB"),
  H1B_NB_riboD_long %>%
    mutate(Disease = "NB"),
  
  H1B_ALL_polyA_long %>%
    mutate(Disease = "ALL"),
  H1B_ALL_riboD_long %>%
    mutate(Disease = "ALL"),
  
  H1B_AML_polyA_long %>%
    mutate(Disease = "AML"),
  H1B_AML_riboD_long %>%
    mutate(Disease = "AML"),
   # -------------------------
  # Median-ratio genes (SS, aRMS, WT, NB, ALL, AML)
  # -------------------------
  avg_SS_polyA_long %>%
    mutate(Disease = "SS"),
  avg_SS_riboD_long %>%
    mutate(Disease = "SS"),

  avg_aRMS_polyA_long %>%
    mutate(Disease = "aRMS"),
  avg_aRMS_riboD_long %>%
    mutate(Disease = "aRMS"),

  avg_WT_polyA_long %>%
    mutate(Disease = "WT"),
  avg_WT_riboD_long %>%
    mutate(Disease = "WT"),

  avg_NB_polyA_long %>%
    mutate(Disease = "NB"),
  avg_NB_riboD_long %>%
    mutate(Disease = "NB"),
  
  avg_ALL_polyA_long %>%
    mutate(Disease = "ALL"),
  avg_ALL_riboD_long %>%
    mutate(Disease = "ALL"),
  
  avg_AML_polyA_long %>%
    mutate(Disease = "AML"),
  avg_AML_riboD_long %>%
    mutate(Disease = "AML"),
)
write_tsv(combined_long, "../../output_data/Fig1E_F/combined_long.tsv.gz")
```

Custom color theme

```{r color theme}
color_theme <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "top",
      legend.title = element_blank(),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.4),
      axis.ticks = element_line(color = "black", linewidth = 0.4),
      strip.text = element_text(face = "bold", size = base_size * 0.9),
      plot.title = element_text(face = "bold", size = base_size * 1.1, hjust = 0.5),
      axis.title.y = element_text(angle = 90, hjust = 0.5, size = 16), 
      axis.title.x = element_text(angle = 0, hjust = 0.5, size = 17), 
      axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
      plot.margin = margin(10, 10, 10, 10)
    )
}

scale_fill_compendia <- function() {
  scale_fill_manual(values = c(
    "riboD" = "#E69F00",  # yellow
    "polyA" = "#0072B2"   # blue
  ))
}
scale_color_compendia <- function() {
  scale_color_manual(values = c( 
    "riboD" = "#E69F00", 
    "polyA" = "#0072B2" ))
  }
```

```{r}
# Map of median-ratio genes
median_gene_map <- list(
  SS      = SS_most_average_gene_hugo$Gene,
  aRMS    = aRMS_most_average_gene_hugo$Gene,
  WT      = WT_most_average_gene_hugo$Gene,
  NB      = NB_most_average_gene_hugo$Gene,
  ALL     = ALL_most_average_gene_hugo$Gene,
  AML     = AML_most_average_gene_hugo$Gene
)
median_gene_map_df <- as.data.frame(median_gene_map)

write_tsv(median_gene_map_df, "../../output_data/Fig1E_F/median_gene_map_df.tsv.gz")

diseases <- unique(combined_long$Disease)

```

Statistical test for significance

```{r selecting correct test}
choose_test <- function(poly, ribo) {
  
  # Shapiro tests
  p_norm_poly  <- shapiro.test(poly)$p.value
  p_norm_ribo  <- shapiro.test(ribo)$p.value
  
  poly_normal <- p_norm_poly  > 0.05
  ribo_normal <- p_norm_ribo  > 0.05
  
  # Both normal → variance test
  if (poly_normal & ribo_normal) {
    
    p_var <- var.test(poly, ribo)$p.value
    equal_var <- p_var > 0.05
    
    # Normal + equal variance → Student t-test
    if (equal_var) {
      return(list(
        p_value = t.test(poly, ribo, var.equal = TRUE)$p.value,
        test_used = "student_t"
      ))
    }
    
    # Normal + unequal variance → Welch t-test
    else {
      return(list(
        p_value = t.test(poly, ribo, var.equal = FALSE)$p.value,
        test_used = "welch_t"
      ))
    }
  }
  
  # Non-normal → Mann–Whitney U test
  return(list(
    p_value = wilcox.test(poly, ribo, exact = FALSE)$p.value,
    test_used = "mann_whitney"
  ))
}
```

Assigning P-value to significance star

```{r sig stars}
p_to_stars <- function(p) 
{ vapply(p, function(x) 
{ if (is.na(x)) return("NA") 
  if (x < 0.0001) return("****") 
  if (x < 0.001) return("***") 
  if (x < 0.01) return("**") 
  if (x < 0.05) return("*") 
  "ns" 
  }, FUN.VALUE = character(1)) }
```

Applying to data

```{r compute significance}
compute_significance_table <- function(df) {
  
  # 1. Clean and deduplicate input
  df_clean <- df %>%
    select(Disease, Gene, Compendia, Expression) %>%
    distinct()
  
  # 2. Collapse to one row per Disease × Gene × Compendia
  df_collapsed <- df_clean %>%
    group_by(Disease, Gene, Compendia) %>%
    summarise(Expression = list(Expression), .groups = "drop")
  
  # 3. Pivot to wide: polyA and riboD in separate columns
  df_wide <- df_collapsed %>%
    tidyr::pivot_wider(
      names_from = Compendia,
      values_from = Expression
    )
  
  # 4. Apply statistical test to each row
  df_results <- df_wide %>%
    rowwise() %>%
    mutate(
      # wrap the returned list so mutate() treats it as a single object
      test = list(choose_test(unlist(polyA), unlist(riboD))),
      p_value = test$p_value,
      test_used = test$test_used,
      stars = p_to_stars(p_value)
    ) %>%
    ungroup() %>%
    select(-test)
  
    # Benjamini–Hochberg FDR correction
  df_results2 <- df_results %>%
    group_by(Disease) %>%
    mutate(padj = p.adjust(p_value, method = "BH")) %>%
    ungroup() %>%
    mutate(stars_adj = p_to_stars(padj))
  
  df_results2
}


sig_results <- compute_significance_table(combined_long)
write_tsv(sig_results, "../../output_data/Fig1E_F/sig_results.tsv.gz")

sig_results
```

Apply stars to figures

```{r Fig1E_SS}
Fig1E_SS <- map(diseases, function(d) {

  median_gene <- median_gene_map[[d]]

  # HIST1H1B expression
  df_h1b <- combined_long %>%
    filter(Disease == d, Gene == "HIST1H1B") %>%
    mutate(PlotType = "HIST1H1B (PolyA-)")

  # Median-ratio gene expression
  df_median <- combined_long %>%
    filter(Disease == d, Gene == median_gene) %>%
    mutate(PlotType = "Median Ratio Gene (PolyA+)")

  # Combine
  df_combined <- bind_rows(df_h1b, df_median)

  # SIGNIFICANCE STARS
  star_df <- tibble(
    Disease = d,
    Gene = c("HIST1H1B", median_gene),
    PlotType = c("HIST1H1B (PolyA-)", "Median Ratio Gene (PolyA+)"),
    stars_adj = sig_results %>%
      filter(Disease == d, Gene %in% c("HIST1H1B", median_gene)) %>%
      arrange(match(Gene, c("HIST1H1B", median_gene))) %>%
      pull(stars_adj)
  )

  # y-position for stars (slightly above max)
  y_star <- max(df_combined$Expression, na.rm = TRUE) * 1.05

  ggplot(df_combined, aes(x = Gene, y = Expression, fill = Compendia)) +
    geom_boxplot(
      alpha = 0.7,
      position = position_dodge(width = 0.8),
      outlier.shape = NA,
      linewidth = 0.4
    ) +
    geom_jitter(
      aes(color = Compendia),
      position = position_jitterdodge(jitter.width = 0.2,
                                      dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6
    ) +
    # ADD STARS
    geom_text(
      data = star_df,
      aes(x = Gene, y = y_star, label = stars_adj),
      inherit.aes = FALSE,
      size = 12
    ) +
    scale_fill_compendia() +
    scale_color_compendia() +
    coord_cartesian(ylim = c(0, 10)) +
    facet_wrap(~ PlotType, ncol = 2, scales = "free_x") +
    labs(
      title = paste(d),
      x = NULL,
      y = "Expression log2(TPM+1)"
    ) +
    color_theme() +
    theme(
      axis.title.y = element_text(size = 20),
      axis.text.y = element_text(size = 14),
      axis.text.x = element_text(size = 20),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2),
      legend.position = "bottom",
      legend.text = element_text(size = 16),       # <-- label text size
      legend.title = element_text(size = 18),      # <-- title text size
      legend.key.size = unit(1, "cm"), # <-- box/key size
      legend.margin = margin(t = -2) 
    ) 
})
Fig1E_SS[[1]]
```

```{r save Fig1E}
ggsave("../../Figures/Fig1E.png", plot = Fig1E_SS[[1]])
ggsave("../../Figures/Fig1E.tif", plot = Fig1E_SS[[1]])
saveRDS(Fig1E_SS[[1]],"../../Figures/Fig1E.rds")
```

```{r Fig1E}
Fig1E <- map(diseases, function(d) {

  median_gene <- median_gene_map[[d]]

  # HIST1H1B expression
  df_h1b <- combined_long %>%
    filter(Disease == d, Gene == "HIST1H1B") %>%
    mutate(PlotType = "HIST1H1B (PolyA-)")

  # Median-ratio gene expression
  df_median <- combined_long %>%
    filter(Disease == d, Gene == median_gene) %>%
    mutate(PlotType = "Median Ratio Gene (PolyA+)")

  # Combine
  df_combined <- bind_rows(df_h1b, df_median)

  # SIGNIFICANCE STARS
  star_df <- tibble(
    Disease = d,
    Gene = c("HIST1H1B", median_gene),
    PlotType = c("HIST1H1B (PolyA-)", "Median Ratio Gene (PolyA+)"),
    stars_adj = sig_results %>%
      filter(Disease == d, Gene %in% c("HIST1H1B", median_gene)) %>%
      arrange(match(Gene, c("HIST1H1B", median_gene))) %>%
      pull(stars_adj)
  )

  # y-position for stars (slightly above max)
  y_star <- max(df_combined$Expression, na.rm = TRUE) * 1.05

  ggplot(df_combined, aes(x = Gene, y = Expression, fill = Compendia)) +
    geom_boxplot(
      alpha = 0.7,
      position = position_dodge(width = 0.8),
      outlier.shape = NA,
      linewidth = 0.4
    ) +
    geom_jitter(
      aes(color = Compendia),
      position = position_jitterdodge(jitter.width = 0.2,
                                      dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6
    ) +
    # ADD STARS
    geom_text(
      data = star_df,
      aes(x = Gene, y = y_star, label = stars_adj),
      inherit.aes = FALSE,
      size = 12
    ) +
    scale_fill_compendia() +
    scale_color_compendia() +
    coord_cartesian(ylim = c(0, 10)) +
    facet_wrap(~ PlotType, ncol = 2, scales = "free_x") +
    labs(
      title = paste(d),
      x = NULL,
      y = "Expression log2(TPM+1)"
    ) +
    color_theme() +
    theme(
      axis.title.y = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2),
      legend.text = element_text(size = 26),       # <-- label text size
      legend.title = element_text(size = 26),      # <-- title text size
      legend.key.size = unit(2, "cm") # <-- box/key size
    )
})
Fig1E
```

## Fig S4

```{r FigS4}
SS <- Fig1E[[1]]
aRMS <- Fig1E[[2]]
WT <- Fig1E[[3]]
NB <- Fig1E[[4]]
ALL <- Fig1E[[5]]
AML <- Fig1E[[6]]

FigS4 <- wrap_plots(SS, aRMS, WT, NB, ALL, AML, ncol = 1) +
  plot_layout(guides = "collect", axis_titles = "collect", axes = "collect") +
  plot_annotation(
    theme = theme(legend.position = "bottom")
  ) &
  theme(
    plot.margin = margin(t = 2, b = 2, l = 4, r = 4),
    plot.title = element_text(face = "bold", size = 30)
  )
FigS4

```

```{r save figs4}
ggsave("../../Figures/FigS4.png", FigS4, width = 20, height = 30, dpi = 500)
ggsave("../../Figures/FigS4.tif", FigS4, width = 20, height = 30, dpi = 500)
```

## Fig 1F

Histogram of median ratio distributions, calculated as the median expression level in PolyA samples divided by the median expression level in RiboD samples

```{r selecting gene ratios}
ss_ratios <- SS_gene_medians %>%
  select(Gene, Ratio = SS_median_ratio) %>%
  mutate(Disease = "SS")
write_tsv(ss_ratios, "../../output_data/Fig1E_F/ss_ratios.tsv.gz")


arms_ratios <- aRMS_gene_medians %>%
  select(Gene = Gene, Ratio = aRMS_median_ratio) %>%
  mutate(Disease = "aRMS")

wt_ratios <- WT_gene_medians %>%
  select(Gene, Ratio = WT_median_ratio) %>%
  mutate(Disease = "WT")

nb_ratios <- NB_gene_medians %>%
  select(Gene = Gene, Ratio = NB_median_ratio) %>%
  mutate(Disease = "NB")

all_ratios <- ALL_gene_medians %>%
  select(Gene = Gene, Ratio = ALL_median_ratio) %>%
  mutate(Disease = "ALL")

aml_ratios <- AML_gene_medians %>%
  select(Gene = Gene, Ratio = AML_median_ratio) %>%
  mutate(Disease = "AML")

#Combining median ratios
ratio_df <- bind_rows(ss_ratios, arms_ratios, wt_ratios, nb_ratios, aml_ratios, all_ratios)
```

```{r theme for fig1F}
theme_Fig1F <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "top",
      legend.title = element_blank(),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.4),
      axis.ticks = element_line(color = "black", linewidth = 0.4),
      strip.text = element_text(face = "bold", size = base_size * 0.9),
      plot.title = element_text(face = "bold", size = base_size * 1.1, hjust = 0.5),
      axis.title.y = element_text(angle = 90, hjust = 0.5, size = 15), 
      plot.margin = margin(10, 10, 10, 10)
    )
}
```

```{r finding HIST1H1B expression}
SS_HIST1H1B <- SS_gene_medians_hugo %>%
  filter(Gene %in% c("HIST1H1B"))
write_tsv(SS_HIST1H1B, "../../output_data/Fig1E_F/SS_HIST1H1B.tsv.gz")

aRMS_HIST1H1B <- aRMS_gene_median_hugo %>%
  filter(Gene %in% c("HIST1H1B"))

WT_HIST1H1B <- WT_gene_medians_hugo %>%
  filter(Gene %in% c("HIST1H1B"))

NB_HIST1H1B <- NB_gene_medians_hugo %>%
  filter(Gene %in% c("HIST1H1B"))

AML_HIST1H1B <- AML_gene_medians_hugo %>%
  filter(Gene %in% c("HIST1H1B"))

ALL_HIST1H1B <- ALL_gene_medians_hugo %>%
  filter(Gene %in% c("HIST1H1B"))
```

```{r Fig1F}
# Fig1F - SS standalone
Fig1F <- ggplot(ss_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  
  geom_vline(xintercept = SS_most_average_gene_hugo$SS_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = SS_most_average_gene_hugo$Gene), angle = 0, vjust = -0.4, hjust = 0.3, color = "#0072B2", size = 8) +
  
  geom_vline(xintercept = SS_HIST1H1B$SS_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = SS_HIST1H1B$Gene), angle = 0, vjust = -5, hjust = 0.2, color = "#E69F00", size = 8) +
  
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
    title = "SS",
    x = "Median Expression Ratio",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 22),
      axis.title.x = element_text(size = 22),
      axis.text.y = element_text(size = 18),
      axis.text.x = element_text(size = 18),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F
```

```{r save Fig1F}
ggsave("../../Figures/Fig1F.png", plot = Fig1F)
ggsave("../../Figures/Fig1F.tif", plot = Fig1F)
saveRDS(Fig1F, "../../Figures/Fig1F.rds")
```

```{r fig1F_aRMS}
Fig1F_aRMS <- ggplot(arms_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  geom_vline(xintercept = aRMS_most_average_gene_hugo$aRMS_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = aRMS_most_average_gene_hugo$Gene), angle = 0, vjust = -0.5, color = "#0072B2", size = 10) +
  
  geom_vline(xintercept = aRMS_HIST1H1B$aRMS_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = aRMS_HIST1H1B$Gene), angle = 0, vjust = -0.5, color = "#E69F00", size = 10) +
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
    title = "aRMS",
    x = "Median Expression Ratio (polyA / riboD)",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 32),
      axis.title.x = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F_aRMS
```

```{r fig1F_SS}
Fig1F_SS <- ggplot(ss_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  
  geom_vline(xintercept = SS_most_average_gene_hugo$SS_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = SS_most_average_gene_hugo$Gene), angle = 0, vjust = -0.5, color = "#0072B2", size = 10) +
  
  geom_vline(xintercept = SS_HIST1H1B$SS_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = SS_HIST1H1B$Gene), angle = 0, vjust = -0.5, color = "#E69F00", size = 10) +
  
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
    title = "SS",
    x = "Median Expression Ratio (polyA / riboD)",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 32),
      axis.title.x = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F_SS
```

```{r fig1f_WT}
Fig1F_WT <- ggplot(wt_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  
  geom_vline(xintercept = WT_most_average_gene_hugo$WT_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = WT_most_average_gene_hugo$Gene), angle = 0, vjust = -0.5, color = "#0072B2", size = 10) +
  
  geom_vline(xintercept = WT_HIST1H1B$WT_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = WT_HIST1H1B$Gene), angle = 0, vjust = -0.5, color = "#E69F00", size = 10) +
  
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
    title = "WT",
    x = "Median Expression Ratio (polyA / riboD)",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 32),
      axis.title.x = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F_WT
```

```{r fig1F_NB}
Fig1F_NB <- ggplot(nb_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  
  geom_vline(xintercept = NB_most_average_gene_hugo$NB_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = NB_most_average_gene_hugo$Gene), angle = 0, vjust = -0.5, color = "#0072B2", size = 10) +
  
  geom_vline(xintercept = NB_HIST1H1B$NB_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = NB_HIST1H1B$Gene), angle = 0, vjust = -0.5, color = "#E69F00", size = 10) +
  
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
    title = "NB",
    x = "Median Expression Ratio (polyA / riboD)",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 32),
      axis.title.x = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F_NB
```

```{r fig1F_ALL}
Fig1F_ALL <- ggplot(all_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  
  geom_vline(xintercept = ALL_most_average_gene_hugo$ALL_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = ALL_most_average_gene_hugo$Gene), angle = 0, vjust = -0.5, color = "#0072B2", size = 10) +
  
  geom_vline(xintercept = ALL_HIST1H1B$AML_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = ALL_HIST1H1B$Gene), angle = 0, vjust = -0.5, color = "#E69F00", size = 10) +
  
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
    title = "ALL",
    x = "Median Expression Ratio (polyA / riboD)",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 32),
      axis.title.x = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F_ALL
```

```{r fig1F_AML}
Fig1F_AML <- ggplot(aml_ratios, aes(x = Ratio)) +
  geom_histogram(binwidth = 0.1, alpha = 0.6, position = "identity") +
  
  geom_vline(xintercept = AML_most_average_gene_hugo$AML_median_ratio, color = "#0072B2", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 2.5, y = 2000, label = AML_most_average_gene_hugo$Gene), angle = 0, vjust = -0.5, color = "#0072B2", size = 10) +
  
  geom_vline(xintercept = AML_HIST1H1B$AML_median_ratio, color = "#E69F00", linetype = "dashed", linewidth = 2) +
  
  geom_text(aes(x = 0.9, y = 2000, label = AML_HIST1H1B$Gene), angle = 0, vjust = -0.5, color = "#E69F00", size = 10) +
  
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 4000)) +
  labs(
   title = "AML",
    x = "Median Expression Ratio (polyA / riboD)",
    y = "Number of Genes"
  ) +
  theme_Fig1F() +
      theme(
      axis.title.y = element_text(size = 32),
      axis.title.x = element_text(size = 32),
      axis.text.y = element_text(size = 28),
      axis.text.x = element_text(size = 28),
      strip.text = element_blank(),
      plot.title = element_text(vjust = -2)
    )

Fig1F_AML
```

## Fig S5

```{r FigS5}
FigS5 <- wrap_plots(Fig1F_SS, Fig1F_aRMS, Fig1F_WT, Fig1F_NB, Fig1F_ALL, Fig1F_AML, ncol = 1) +
  plot_layout(guides = "collect", axis_titles = "collect", axes = "collect") +
  plot_annotation(
    theme = theme(legend.position = "bottom")
  ) &
  theme(
    plot.margin = margin(t = 2, b = 2, l = 4, r = 4),
    plot.title = element_text(face = "bold", size = 30),
    legend.title = NULL
  )
FigS5
```

```{r save figs5}
ggsave("../../Figures/FigS5.png", FigS5, width = 20, height = 30, dpi = 500)
ggsave("../../Figures/FigS5.tif", FigS5, width = 20, height = 30, dpi = 500)
```

Session Info

```{r}
sessioninfo::session_info()
```
