---
title: "correlation_analysis_TEBA"
author: Maryke Grobler
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

# correlation_analysis

Maryke Grobler

## Correlation analysis

The goal of this analysis is to determine which samples are the most highly correlated within each disease type, then subset the gene expression to only include those samples. The subsetted expression file will be used for downstream differential expression analysis.

This is using gene expression from the Treehouse Tumor v25 RiboD and PolyA compendia. Diseases include synovial sarcoma (SS), alveolar rhabdomyosarcoma (aRMS), Wilms tumor (WT), neuroblastoma (NB), acute myeloid leukemia (AML), and acute lymphoblastic leukemia (ALL).

```{r install packages}
library(tidyverse)
library(corrr)
```

```{r metadata files}
polyA_clinical <- read_tsv("https://xena.treehouse.gi.ucsc.edu:443/download/clinical_Treehouse-Tumor-Compendium-25.01-PolyA_20250131v1.tsv")

riboD_clinical <- read_tsv("https://xena.treehouse.gi.ucsc.edu:443/download/clinical_Treehouse-Tumor-Compendium-25.01-RiboD_20250306v1.tsv")
```

```{r reformat}
# create column in each file that specifies with compendia the file is from
riboD_clinical_lib <- riboD_clinical %>%
  mutate(lib_prep="riboD")

polyA_clinical_lib <- polyA_clinical %>%
  mutate(lib_prep="polyA")

# merge clinical files
clin_merged <- bind_rows(polyA_clinical_lib, riboD_clinical_lib)
```

## SS and aRMS

```{r load in expression SS aRMS}
expression_SS_aRMS <- read_tsv("../input_data/expression_SS_aRMS.tsv.gz", show_col_types = FALSE)
```

```{r SS_aRMS}
SS_aRMS <- c("synovial sarcoma", "alveolar rhabdomyosarcoma")

clin_merged_SS_aRMS <- clin_merged %>%
  filter(disease %in% c(SS_aRMS)) %>%
  group_by(disease) %>%
  mutate(disease = case_when(
    disease == "synovial sarcoma" ~ "SS",
    disease == "alveolar rhabdomyosarcoma" ~ "aRMS"
    )) %>%
  unite(col = "disease_and_prep", disease, lib_prep, sep = "_") %>%
  select(th_dataset_id, disease_and_prep)

#write_tsv(clin_merged_SS_aRMS, "../data/clin_merged_SS_aRMS.tsv")
```

The datasets include some samples from the same patient, which may be skew the correlation scores since those samples would be more highly correlated to each other than any of the other samples.

```{r filter out duplicates SS aRMS}
# filtering out samples that contain S02 or S03
# in the clinical file
clin_SS_aRMS_dups_removed <- clin_merged_SS_aRMS %>%
  filter(!grepl("S02$", th_dataset_id),
         !grepl("S03$", th_dataset_id))

# in the expression file
expr_SS_aRMS_dups_removed <- expression_SS_aRMS %>%
  pivot_longer(cols = -Gene,
               names_to = "THID",
               values_to = "est_count") %>%
  filter(!grepl("S02$", THID),
         !grepl("S03$", THID)) %>%
  pivot_wider(names_from = "THID",
              values_from = "est_count")
```

```{r genes to row names SS aRMS}
# the first column in expr_SS_aRMS_dups_removed is "Gene", but the gene names should be the names of each row, not as a separate column.

cleaned_expr_SS_aRMS_dups_rem <- expr_SS_aRMS_dups_removed %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")
```

```{r filter out low genes SS aRMS}
# filtering out genes with a count less than 10 across all samples
filt_expr_SS_aRMS_dups_rem <- cleaned_expr_SS_aRMS_dups_rem %>%
  filter(if_all(everything(), ~.x > 10))
```

```{r separate files SS aRMS}
# filtering clinical file 
# DR means Duplicate samples Removed
clin_SS_polyA_DR <- clin_SS_aRMS_dups_removed %>%
  filter(disease_and_prep == "SS_polyA")

clin_SS_riboD_DR <- clin_SS_aRMS_dups_removed %>%
  filter(disease_and_prep == "SS_riboD")

clin_aRMS_polyA_DR <- clin_SS_aRMS_dups_removed %>%
  filter(disease_and_prep == "aRMS_polyA")

clin_aRMS_riboD_DR <- clin_SS_aRMS_dups_removed %>%
  filter(disease_and_prep == "aRMS_riboD")

# filtering expression file
expression_SS_polyA_DR <- filt_expr_SS_aRMS_dups_rem %>%
  select(all_of(c(clin_SS_polyA_DR$th_dataset_id)))

expression_SS_riboD_DR <- filt_expr_SS_aRMS_dups_rem %>%
  select(all_of(c(clin_SS_riboD_DR$th_dataset_id)))

expression_aRMS_polyA_DR <- filt_expr_SS_aRMS_dups_rem %>%
  select(all_of(c(clin_aRMS_polyA_DR$th_dataset_id)))

expression_aRMS_riboD_DR <- filt_expr_SS_aRMS_dups_rem %>%
  select(all_of(c(clin_aRMS_riboD_DR$th_dataset_id)))
```

```{r calculate correlation SS aRMS}
cor_SS_polyA_DR <- correlate(expression_SS_polyA_DR, method = "spearman", quiet = TRUE)
cor_SS_riboD_DR <- correlate(expression_SS_riboD_DR, method = "spearman", quiet = TRUE)
cor_aRMS_polyA_DR <- correlate(expression_aRMS_polyA_DR, method = "spearman", quiet = TRUE)
cor_aRMS_riboD_DR <- correlate(expression_aRMS_riboD_DR, method = "spearman", quiet = TRUE)
```

```{r stats for correlation SS polyA}
# SS_polyA
sum_cor_SS_polyA_DR <- cor_SS_polyA_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_SS_polyA_DR, n = 5)
```

```{r stats for correlation SS riboD}
# SS_riboD
sum_cor_SS_riboD_DR <- cor_SS_riboD_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise( #calculate the stats
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_SS_riboD_DR, n = 5)
```

```{r stats for correlation aRMS polyA}
# aRMS_polyA
# convert to long format, reshapes the data from wide (74 columns) to long format (one row per sample pair)
sum_cor_aRMS_polyA_DR <- cor_aRMS_polyA_DR %>%
  pivot_longer(cols = -term,
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_aRMS_polyA_DR, n = 5)
```

```{r stats for correlation aRMS riboD}
# aRMS_riboD
# convert to long format, reshapes the data from wide (74 columns) to long format (one row per sample pair)
sum_cor_aRMS_riboD_DR <- cor_aRMS_riboD_DR %>%
  pivot_longer(cols = -term,
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_aRMS_riboD_DR, n = 5)
```

```{r selecting highly correlated samples SS aRMS}
# this was a sample that was only 18th out of 20 most correlated samples, but it clustered together with the other SS riboD samples, so I'm including it
THR49_SS_riboD <- sum_cor_SS_riboD_DR %>%
  filter(term == "THR49_4787_S01") 

filtered_SS_riboD_list <- sum_cor_SS_riboD_DR %>%
  filter(term != "TH38_1433_S01") %>% # this was 10 out of 20 in the most correlated, but it clustered in the SS polyA samples, so I'm removing it
  slice_head(n = 14) %>% #taking the top most correlated samples
  rbind(THR49_SS_riboD) %>% #adding this THID 
  select(term) %>%
  mutate(disease_and_prep = "SS_riboD") # need to add back in the category

filtered_SS_polyA_list <- sum_cor_SS_polyA_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "SS_polyA")

filtered_aRMS_polyA_list <- sum_cor_aRMS_polyA_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "aRMS_polyA")

filtered_aRMS_riboD_list <- sum_cor_aRMS_riboD_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "aRMS_riboD")
```

```{r final SS aRMS list}
filtered_SS_aRMS_list <- bind_rows(filtered_SS_riboD_list, filtered_SS_polyA_list, filtered_aRMS_riboD_list, filtered_aRMS_polyA_list)

write_tsv(filtered_SS_aRMS_list, "../input_data/filtered_SS_aRMS_list.tsv")
```

## WT and NB

```{r load in expression WT NB}
expression_WT_NB <- read_tsv("../input_data/expression_WT_NB.tsv.gz", show_col_types = FALSE)
```

```{r WT_NB}
WT_NB <- c("wilms tumor", "neuroblastoma")

clin_merged_WT_NB <- clin_merged %>%
  filter(disease %in% c(WT_NB)) %>%
  group_by(disease) %>%
  mutate(disease = case_when(
    disease == "wilms tumor" ~ "WT",
    disease == "neuroblastoma" ~ "NB"
    )) %>%
  unite(col = "disease_and_prep", disease, lib_prep, sep = "_") %>%
  select(th_dataset_id, disease_and_prep) 

#write_tsv(clin_merged_WT_NB, "../data/clin_merged_WT_NB.tsv")
```

```{r filtering out duplicates WT NB}
# filtering out samples that contain S02, S03, S04
# in the clinical file
clin_WT_NB_dups_removed <- clin_merged_WT_NB %>%
  filter(!grepl("S02$", th_dataset_id), # the $ in a regular expression indicates the end of a line
         !grepl("S03$", th_dataset_id),
         !grepl("S04$", th_dataset_id))

# in the expression file
expr_WT_NB_dups_removed <- expression_WT_NB %>%
  pivot_longer(cols = -Gene,
               names_to = "THID",
               values_to = "est_count") %>%
  filter(!grepl("S02$", THID),
         !grepl("S03$", THID),
         !grepl("S04$", THID)) %>%
  pivot_wider(names_from = "THID",
              values_from = "est_count")
```

```{r genes to row names WT NB}
# the first column in expr_SS_aRMS_dups_removed is "Gene", but the gene names should be the names of each row, not as a separate column.

cleaned_expr_WT_NB_dups_removed <- expr_WT_NB_dups_removed %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")
```

```{r filter out low genes WT NB}
filt_expr_WT_NB_dups_rem <- cleaned_expr_WT_NB_dups_removed %>%
  filter(if_all(everything(), ~.x > 10))
```

```{r separate files WT NB}
# filtering clinical file 
# DR means Duplicate samples Removed
clin_WT_polyA_DR <- clin_WT_NB_dups_removed %>%
  filter(disease_and_prep == "WT_polyA")

clin_WT_riboD_DR <- clin_WT_NB_dups_removed %>%
  filter(disease_and_prep == "WT_riboD")

clin_NB_polyA_DR <- clin_WT_NB_dups_removed %>%
  filter(disease_and_prep == "NB_polyA")

clin_NB_riboD_DR <- clin_WT_NB_dups_removed %>%
  filter(disease_and_prep == "NB_riboD")

# filtering expression file
expression_WT_polyA_DR <- filt_expr_WT_NB_dups_rem %>%
  select(all_of(c(clin_WT_polyA_DR$th_dataset_id)))

expression_WT_riboD_DR <- filt_expr_WT_NB_dups_rem %>%
  select(all_of(c(clin_WT_riboD_DR$th_dataset_id)))

expression_NB_polyA_DR <- filt_expr_WT_NB_dups_rem %>%
  select(all_of(c(clin_NB_polyA_DR$th_dataset_id)))

expression_NB_riboD_DR <- filt_expr_WT_NB_dups_rem %>%
  select(all_of(c(clin_NB_riboD_DR$th_dataset_id)))
```

```{r calculate correlation WT NB}
# Run correlate()
cor_WT_polyA_DR <- correlate(expression_WT_polyA_DR, method = "spearman", quiet = TRUE)
cor_WT_riboD_DR <- correlate(expression_WT_riboD_DR, method = "spearman", quiet = TRUE)
cor_NB_polyA_DR <- correlate(expression_NB_polyA_DR, method = "spearman", quiet = TRUE)
cor_NB_riboD_DR <- correlate(expression_NB_riboD_DR, method = "spearman", quiet = TRUE)
```

```{r statistics WT polyA}
# WT_polyA
sum_cor_WT_polyA_DR <- cor_WT_polyA_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_WT_polyA_DR, n = 5)
```

```{r stats WT riboD}
# WT_riboD
sum_cor_WT_riboD_DR <- cor_WT_riboD_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise( #calculate the stats
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_WT_riboD_DR, n = 5)
```

```{r stats NB polyA}
# NB_polyA
sum_cor_NB_polyA_DR <- cor_NB_polyA_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_NB_polyA_DR, n = 5)
```

```{r stats NB riboD}
# NB_riboD
sum_cor_NB_riboD_DR <- cor_NB_riboD_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise( #calculate the stats
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_NB_riboD_DR, n = 5)
```

The top 15 most correlated samples were all in the same cluster in the tumormap, so there was minimal issue in creating the subsetted sample IDs.

```{r selecting highly correlated samples WT NB}
filtered_WT_polyA_list <- sum_cor_WT_polyA_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "WT_polyA")

filtered_WT_riboD_list <- sum_cor_WT_riboD_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "WT_riboD")

filtered_NB_polyA_list <- sum_cor_NB_polyA_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "NB_polyA")

filtered_NB_riboD_list <- sum_cor_NB_riboD_DR %>%
  slice_head(n = 15) %>%
  select(term) %>%
  mutate(disease_and_prep = "NB_riboD")
```

```{r combining and saving sample IDs WT NB}
filtered_WT_NB_list <- bind_rows(filtered_WT_riboD_list, filtered_WT_polyA_list, filtered_NB_riboD_list, filtered_NB_polyA_list)

write_tsv(filtered_WT_NB_list, "../input_data/filtered_WT_NB_list.tsv")
```

## ALL and AML

```{r load in expression}
expression_ALL_AML <- read_tsv("../input_data/expression_ALL_AML.tsv.gz", show_col_types = FALSE)
```

```{r ALL_AML}
ALL_AML <- c("acute lymphoblastic leukemia", "acute myeloid leukemia")

clin_merged_ALL_AML <- clin_merged %>%
  filter(disease %in% c(ALL_AML)) %>%
  group_by(disease) %>%
  mutate(disease = case_when(
    disease == "acute myeloid leukemia" ~ "AML",
    disease == "acute lymphoblastic leukemia" ~ "ALL"
    )) %>%
  unite(col = "disease_and_prep", disease, lib_prep, sep = "_") %>%
  select(th_dataset_id, disease_and_prep)

#write_tsv(clin_merged_ALL_AML, "../data/clin_merged_ALL_AML.tsv")
```

```{r remove duplicates and TARGET ALL samples}
# in the clinical file
clin_ALL_AML_noTARGET_dups_removed <- clin_merged_ALL_AML %>%
  filter(!grepl("TARGET", th_dataset_id),
         !grepl("S02$", th_dataset_id),
         !grepl("S03$", th_dataset_id),
         !grepl("S04$", th_dataset_id))

# in the expression file
expr_ALL_AML_noTARGET_dups_removed <- expression_ALL_AML %>%
  pivot_longer(cols = -Gene,
               names_to = "THID",
               values_to = "est_count") %>%
  filter(!grepl("TARGET", THID),
         !grepl("S02$", THID),
         !grepl("S03$", THID),
         !grepl("S04$", THID)) %>%
  pivot_wider(names_from = "THID",
              values_from = "est_count")
```

```{r genes to row names ALL AML}
cleaned_expr_ALL_AML_noTARGET_DR <- expr_ALL_AML_noTARGET_dups_removed %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")
```

```{r filter out low genes ALL AML}
filt_expr_ALL_AML_dups_rem <- cleaned_expr_ALL_AML_noTARGET_DR %>%
  filter(if_all(everything(), ~.x > 10))
```

```{r separate files ALL AML}
# filtering clinical file 
# DR means Duplicate samples Removed
clin_ALL_polyA_DR <- clin_ALL_AML_noTARGET_dups_removed %>%
  filter(disease_and_prep == "ALL_polyA")

clin_ALL_riboD_DR <- clin_ALL_AML_noTARGET_dups_removed %>%
  filter(disease_and_prep == "ALL_riboD")

clin_AML_polyA_DR <- clin_ALL_AML_noTARGET_dups_removed %>%
  filter(disease_and_prep == "AML_polyA")

clin_AML_riboD_DR <- clin_ALL_AML_noTARGET_dups_removed %>%
  filter(disease_and_prep == "AML_riboD")

# filtering expression file
expression_ALL_polyA_DR <- filt_expr_ALL_AML_dups_rem %>%
  select(all_of(c(clin_ALL_polyA_DR$th_dataset_id)))

expression_ALL_riboD_DR <- filt_expr_ALL_AML_dups_rem %>%
  select(all_of(c(clin_ALL_riboD_DR$th_dataset_id)))

expression_AML_polyA_DR <- filt_expr_ALL_AML_dups_rem %>%
  select(all_of(c(clin_AML_polyA_DR$th_dataset_id)))

expression_AML_riboD_DR <- filt_expr_ALL_AML_dups_rem %>%
  select(all_of(c(clin_AML_riboD_DR$th_dataset_id)))
```

```{r calculate correlation ALL AML }
cor_ALL_polyA_DR <- correlate(expression_ALL_polyA_DR, method = "spearman", quiet = TRUE)
cor_ALL_riboD_DR <- correlate(expression_ALL_riboD_DR, method = "spearman", quiet = TRUE)
cor_AML_polyA_DR <- correlate(expression_AML_polyA_DR, method = "spearman", quiet = TRUE)
cor_AML_riboD_DR <- correlate(expression_AML_riboD_DR, method = "spearman", quiet = TRUE)
```

```{r stats ALL polyA}
# ALL_polyA
sum_cor_ALL_polyA_DR <- cor_ALL_polyA_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_ALL_polyA_DR, n = 5)
```

```{r stats ALL riboD}
# ALL_riboD
sum_cor_ALL_riboD_DR <- cor_ALL_riboD_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_ALL_riboD_DR, n = 5)
```

```{r stats AML polyA}
# AML_polyA
sum_cor_AML_polyA_DR <- cor_AML_polyA_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_AML_polyA_DR, n = 5)
```

```{r stats AML riboD}
# AML_riboD
sum_cor_AML_riboD_DR <- cor_AML_riboD_DR %>%
  pivot_longer(cols = -term, # convert to long format, reshapes the data from wide (50 columns) to long format (one row per sample pair)
               names_to = "other_sample",
               values_to = "correlation") %>%
  filter(term != other_sample) %>%  #removes the self-correlations
  group_by(term) %>%
  summarise(
    min_value = min(correlation, na.rm = TRUE),
    max_value = max(correlation, na.rm = TRUE),
    median_value = median(correlation, na.rm = TRUE),
    mean_value = mean(correlation, na.rm = TRUE),
    stdev = sd(correlation, na.rm = TRUE)
  ) %>%
  arrange(desc(median_value))
head(sum_cor_AML_riboD_DR, n = 5)
```

```{r selecting highly correlated samples ALL AML}
filtered_ALL_polyA_list <- sum_cor_ALL_polyA_DR %>%
  slice_head(n = 20) %>%
  select(term) %>%
  mutate(disease_and_prep = "ALL_polyA")

filtered_ALL_riboD_list <- sum_cor_ALL_riboD_DR %>%
  slice_head(n = 20) %>% # taking the top most correlated samples
  select(term) %>%
  mutate(disease_and_prep = "ALL_riboD") # need to add back in the category

filtered_AML_polyA_list <- sum_cor_AML_polyA_DR %>%
  slice_head(n = 20) %>%
  select(term) %>%
  mutate(disease_and_prep = "AML_polyA")

filtered_AML_riboD_list <- sum_cor_AML_riboD_DR %>%
  slice_head(n = 20) %>%
  select(term) %>%
  mutate(disease_and_prep = "AML_riboD")
```

```{r combining and saving sample IDs ALL AML}
filtered_ALL_AML_list <- bind_rows(filtered_ALL_polyA_list, filtered_ALL_riboD_list, filtered_AML_polyA_list, filtered_AML_riboD_list)

write_tsv(filtered_ALL_AML_list, "../input_data/filtered_ALL_AML_list.tsv")
```

### Session Info

```{r session info}
sessioninfo::session_info()
```
