---
title: "ALL_AML_DESeq2"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## ALL v AML

This notebook is analyzing the expression profiles of acute myeloid leukemia (AML) and acute lymphoblastic leukemia (ALL) tumors from the v25 riboD and polyA compendia.

Goal: determine the extent that genes get called as differentially expressed as a result of the transcript enrichment method of the library preparation process.

Use: subsetted list of tumors for each disease in each library prep method (based on correlation analysis and tumor map analysis done before)

The comparisons we’ll be making (ALL = acute lymphoblastic leukemia, AML = acute myeloid leukemia):

- AML polyA vs ALL polyA

- AML riboD vs ALL riboD

- AML polyA vs ALL riboD

- AML riboD vs ALL polyA

```{r load libraries}
library(tidyverse)
library(DESeq2)
```

```{r}
# sample metadata
# This includes the top 20 samples from ALL polyA, ALL riboD, AML polyA, AML riboD
# The top 20 samples came from the correlation analysis, where this tsv file was generated
filtered_ALL_AML_list <- read_tsv("../../input_data/filtered_ALL_AML_list.tsv", show_col_types = FALSE)

expression_ALL_AML <- read_tsv("../../input_data/expression_ALL_AML.tsv.gz", show_col_types = FALSE)

gene_names <- read_tsv("../../input_data/EnsGeneID_Hugo_Observed_Conversions.txt", show_col_types = FALSE)
```

```{r}
# selecting only the 20 most correlated samples for each category,
# based on the metadata file
interested_expression <- expression_ALL_AML %>%
  select(all_of(c("Gene", filtered_ALL_AML_list$term)))
```

Since filtering out genes with a count less than 10 across all samples will filter out relevant genes, we’ll filter out genes with a count less than 10 in 11 or more samples (i.e., we’re keeping genes that have a count greater than 10 in at least 10 samples).

```{r}
# Explanation:
# across(where(is.numeric)) is selecting all the numeric columns
# i.e., all the sample columns. This deselects the Gene column, since that is a character column

# rowSums() counts how many samples meet the threshold for each gene,
# i.e., have a value of at least 10

# then keeps only the genes that meet that threshold in at least 10 samples
updated_expression <- interested_expression %>%
  filter(rowSums(across(where(is.numeric)) >= 10) >= 10)
```

### **AML polyA vs ALL polyA**

```{r}
# metadata
AML_polyA_ALL_polyA_meta <- filtered_ALL_AML_list %>%
  filter(disease_and_prep == "AML_polyA" | disease_and_prep == "ALL_polyA")
# the pipe | means we're taking all samples called ALL_riboD OR ALL_polyA in column disease_and_prep

# expression
AML_polyA_ALL_polyA_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", AML_polyA_ALL_polyA_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_AML_polyA_ALL_polyA_updated_expr <- AML_polyA_ALL_polyA_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_AML_polyA_ALL_polyA_meta <- AML_polyA_ALL_polyA_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_AML_polyA_ALL_polyA_meta) %in% colnames(fix_AML_polyA_ALL_polyA_updated_expr))

# check in order is the same
all(rownames(fix_AML_polyA_ALL_polyA_meta) == colnames(fix_AML_polyA_ALL_polyA_updated_expr))
```

```{r}
# format data
round_AML_polyA_ALL_polyA_updated_expr = round(fix_AML_polyA_ALL_polyA_updated_expr)

AML_polyA_ALL_polyA_updated_expr_matrix <- as.matrix(round_AML_polyA_ALL_polyA_updated_expr)

fix_AML_polyA_ALL_polyA_meta$disease_and_prep <- as.factor(fix_AML_polyA_ALL_polyA_meta$disease_and_prep)
```

```{r}
# set reference levels
# we want to make sure the references are consistent across the comparisons
levels(fix_AML_polyA_ALL_polyA_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_AML_polyA_ALL_polyA_updated <- DESeqDataSetFromMatrix(countData = AML_polyA_ALL_polyA_updated_expr_matrix,
                              colData = fix_AML_polyA_ALL_polyA_meta,
                              design = ~disease_and_prep)

dds_AML_polyA_ALL_polyA_updated
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
AML_polyA_ALL_polyA_updated_genes_to_keep <- rowSums(counts(dds_AML_polyA_ALL_polyA_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_AML_pA_ALL_pA_updated <- dds_AML_polyA_ALL_polyA_updated[which(AML_polyA_ALL_polyA_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_AML_pA_ALL_pA_updated_object <- DESeq(dds_AML_pA_ALL_pA_updated)
```

```{r}
# create results table
result_dds_AML_pA_ALL_pA_updated <- results(dds_AML_pA_ALL_pA_updated_object)
# summarize results table
summary(result_dds_AML_pA_ALL_pA_updated, alpha = 0.05)
```

```{r}
# convert results to dataframe
result_dds_AML_pA_ALL_pA_updated_df <- result_dds_AML_pA_ALL_pA_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
AML_polyA_ALL_polyA_LFC1great_padj <- result_dds_AML_pA_ALL_pA_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
AML_polyA_ALL_polyA_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
AML_polyA_ALL_polyA_LFC1less_padj <- result_dds_AML_pA_ALL_pA_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
AML_polyA_ALL_polyA_LFC1less_padj
```

```{r}
hugo_results_AML_polyA_ALL_polyA_updated <- result_dds_AML_pA_ALL_pA_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

# To check whether we lost any DEGs during the HugoID conversion,
# I repeated the above chunk to count the DEGs after converting to HugoIDs
# the numbers were the same

write_tsv(hugo_results_AML_polyA_ALL_polyA_updated, path = "../../output_data/ALL_AML/hugo_results_AML_polyA_ALL_polyA_updated.tsv.gz")

```

### **AML riboD vs ALL riboD**

```{r}
# metadata
AML_riboD_ALL_riboD_meta <- filtered_ALL_AML_list %>%
  filter(disease_and_prep == "AML_riboD" | disease_and_prep == "ALL_riboD")

# expression
AML_riboD_ALL_riboD_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", AML_riboD_ALL_riboD_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_AML_riboD_ALL_riboD_updated_expr <- AML_riboD_ALL_riboD_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_AML_riboD_ALL_riboD_meta <- AML_riboD_ALL_riboD_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_AML_riboD_ALL_riboD_meta) %in% colnames(fix_AML_riboD_ALL_riboD_updated_expr))

# check in order is the same
all(rownames(fix_AML_riboD_ALL_riboD_meta) == colnames(fix_AML_riboD_ALL_riboD_updated_expr))
```

```{r}
# format data
round_AML_riboD_ALL_riboD_updated_expr = round(fix_AML_riboD_ALL_riboD_updated_expr)

AML_riboD_ALL_riboD_updated_expr_matrix <- as.matrix(round_AML_riboD_ALL_riboD_updated_expr)

fix_AML_riboD_ALL_riboD_meta$disease_and_prep <- as.factor(fix_AML_riboD_ALL_riboD_meta$disease_and_prep)
```

```{r}
# set reference levels
# we want to make sure the references are consistent across the comparisons
levels(fix_AML_riboD_ALL_riboD_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_AML_riboD_ALL_riboD_updated <- DESeqDataSetFromMatrix(countData = AML_riboD_ALL_riboD_updated_expr_matrix,
                              colData = fix_AML_riboD_ALL_riboD_meta,
                              design = ~disease_and_prep)
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
AML_riboD_ALL_riboD_updated_genes_to_keep <- rowSums(counts(dds_AML_riboD_ALL_riboD_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_AML_riboD_ALL_riboD_updated <- dds_AML_riboD_ALL_riboD_updated[which(AML_riboD_ALL_riboD_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_AML_riboD_ALL_riboD_updated_object <- DESeq(dds_AML_riboD_ALL_riboD_updated)
```

```{r}
# create results table
results_AML_riboD_ALL_riboD_updated <- results(dds_AML_riboD_ALL_riboD_updated_object)

# summarize results table
summary(results_AML_riboD_ALL_riboD_updated, alpha = 0.05)
```

```{r}
# first we convert to df
results_AML_riboD_ALL_riboD_updated_df <- results_AML_riboD_ALL_riboD_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
AML_riboD_ALL_riboD_LFC1great_padj <- results_AML_riboD_ALL_riboD_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
AML_riboD_ALL_riboD_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
AML_riboD_ALL_riboD_LFC1less_padj <- results_AML_riboD_ALL_riboD_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
AML_riboD_ALL_riboD_LFC1less_padj
```

```{r}
hugo_results_AML_riboD_ALL_riboD_updated <- results_AML_riboD_ALL_riboD_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_AML_riboD_ALL_riboD_updated, path = "../../output_data/ALL_AML/hugo_results_AML_riboD_ALL_riboD_updated.tsv.gz")
```

### **AML polyA vs ALL riboD**

```{r}
# metadata
AML_polyA_ALL_riboD_meta <- filtered_ALL_AML_list %>%
  filter(disease_and_prep == "AML_polyA" | disease_and_prep == "ALL_riboD")
# the pipe | means we're taking all samples called ALL_riboD OR ALL_polyA in column disease_and_prep

# expression
AML_polyA_ALL_riboD_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", AML_polyA_ALL_riboD_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_AML_polyA_ALL_riboD_updated_expr <- AML_polyA_ALL_riboD_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_AML_polyA_ALL_riboD_meta <- AML_polyA_ALL_riboD_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_AML_polyA_ALL_riboD_meta) %in% colnames(fix_AML_polyA_ALL_riboD_updated_expr))

# check in order is the same
all(rownames(fix_AML_polyA_ALL_riboD_meta) == colnames(fix_AML_polyA_ALL_riboD_updated_expr))
```

```{r}
# format data
round_AML_polyA_ALL_riboD_updated_expr = round(fix_AML_polyA_ALL_riboD_updated_expr)

AML_polyA_ALL_riboD_updated_expr_matrix <- as.matrix(round_AML_polyA_ALL_riboD_updated_expr)

fix_AML_polyA_ALL_riboD_meta$disease_and_prep <- as.factor(fix_AML_polyA_ALL_riboD_meta$disease_and_prep)
```

```{r}
# check reference levels
levels(fix_AML_polyA_ALL_riboD_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_AML_polyA_ALL_riboD_updated <- DESeqDataSetFromMatrix(countData = AML_polyA_ALL_riboD_updated_expr_matrix,
                              colData = fix_AML_polyA_ALL_riboD_meta,
                              design = ~disease_and_prep)
dds_AML_polyA_ALL_riboD_updated
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
AML_polyA_ALL_riboD_updated_genes_to_keep <- rowSums(counts(dds_AML_polyA_ALL_riboD_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_AML_polyA_ALL_riboD_updated <- dds_AML_polyA_ALL_riboD_updated[which(AML_polyA_ALL_riboD_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_AML_polyA_ALL_riboD_updated_object <- DESeq(dds_AML_polyA_ALL_riboD_updated)
```

```{r}
# create results table
results_dds_AML_polyA_ALL_riboD_updated <- results(dds_AML_polyA_ALL_riboD_updated_object)
# summarize results table
summary(results_dds_AML_polyA_ALL_riboD_updated, alpha = 0.05)
```

```{r}
# convert results to dataframe
results_dds_AML_polyA_ALL_riboD_updated_df <- results_dds_AML_polyA_ALL_riboD_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
AML_polyA_ALL_riboD_LFC1great_padj <- results_dds_AML_polyA_ALL_riboD_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
AML_polyA_ALL_riboD_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
AML_polyA_ALL_riboD_LFC1less_padj <- results_dds_AML_polyA_ALL_riboD_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
AML_polyA_ALL_riboD_LFC1less_padj
```

```{r}
hugo_results_AML_polyA_ALL_riboD_updated <- results_dds_AML_polyA_ALL_riboD_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_AML_polyA_ALL_riboD_updated, path = "../../output_data/ALL_AML/hugo_results_AML_polyA_ALL_riboD_updated.tsv.gz")
```

### **AML riboD vs ALL polyA**

```{r}
# metadata
AML_riboD_ALL_polyA_meta <- filtered_ALL_AML_list %>%
  filter(disease_and_prep == "AML_riboD" | disease_and_prep == "ALL_polyA")

# expression
AML_riboD_ALL_polyA_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", AML_riboD_ALL_polyA_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_AML_riboD_ALL_polyA_updated_expr <- AML_riboD_ALL_polyA_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_AML_riboD_ALL_polyA_meta <- AML_riboD_ALL_polyA_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_AML_riboD_ALL_polyA_meta) %in% colnames(fix_AML_riboD_ALL_polyA_updated_expr))

# check in order is the same
all(rownames(fix_AML_riboD_ALL_polyA_meta) == colnames(fix_AML_riboD_ALL_polyA_updated_expr))
```

```{r}
# format data
round_AML_riboD_ALL_polyA_updated_expr = round(fix_AML_riboD_ALL_polyA_updated_expr)

AML_riboD_ALL_polyA_updated_expr_matrix <- as.matrix(round_AML_riboD_ALL_polyA_updated_expr)

fix_AML_riboD_ALL_polyA_meta$disease_and_prep <- as.factor(fix_AML_riboD_ALL_polyA_meta$disease_and_prep)
```

```{r}
# check reference levels
levels(fix_AML_riboD_ALL_polyA_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_AML_riboD_ALL_polyA_updated <- DESeqDataSetFromMatrix(countData = AML_riboD_ALL_polyA_updated_expr_matrix,
                              colData = fix_AML_riboD_ALL_polyA_meta,
                              design = ~disease_and_prep)
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
AML_riboD_ALL_polyA_updated_genes_to_keep <- rowSums(counts(dds_AML_riboD_ALL_polyA_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_AML_riboD_ALL_polyA_updated <- dds_AML_riboD_ALL_polyA_updated[which(AML_riboD_ALL_polyA_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_AML_riboD_ALL_polyA_updated_object <- DESeq(dds_AML_riboD_ALL_polyA_updated)
```

```{r}
# create results table
results_AML_riboD_ALL_polyA_updated <- results(dds_AML_riboD_ALL_polyA_updated_object)

# summarize results table
summary(results_AML_riboD_ALL_polyA_updated, alpha = 0.05)
```

```{r}
# first we convert to df
results_AML_riboD_ALL_polyA_updated_df <- results_AML_riboD_ALL_polyA_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
AML_riboD_ALL_polyA_LFC1great_padj <- results_AML_riboD_ALL_polyA_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
AML_riboD_ALL_polyA_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
AML_riboD_ALL_polyA_LFC1less_padj <- results_AML_riboD_ALL_polyA_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
AML_riboD_ALL_polyA_LFC1less_padj
```

```{r}
hugo_results_AML_riboD_ALL_polyA_updated <- results_AML_riboD_ALL_polyA_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_AML_riboD_ALL_polyA_updated, path = "../../output_data/ALL_AML/hugo_results_AML_riboD_ALL_polyA_updated.tsv.gz")
```

#### Session Info

```{r}
sessioninfo::session_info()
```
