---
title: "all_disease_comparisons"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## **Analysis of DESeq2 results across all disease comparisons**

This project is going to be analyzing the results from the differential expression analysis of the following disease comparisons, each with riboD and polyA samples:

SS versus aRMS

WT versus NB

ALL versus AML

The goal is to find common patterns of genes that are artifactually differentially expressed when comparing samples prepared by different transcript enrichment types.

```{r load packages}
library(tidyverse)
library(VennDiagram)
```

```{r load in files}
# load in results files from the SS:aRMS comparisons (aRMS is the reference)
armsPolyA_ssPolyA <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_polyA_SS_polyA_updated.tsv.gz")
armsRiboD_ssRiboD <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_riboD_SS_riboD_updated.tsv.gz")
armsPolyA_ssRiboD <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_polyA_SS_riboD_updated.tsv.gz")
armsRiboD_ssPolyA <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_riboD_SS_polyA_updated.tsv.gz")

# load in results from the WT:NB comparisons (NB is the reference)
nbPolyA_wtPolyA <- read_tsv("../../output_data/NB_WT/hugo_results_WT_polyA_NB_polyA_updated.tsv.gz") 
nbRiboD_wtRiboD <- read_tsv("../../output_data/NB_WT/hugo_results_WT_riboD_NB_riboD_updated.tsv.gz") 
nbPolyA_wtRiboD <- read_tsv("../../output_data/NB_WT/hugo_results_WT_riboD_NB_polyA_updated.tsv.gz")
nbRiboD_wtPolyA <- read_tsv("../../output_data/NB_WT/hugo_results_WT_polyA_NB_riboD_updated.tsv.gz")

# load in results from the ALL:AML comparisons (ALL is the reference)
allPolyA_amlPolyA <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_polyA_ALL_polyA_updated.tsv.gz")
allRiboD_amlRiboD <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_riboD_ALL_riboD_updated.tsv.gz")
allPolyA_amlRiboD <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_riboD_ALL_polyA_updated.tsv.gz")
allRiboD_amlPolyA <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_polyA_ALL_riboD_updated.tsv.gz")
```

### **Get significant up genes**

```{r SS aRMS up genes}
# for SS and aRMS
# Filter for genes with a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05
armsPolyA_ssPolyA_great1 <- armsPolyA_ssPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "armsPolyA_ssPolyA_great1") %>%
  relocate(comparison)

armsRiboD_ssRiboD_great1 <- armsRiboD_ssRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ARMSriboD_SSriboD_great1") %>%
  relocate(comparison)

armsRiboD_ssPolyA_great1 <- armsRiboD_ssPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ARMSriboD_SSpolyA_great1") %>%
  relocate(comparison)

armsPolyA_ssRiboD_great1 <- armsPolyA_ssRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ARMSpolyA_SSriboD_great1") %>%
  relocate(comparison)
```

```{r NB WT up genes}
# for WT and NB
# Filter for genes with a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05

nbPolyA_wtPolyA_great1 <- nbPolyA_wtPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "NBpolyA_WTpolyA_great1") %>%
  relocate(comparison)

nbRiboD_wtRiboD_great1 <- nbRiboD_wtRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "NBriboD_WTriboD_great1") %>%
  relocate(comparison)
 
nbRiboD_wtPolyA_great1 <- nbRiboD_wtPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "NBriboD_WTpolyA_great1") %>%
  relocate(comparison)

nbPolyA_wtRiboD_great1 <- nbPolyA_wtRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "NBpolyA_WTriboD_great1") %>%
  relocate(comparison)
```

```{r ALL AML up genes}
# for ALL and AML
# Filter for genes with a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05

allPolyA_amlPolyA_great1 <- allPolyA_amlPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ALLpolyA_AMLpolyA_great1") %>%
  relocate(comparison)

allRiboD_amlRiboD_great1 <- allRiboD_amlRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ALLriboD_AMLriboD_great1") %>%
  relocate(comparison)

allRiboD_amlPolyA_great1 <- allRiboD_amlPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ALLriboD_AMLpolyA_great1") %>%
  relocate(comparison)

allPolyA_amlRiboD_great1 <- allPolyA_amlRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "ALLpolyA_AMLriboD_great1") %>%
  relocate(comparison)
```

### **Unique gene lists - upregulated**

PolyA biased (in common between the polyAtruth and riboDtruth comparisons) and RiboD biased (in common between the polyAtruth and riboDtruth comparisons)

#### up in SS relative to aRMS

```{r armsRiboD_ssPolyA_great1}
# polyA biased for SS and aRMS comparison
# genes that are uniquely upregulated in ssPolyA relative to armsRiboD (polyA biased)
print(paste("total number of up genes in SS polyA relative to aRMS riboD (polyA biased) is", nrow(armsRiboD_ssPolyA_great1)))

up_arms_ss_PolyAunbiased_uniquetoPolyAbiased <- armsRiboD_ssPolyA_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>% # not in polyA unbiased
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") # not in riboD biased
print(paste("number of genes unique to polyA biased wrt polyAunbiased and riboD biased is", nrow(up_arms_ss_PolyAunbiased_uniquetoPolyAbiased)))

up_arms_ss_RiboDunbiased_uniquetoPolyAbiased <- armsRiboD_ssPolyA_great1 %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>% # not in riboD unbiased
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") # not in riboD biased
print(paste("number of genes unique to polyA biased wrt riboDunbiased and riboDbiased is", nrow(up_arms_ss_RiboDunbiased_uniquetoPolyAbiased)))
```

```{r up_uniquetoPolyAunbiased}
up_arms_ss_PolyAunbiased_uniquetoPolyAunbiased <- armsPolyA_ssPolyA_great1 %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene") %>% # not in polyA biased
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") # not in riboD biased
print(paste("number of genes unique to polyA unbiased wrt polyAbiased and riboD biased is", nrow(up_arms_ss_PolyAunbiased_uniquetoPolyAunbiased)))
```

```{r up_uniquetoRiboDunbiased}
up_arms_ss_RiboDunbiased_uniquetoRiboDunbiased <- armsRiboD_ssRiboD_great1 %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene") %>% # not in polyA biased
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") # not in riboD biased
print(paste("number of genes unique to RiboD unbiased wrt polyA biased and riboD biased is", nrow(up_arms_ss_RiboDunbiased_uniquetoRiboDunbiased)))
```

```{r up_polyAunbiased_riboDunbiased}
# are there any genes found to be differentially upregulated in the polyA unbiased and riboD unbiased that are not found in the biased comparisons?
up_arms_ss_polyAunbiased_riboDunbiased <- up_arms_ss_PolyAunbiased_uniquetoPolyAunbiased %>%
  inner_join(up_arms_ss_RiboDunbiased_uniquetoRiboDunbiased, by = "Gene")
print(paste("number of genes in both unique_to_polyAunbiased and unique_to_riboDunbiased is", nrow(up_arms_ss_polyAunbiased_riboDunbiased)))
```

```{r up_arms_ss_polyAbiased}
# need to combine these to get the genes in common between the two venn diagram polyA biased sections
up_arms_ss_polyAbiased <- inner_join(up_arms_ss_PolyAunbiased_uniquetoPolyAbiased, up_arms_ss_RiboDunbiased_uniquetoPolyAbiased, by = "Gene")
print(paste("number of genes in common as unique to polyA biased is", nrow(up_arms_ss_polyAbiased)))

# write_tsv(up_arms_ss_polyAbiased, "../../output_data/all_disease_comparisons/up_arms_ss_polyAbiased.tsv.gz")
```

```{r armsPolyA_ssRiboD_great1}
# riboD biased for SS and aRMS comparison
# genes that are uniquely upregulated in ssRiboD relative to armsPolyA (riboD biased)
print(paste("total number of genes in riboD_biased is", nrow(armsPolyA_ssRiboD_great1)))

up_arms_ss_polyAunbiased_uniquetoRiboDbiased <- armsPolyA_ssRiboD_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt polyAunbiased is", nrow(up_arms_ss_polyAunbiased_uniquetoRiboDbiased)))

up_arms_ss_riboDunbiased_uniquetoRiboDbiased <- armsPolyA_ssRiboD_great1 %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt riboDunbiased is", nrow(up_arms_ss_riboDunbiased_uniquetoRiboDbiased)))
```

```{r up_arms_ss_riboDbiased}
# need to combine these to get the ones in common between the two venn diagram riboD biased sections
up_arms_ss_riboDbiased <- inner_join(up_arms_ss_polyAunbiased_uniquetoRiboDbiased, up_arms_ss_riboDunbiased_uniquetoRiboDbiased, by = "Gene")

print(paste("number of genes unique to riboD biased is", nrow(up_arms_ss_riboDbiased)))

# write_tsv(up_arms_ss_riboDbiased, "../../output_data/all_disease_comparisons/up_arms_ss_riboDbiased.tsv.gz")
```

#### up in WT relative to NB

```{r nbRiboD_wtPolyA_great1}
# first false for WT and NB comparison
# genes that are uniquely upregulated in wtPolyA relative to nbRiboD (polyA biased)
print(paste("total number of genes in polyA biased is", nrow(nbRiboD_wtPolyA_great1)))

up_nb_wt_PolyAunbiased_uniquetoPolyAbiased <- nbRiboD_wtPolyA_great1 %>%
  anti_join(nbPolyA_wtPolyA_great1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_great1, by = "Gene")
print(paste("number of genes unique to polyAbiased wrt polyAunbiased is", nrow(up_nb_wt_PolyAunbiased_uniquetoPolyAbiased)))

up_nb_wt_RiboDunbiased_uniquetoPolyAbiased <- nbRiboD_wtPolyA_great1 %>%
  anti_join(nbRiboD_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_great1, by = "Gene")
print(paste("number of genes unique to polyAbiased wrt riboDunbiased is", nrow(up_nb_wt_RiboDunbiased_uniquetoPolyAbiased)))
```

```{r up_nb_wt_polyAbiased}
# need to combine these to get the ones in common between the two venn diagram first false sections
up_nb_wt_polyAbiased <- inner_join(up_nb_wt_PolyAunbiased_uniquetoPolyAbiased, up_nb_wt_RiboDunbiased_uniquetoPolyAbiased, by = "Gene")

print(paste("number of genes unique to polyAbiased is", nrow(up_nb_wt_polyAbiased)))

# write_tsv(up_nb_wt_polyAbiased, "../../output_data/all_disease_comparisons/up_nb_wt_polyAbiased.tsv.gz")
```

```{r nbPolyA_wtRiboD_great1}
# riboD biased for WT and NB comparison
# genes that are uniquely upregulated in wtRiboD relative to nbPolyA (riboD biased)
print(paste("total number of genes in riboD biased is", nrow(nbPolyA_wtRiboD_great1)))

up_nb_wt_polyAunbiased_uniquetoRiboDbiased <- nbPolyA_wtRiboD_great1 %>%
  anti_join(nbPolyA_wtPolyA_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_great1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt polyAunbiased is", nrow(up_nb_wt_polyAunbiased_uniquetoRiboDbiased)))

up_nb_wt_riboDunbiased_uniquetoRiboDbiased <- nbPolyA_wtRiboD_great1 %>%
  anti_join(nbRiboD_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_great1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt riboDunbiased is", nrow(up_nb_wt_riboDunbiased_uniquetoRiboDbiased)))
```

```{r up_nb_wt_riboDbiased}
# need to combine these to get the ones in common between the two venn diagram riboD biased sections
up_nb_wt_riboDbiased <- inner_join(up_nb_wt_polyAunbiased_uniquetoRiboDbiased, up_nb_wt_riboDunbiased_uniquetoRiboDbiased, by = "Gene")

print(paste("number of genes unique to riboD biased is", nrow(up_nb_wt_riboDbiased)))

# write_tsv(up_nb_wt_riboDbiased, "../../output_data/all_disease_comparisons/up_nb_wt_riboDbiased.tsv.gz")
```

#### up in AML relative to ALL

```{r allRiboD_amlPolyA_great1}
# polyA biased for ALL and AML comparison
# genes that are uniquely upregulated in amlPolyA relative to allRiboD (polyA biased)
print(paste("total number of genes in polyA biased is", nrow(allRiboD_amlPolyA_great1)))

up_all_aml_PolyAunbiased_uniquetoPolyAbiased <- allRiboD_amlPolyA_great1 %>%
  anti_join(allPolyA_amlPolyA_great1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_great1, by = "Gene")
print(paste("number of genes unique to polyA biased wrt polyAunbiased is", nrow(up_all_aml_PolyAunbiased_uniquetoPolyAbiased)))

up_all_aml_RiboDunbiased_uniquetoPolyAbiased <- allRiboD_amlPolyA_great1 %>%
  anti_join(allRiboD_amlRiboD_great1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_great1, by = "Gene")
print(paste("number of genes unique to polyA biased wrt riboDunbiased is", nrow(up_all_aml_RiboDunbiased_uniquetoPolyAbiased)))
```

```{r up_all_aml_polyAbiased}
# need to combine these to get the ones in common between the two venn diagram first false sections
up_all_aml_polyAbiased <- inner_join(up_all_aml_PolyAunbiased_uniquetoPolyAbiased, up_all_aml_RiboDunbiased_uniquetoPolyAbiased, by = "Gene")

print(paste("number of genes unique to polyA biased is", nrow(up_all_aml_polyAbiased)))

# write_tsv(up_all_aml_polyAbiased, "../../output_data/all_disease_comparisons/up_all_aml_polyAbiased.tsv.gz")
```

```{r allPolyA_amlRiboD_great1}
# riboD biased for ALL and AML comparison
# genes that are uniquely upregulated in amlRiboD relative to allPolyA (riboD biased)
print(paste("total number of genes in riboD biased is", nrow(allPolyA_amlRiboD_great1)))

up_all_aml_polyAunbiased_uniquetoRiboDbiased <- allPolyA_amlRiboD_great1 %>%
  anti_join(allPolyA_amlPolyA_great1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_great1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt polyAunbiased is", nrow(up_all_aml_polyAunbiased_uniquetoRiboDbiased)))

up_all_aml_riboDunbiased_uniquetoRiboDbiased <- allPolyA_amlRiboD_great1 %>%
  anti_join(allRiboD_amlRiboD_great1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_great1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt riboDunbiased is", nrow(up_all_aml_riboDunbiased_uniquetoRiboDbiased)))
```

```{r up_all_aml_riboDbiased}
# need to combine these to get the ones in common between the two venn diagram second false sections
up_all_aml_riboDbiased <- inner_join(up_all_aml_polyAunbiased_uniquetoRiboDbiased, up_all_aml_riboDunbiased_uniquetoRiboDbiased, by = "Gene")

print(paste("number of genes unique to riboD biased is", nrow(up_all_aml_riboDbiased)))

# write_tsv(up_all_aml_riboDbiased, "../../output_data/all_disease_comparisons/up_all_aml_riboDbiased.tsv.gz")
```

### Venn Diagrams

```{r up_polyAbiased_all_VD}
# up genes in the polyAbiased category of ss:arms, wt:nb, aml:all
up_polyAbiased_all <- list(
  "Up in SS polyA relative to aRMS riboD (polyA biased)" = up_arms_ss_polyAbiased$Gene,
  "Up in WT polyA relative to NB riboD (polyA biased)" = up_nb_wt_polyAbiased$Gene,
  "Up in AML polyA relative to ALL riboD (polyA biased)" = up_all_aml_polyAbiased$Gene
)

# write_rds(up_polyAbiased_all, "../../output_data/all_disease_comparisons/up_polyAbiased_all.rds")

up_polyAbiased_all_VD <- venn.diagram(
  x = up_polyAbiased_all,
  category.names = c(
    "Up in SS polyA relative to \n aRMS riboD (polyAbiased)",
    "Up in WT polyA relative to \n NB riboD (polyAbiased)",
    "Up in AML polyA relative to \n ALL riboD (polyAbiased)"
    ),
  filename = NULL, # Save as a rds file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#0072B2", "#0098ed", "#87CEEB"),
  # cat.col = c("#0072B2", "#0072B2", "#0072B2"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in PolyA disease relative to \n reference RiboD disease (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.70, # Font size for main title
  disable.logging = TRUE
)
up_polyAbiased_all_VD
```

```{r up_riboDbiased_all_VD}
# up genes in the riboD biased category of ss:arms, wt:nb, aml:all
up_riboDbiased_all <- list(
  "Up in SS riboD relative to aRMS polyA (riboD biased)" = up_arms_ss_riboDbiased$Gene,
  "Up in WT riboD relative to NB polyA (riboD biased)" = up_nb_wt_riboDbiased$Gene,
  "Up in AML riboD relative to ALL polyA (riboD biased)" = up_all_aml_riboDbiased$Gene
)

# write_rds(up_riboDbiased_all, "../../output_data/all_disease_comparisons/up_riboDbiased_all.rds")

up_riboDbiased_all_VD <- venn.diagram(
  x = up_riboDbiased_all,
  category.names = c(
    "Up in SS riboD relative to \n aRMS polyA (riboD biased)",
    "Up in WT riboD relative to \n NB polyA (riboD biased)",
    "Up in AML riboD relative to \n ALL polyA (riboD biased)"
    ),
  filename = NULL, # Save as a rds file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#E69F00", "#ff9b00", "#e6b200"),
  # cat.col = c("#E69F00", "#E69F00", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.60, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in RiboD disease relative to \n reference PolyA disease (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.70, # Font size for main title
  disable.logging = TRUE
)
up_riboDbiased_all_VD
```

```{r up in common}
# gene list for the genes in common between all three polyA_biased comparisons
up_polyAbiased_common <- up_arms_ss_polyAbiased %>%
  inner_join(up_nb_wt_polyAbiased, by = "Gene") %>%
  inner_join(up_all_aml_polyAbiased, by = "Gene")
print(nrow(up_polyAbiased_common))

# gene list for the genes in common between all three riboD_biased comparisons
up_riboDbiased_common <- up_arms_ss_riboDbiased %>%
  inner_join(up_nb_wt_riboDbiased, by = "Gene") %>%
  inner_join(up_all_aml_riboDbiased, by = "Gene")
print(nrow(up_riboDbiased_common))
```

### Get significant down genes

```{r aRMS SS down genes}
# for SS and aRMS
# Filter for genes with a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05
armsPolyA_ssPolyA_less1 <- armsPolyA_ssPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "armsPolyA_ssPolyA_less1") %>%
  relocate(comparison)

armsRiboD_ssRiboD_less1 <- armsRiboD_ssRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ARMSriboD_SSriboD_less1") %>%
  relocate(comparison)

armsRiboD_ssPolyA_less1 <- armsRiboD_ssPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ARMSriboD_SSpolyA_less1") %>%
  relocate(comparison)

armsPolyA_ssRiboD_less1 <- armsPolyA_ssRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ARMSpolyA_SSriboD_less1") %>%
  relocate(comparison)
```

```{r NB WT down genes}
# for WT and NB
# Filter for genes with a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05

nbPolyA_wtPolyA_less1 <- nbPolyA_wtPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "NBpolyA_WTpolyA_less1") %>%
  relocate(comparison)

nbRiboD_wtRiboD_less1 <- nbRiboD_wtRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "NBriboD_WTriboD_less1") %>%
  relocate(comparison)
 
nbRiboD_wtPolyA_less1 <- nbRiboD_wtPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "NBriboD_WTpolyA_less1") %>%
  relocate(comparison)

nbPolyA_wtRiboD_less1 <- nbPolyA_wtRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "NBpolyA_WTriboD_less1") %>%
  relocate(comparison)
```

```{r ALL AML down genes}
# for ALL and AML
# Filter for genes with a Log2 fold change less than or equal to 1 and have an adjusted p-value less than 0.05

allPolyA_amlPolyA_less1 <- allPolyA_amlPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ALLpolyA_AMLpolyA_less1") %>%
  relocate(comparison)

allRiboD_amlRiboD_less1 <- allRiboD_amlRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ALLriboD_AMLriboD_less1") %>%
  relocate(comparison)

allRiboD_amlPolyA_less1 <- allRiboD_amlPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ALLriboD_AMLpolyA_less1") %>%
  relocate(comparison)

allPolyA_amlRiboD_less1 <- allPolyA_amlRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "ALLpolyA_AMLriboD_less1") %>%
  relocate(comparison)
```

### **Unique gene lists - downregulated**

PolyA biased (in common between the polyAtruth and riboDtruth comparisons) and RiboD biased (in common between the polyAtruth and riboDtruth comparisons)

#### **Down in SS relative to aRMS**

```{r armsRiboD_ssPolyA_less1}
# polyA biased for SS and aRMS comparison
# genes that are uniquely downregulated in ssPolyA relative to armsRiboD (first false)

print(paste("total number of downregulated genes in SS polyA relative to aRMS riboD (polyA biased) is", nrow(armsRiboD_ssPolyA_less1)))

down_arms_ss_PolyAunbiased_uniquetoPolyAbiased <- armsRiboD_ssPolyA_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>% # not in polyA unbiased
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") # not in riboD biased
print(paste("number of genes unique to polyAbiased wrt polyAunbiased and riboDbiased is", nrow(down_arms_ss_PolyAunbiased_uniquetoPolyAbiased)))

down_arms_ss_RiboDunbiased_uniquetoPolyAbiased <- armsRiboD_ssPolyA_less1 %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>% # not in riboD unbiased
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") # not in riboD biased
print(paste("number of genes unique to polyAbiased wrt riboDunbiased and riboDbiased is", nrow(down_arms_ss_RiboDunbiased_uniquetoPolyAbiased)))
```

```{r}
# need to combine these to get the ones in common between the two venn diagram polyA biased sections
down_arms_ss_polyAbiased <- inner_join(down_arms_ss_PolyAunbiased_uniquetoPolyAbiased, down_arms_ss_RiboDunbiased_uniquetoPolyAbiased, by = "Gene")
print(paste("number of genes in common as unique to PolyA biased is", nrow(down_arms_ss_polyAbiased)))

# write_tsv(down_arms_ss_polyAbiased, "../../output_data/all_disease_comparisons/down_arms_ss_polyAbiased.tsv.gz")
```

```{r armsPolyA_ssRiboD_less1}
# riboDbiased for SS and aRMS comparison
# genes that are uniquely downregulated in ssRiboD relative to armsPolyA (riboDbiased)
print(paste("total number of genes in riboD biased is", nrow(armsPolyA_ssRiboD_less1)))

down_arms_ss_polyAunbiased_uniquetoRiboDbiased <- armsPolyA_ssRiboD_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt polyAunbiased is", nrow(down_arms_ss_polyAunbiased_uniquetoRiboDbiased)))

down_arms_ss_riboDunbiased_uniquetoRiboDbiased <- armsPolyA_ssRiboD_less1 %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt riboDunbiased is", nrow(down_arms_ss_riboDunbiased_uniquetoRiboDbiased)))
```

```{r down_arms_ss_riboDbiased}
# need to combine these to get the ones in common between the two venn diagram second false sections
down_arms_ss_riboDbiased <- inner_join(down_arms_ss_polyAunbiased_uniquetoRiboDbiased, down_arms_ss_riboDunbiased_uniquetoRiboDbiased, by = "Gene")
print(paste("number of genes unique to riboD biased is", nrow(down_arms_ss_riboDbiased)))

# write_tsv(down_arms_ss_riboDbiased, "../../output_data/all_disease_comparisons/down_arms_ss_riboDbiased.tsv.gz")
```

```{r down_uniquetoPolyAunbiased}
down_arms_ss_PolyAunbiased_uniquetoPolyAunbiased <- armsPolyA_ssPolyA_less1 %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene") %>% # not in polyA biased
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") # not in riboD biased
print(paste("number of genes uniquely downregulated in polyA unbiased wrt polyAbiased and riboD biased is", nrow(down_arms_ss_PolyAunbiased_uniquetoPolyAunbiased)))
```

```{r down_uniquetoRiboDunbiased}
down_arms_ss_RiboDunbiased_uniquetoRiboDunbiased <- armsRiboD_ssRiboD_less1 %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene") %>% # not in polyA biased
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") # not in riboD biased
print(paste("number of genes uniquely downregulated in riboD unbiased wrt polyAbiased and riboD biased is", nrow(down_arms_ss_RiboDunbiased_uniquetoRiboDunbiased)))
```

```{r}
# are there any genes found to be differentially downregulated in the polyA unbiased and riboD unbiased that are not found in the biased comparisons?
down_arms_ss_polyAunbiased_riboDunbiased <- down_arms_ss_PolyAunbiased_uniquetoPolyAunbiased %>%
  inner_join(down_arms_ss_RiboDunbiased_uniquetoRiboDunbiased, by = "Gene")
print(paste("number of genes in both unique_to_polyAunbiased and unique_to_riboDunbiased is", nrow(down_arms_ss_polyAunbiased_riboDunbiased)))
```

#### **Down in WT relative to NB**

```{r nbRiboD_wtPolyA_less1}
# polyA biased for WT and NB comparison
# genes that are uniquely downregulated in wtPolyA relative to nbRiboD
print(paste("total number of genes in polyA biased is", nrow(nbRiboD_wtPolyA_less1)))

down_nb_wt_PolyAunbiased_uniquetoPolyAbiased <- nbRiboD_wtPolyA_less1 %>%
  anti_join(nbPolyA_wtPolyA_less1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_less1, by = "Gene")
print(paste("number of genes unique to polyA biased wrt polyA unbiased is", nrow(down_nb_wt_PolyAunbiased_uniquetoPolyAbiased)))

down_nb_wt_RiboDunbiased_uniquetoPolyAbiased <- nbRiboD_wtPolyA_less1 %>%
  anti_join(nbRiboD_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_less1, by = "Gene")
print(paste("number of genes unique to polyA biased wrt riboD unbiased is", nrow(down_nb_wt_RiboDunbiased_uniquetoPolyAbiased)))
```

```{r down_nb_wt_polyAbiased}
# need to combine these to get the ones in common between the two venn diagram polyA biased sections
down_nb_wt_polyAbiased <- inner_join(down_nb_wt_PolyAunbiased_uniquetoPolyAbiased, down_nb_wt_RiboDunbiased_uniquetoPolyAbiased, by = "Gene")

print(paste("number of genes unique to polyA biased is", nrow(down_nb_wt_polyAbiased)))

# write_tsv(down_nb_wt_polyAbiased, "../../output_data/all_disease_comparisons/down_nb_wt_polyAbiased.tsv.gz")
```

```{r nbPolyA_wtRiboD_less1}
# riboD biased for WT and NB comparison
# genes that are uniquely downregulated in wtRiboD relative to nbPolyA (riboD biased)
print(paste("total number of genes in riboD biased is", nrow(nbPolyA_wtRiboD_less1)))

down_nb_wt_polyAunbiased_uniquetoRiboDbiased <- nbPolyA_wtRiboD_less1 %>%
  anti_join(nbPolyA_wtPolyA_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_less1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt polyAunbiased is", nrow(down_nb_wt_polyAunbiased_uniquetoRiboDbiased)))

down_nb_wt_riboDunbiased_uniquetoRiboDbiased <- nbPolyA_wtRiboD_less1 %>%
  anti_join(nbRiboD_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_less1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt riboDunbiased is", nrow(down_nb_wt_riboDunbiased_uniquetoRiboDbiased)))
```

```{r down_nb_wt_riboDbiased}
# need to combine these to get the ones in common between the two venn diagram second false sections
down_nb_wt_riboDbiased <- inner_join(down_nb_wt_polyAunbiased_uniquetoRiboDbiased, down_nb_wt_riboDunbiased_uniquetoRiboDbiased, by = "Gene")

print(paste("number of genes unique to riboD biased is", nrow(down_nb_wt_riboDbiased)))

# write_tsv(down_nb_wt_riboDbiased, "../../output_data/all_disease_comparisons/down_nb_wt_riboDbiased.tsv.gz")
```

#### **Down in AML relative to ALL**

```{r allRiboD_amlPolyA_less1}
# polyA biased for ALL and AML comparison
# genes that are uniquely downregulated in amlPolyA relative to allRiboD (first false)
print(paste("total number of genes in polyA biased is", nrow(allRiboD_amlPolyA_less1)))

down_all_aml_PolyAunbiased_uniquetoPolyAbiased <- allRiboD_amlPolyA_less1 %>%
  anti_join(allPolyA_amlPolyA_less1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_less1, by = "Gene")
print(paste("number of genes unique to polyA biased wrt polyAunbiased is", nrow(down_all_aml_PolyAunbiased_uniquetoPolyAbiased)))

down_all_aml_RiboDunbiased_uniquetoPolyAbiased <- allRiboD_amlPolyA_less1 %>%
  anti_join(allRiboD_amlRiboD_less1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_less1, by = "Gene")
print(paste("number of genes unique to polyA biased wrt riboDunbiased is", nrow(down_all_aml_RiboDunbiased_uniquetoPolyAbiased)))
```

```{r down_all_aml_polyAbiased}
# need to combine these to get the ones in common between the two venn diagram first false sections
down_all_aml_polyAbiased <- inner_join(down_all_aml_PolyAunbiased_uniquetoPolyAbiased, down_all_aml_RiboDunbiased_uniquetoPolyAbiased, by = "Gene")

print(paste("number of genes unique to polyA biased is", nrow(down_all_aml_polyAbiased)))

# write_tsv(down_all_aml_polyAbiased, "../../output_data/all_disease_comparisons/down_all_aml_polyAbiased.tsv.gz")
```

```{r allPolyA_amlRiboD_less1}
# riboD biased for ALL and AML comparison
# genes that are uniquely downregulated in amlRiboD relative to allPolyA
print(paste("total number of genes in riboD biased is", nrow(allPolyA_amlRiboD_less1)))

down_all_aml_polyAunbiased_uniquetoRiboDbiased <- allPolyA_amlRiboD_less1 %>%
  anti_join(allPolyA_amlPolyA_less1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_less1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt polyAunbiased is", nrow(down_all_aml_polyAunbiased_uniquetoRiboDbiased)))

down_all_aml_riboDunbiased_uniquetoRiboDbiased <- allPolyA_amlRiboD_less1 %>%
  anti_join(allRiboD_amlRiboD_less1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_less1, by = "Gene")
print(paste("number of genes unique to riboD biased wrt riboDunbiased is", nrow(down_all_aml_riboDunbiased_uniquetoRiboDbiased)))
```

```{r down_all_aml_riboDbiased}
# need to combine these to get the ones in common between the two venn diagram second false sections
down_all_aml_riboDbiased <- inner_join(down_all_aml_polyAunbiased_uniquetoRiboDbiased, down_all_aml_riboDunbiased_uniquetoRiboDbiased, by = "Gene")

print(paste("number of genes unique to riboD biased is", nrow(down_all_aml_riboDbiased)))

# write_tsv(down_all_aml_riboDbiased, "../../output_data/all_disease_comparisons/down_all_aml_riboDbiased.tsv.gz")
```

### Venn diagrams

```{r down_polyAbiased_all_VD}
# down genes in the first false category of ss:arms, wt:nb, aml:all
down_polyAbiased_all <- list(
  "Down SS polyA relative to aRMS riboD (polyA biased)" = down_arms_ss_polyAbiased$Gene,
  "Down in WT polyA relative to NB riboD (polyA biased)" = down_nb_wt_polyAbiased$Gene,
  "Down in AML polyA relative to ALL riboD (polyA biased)" = down_all_aml_polyAbiased$Gene
)

# write_rds(down_polyAbiased_all, "../../output_data/all_disease_comparisons/down_polyAbiased_all.rds")


down_polyAbiased_all_VD <- venn.diagram(
  x = down_polyAbiased_all,
  category.names = c(
    "Down in SS polyA relative to \n aRMS riboD (polyA biased)",
    "Down in WT polyA relative to \n NB riboD (polyA biased)",
    "Down in AML polyA relative to \n ALL riboD (polyA biased)"
    ),
  filename = NULL, # Save as a rds file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#0072B2", "#0098ed", "#87CEEB"),
  # cat.col = c("#0072B2", "#0072B2", "#0072B2"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in PolyA disease relative to \n reference RiboD disease (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.70, # Font size for main title
  disable.logging = TRUE
)
down_polyAbiased_all_VD
```

```{r down_riboDbiased_all_VD}
# down genes in the second false category of ss:arms, wt:nb, aml:all
down_riboDbiased_all <- list(
  "Down in SS riboD relative to aRMS polyA (second false)" = down_arms_ss_riboDbiased$Gene,
  "Down in WT riboD relative to NB polyA (second false)" = down_nb_wt_riboDbiased$Gene,
  "Down in AML riboD relative to ALL polyA (second false)" = down_all_aml_riboDbiased$Gene
)

# write_rds(down_riboDbiased_all, "../../output_data/all_disease_comparisons/down_riboDbiased_all.rds")


down_riboDbiased_all_VD <- venn.diagram(
  x = down_riboDbiased_all,
  category.names = c(
    "Down in SS riboD relative to \n aRMS polyA (riboD biased)",
    "Down in WT riboD relative to \n NB polyA (riboD biased)",
    "Down in AML riboD relative to \n ALL polyA (riboD biased)"
    ),
  filename = NULL, # Save as a rds file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#E69F00", "#ff9b00", "#e6b200"),
  # cat.col = c("#E69F00", "#E69F00", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.60, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in RiboD disease relative to \n reference PolyA disease (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.70, # Font size for main title
  disable.logging = TRUE
)
down_riboDbiased_all_VD
```

```{r down genes in common}
# gene list for the genes in common between all three polyA biased comparisons
down_polyAbiased_common <- down_arms_ss_polyAbiased %>%
  inner_join(down_nb_wt_polyAbiased, by = "Gene") %>%
  inner_join(down_all_aml_polyAbiased, by = "Gene")
print(nrow(down_polyAbiased_common))

# gene list for the genes in common between all three riboD biased comparisons
down_riboDbiased_common <- down_arms_ss_riboDbiased %>%
  inner_join(down_nb_wt_riboDbiased, by = "Gene") %>%
  inner_join(down_all_aml_riboDbiased, by = "Gene")
print(nrow(down_riboDbiased_common))
```

## **Visualizing venn diagram similarities**

### **Up genes in PolyAunbiased v biases**

Taking the venn diagram numbers (proportions of genes in each region) from DESeq2_results_analysis for each disease comparisons (SS:aRMS, WT:NB, ALL:AML), visualizing how the regions of the VDs are similar/different (upregulated genes)

```{r up_polyAunbiased_vs_biases}
up_polyAunbiased_vs_biases <- data.frame(
  region  = rep(c("polyA unbiased only", 
                  "polyA biased only", 
                  "riboD biased only",
                  "polyA unbiased & polyA biased",  
                  "polyA unbiased & riboD biased",  
                  "polyA biased & riboD biased",  
                  "polyA unbiased, polyA biased, riboD biased"), times = 3),
  disease_comparison = rep(c("up in SS relative to aRMS", "up in WT relative to NB", "up in AML relative to ALL"), each = 7),
  percentage   = c(
    # up in SS relative to aRMS
    1.37, 15.8, 45.1, 7.63, 12.4, 0.286, 17.4,
    # up in WT relative to NB
    1.6, 19.3, 39.9, 8.72, 15.8, 0.535, 14.1,
    # up in AML relative to ALL
    #5.17,  21.6, 28.7, 11.5, 9.5, 1.56, 22 # these were numbers from Anusha's old samples?
    2.52, 20.1, 24.6, 7.52, 25.1, 0.581, 19.5
  )
)
```

```{r fix region polyAunbiased}
# Fix region order for the x-axis
up_polyAunbiased_vs_biases$region <- factor(up_polyAunbiased_vs_biases$region,
                      levels = c("polyA unbiased only", 
                                 "polyA biased only", 
                                 "riboD biased only", 
                                 "polyA unbiased & polyA biased",  
                                 "polyA unbiased & riboD biased",  
                                 "polyA biased & riboD biased", 
                                 "polyA unbiased, polyA biased, riboD biased"))
```

```{r up_polyAunbiased_vs_biases_plot}
up_polyAunbiased_vs_biases_plot <- ggplot(up_polyAunbiased_vs_biases, aes(x = region, y = percentage, fill = disease_comparison)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_dodge(width = 0.75),
#            vjust = -0.4, 
            size = 2.8, 
            angle = 90,
            color = "grey30") +
  scale_fill_manual(values = c(
    "up in SS relative to aRMS" = "#f0eca3",
    "up in WT relative to NB" = "#cda5e6",
    "up in AML relative to ALL" = "#a1d6a1"
  )) +
  labs(
    title    = "Venn Diagram Region Percentages  - upregulated polyA unbiased vs biases",
    subtitle = "Each group of bars shows one intersection region across the three disease contexts",
    x        = "Venn Diagram Region",
    y        = "Percentage",
    fill     = "Disease comparison"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )

up_polyAunbiased_vs_biases_plot
```

```{r up_polyAunbiased_vs_biases_plot_bydisease}
up_polyAunbiased_vs_biases_plot_bydisease <- ggplot(up_polyAunbiased_vs_biases, aes(x = disease_comparison, y = percentage, fill = region)) +
  geom_col(position = position_stack(), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_stack(vjust = 0.5),
#            vjust = -0.4,
            size = 2.8,
            color = "black") +
  scale_fill_manual(values = c(
    "polyA unbiased only" = "#BB5566",
    "polyA biased only" = "#0072B2",
    "riboD biased only" = "#E69F00",
    "polyA unbiased & polyA biased" = "#8391b5",
    "polyA unbiased & riboD biased" = "#e2a86c",
    "polyA biased & riboD biased" = "#bfad80",
    "polyA unbiased, polyA biased, riboD biased" = "#b7986a"
  )) +
  labs(
    title    = "Percentage of upregulated genes in each region - polyA unbiased vs biases",
    subtitle = "Each bar shows the percentage of genes in each \nvenn diagram region for a given disease comparison",
    x        = "Disease comparison",
    y        = "Percentage",
    fill     = "Venn diagram region"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", 
                                    size = 15,
                                    margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )

up_polyAunbiased_vs_biases_plot_bydisease
```

### **Up genes in riboD unbiased v biases**

```{r up_riboDunbiased_vs_biases}
up_riboDunbiased_vs_biases <- data.frame(
  region  = rep(c("riboD unbiased only", 
                  "polyA biased only", 
                  "riboD biased only",
                  "riboD unbiased & polyA biased",
                  "riboD unbiased & riboD biased",  
                  "polyA biased & riboD biased",  
                  "riboD unbiased, polyA biased, riboD biased"), times = 3),
  disease_comparison = rep(c("up in SS relative to aRMS", "up in WT relative to NB", "up in AML relative to ALL"), each = 7),
  percentage   = c(
    # up in SS relative to aRMS
    0.962, 17.2, 48.4, 6.31, 9.32, 1.29, 16.5,
    # up in WT relative to NB
    1.2, 21.8, 46.6, 6.33, 9.33, 1.49, 13.2,
    # up in AML relative to ALL
    #4.53, 24.3, 29.5, 9.02, 8.95, 1.84, 21.8 - according to Anusha's old samples?
    1.25, 19, 44.4, 8.97, 5.99, 3.06, 17.3
  )
)
```

```{r}
# Fix region order for the x-axis
up_riboDunbiased_vs_biases$region <- factor(up_riboDunbiased_vs_biases$region,
                      levels = c("riboD unbiased only", 
                                 "polyA biased only", 
                                 "riboD biased only", 
                                 "riboD unbiased & polyA biased",  
                                 "riboD unbiased & riboD biased",  
                                 "polyA biased & riboD biased", 
                                 "riboD unbiased, polyA biased, riboD biased"))
```

```{r up_riboDunbiased_vs_biases_plot}
up_riboDunbiased_vs_biases_plot <- ggplot(up_riboDunbiased_vs_biases, aes(x = region, y = percentage, fill = disease_comparison)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_dodge(width = 0.75),
#            vjust = -0.4, 
            size = 2.8, 
            angle = 90,
            color = "grey30") +
  scale_fill_manual(values = c(
    "up in SS relative to aRMS" = "#f0eca3",
    "up in WT relative to NB" = "#cda5e6",
    "up in AML relative to ALL" = "#a1d6a1"
  )) +
  labs(
    title    = "Venn Diagram Region Percentages - upregulated riboD unbiased vs biases",
    subtitle = "Each group of bars shows one intersection region across the three disease contexts",
    x        = "Venn Diagram Region",
    y        = "Percentage",
    fill     = "Disease comparison"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )

up_riboDunbiased_vs_biases_plot
```

```{r up_riboDunbiased_vs_biases_plot_bydisease}
up_riboDunbiased_vs_biases_plot_bydisease <- ggplot(up_riboDunbiased_vs_biases, aes(x = disease_comparison, y = percentage, fill = region)) +
  geom_col(position = position_stack(), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_stack(vjust = 0.5),
#            vjust = -0.4,
            size = 2.8,
            color = "black") +
  scale_fill_manual(values = c(
    "riboD unbiased only" = "#BB5566",
    "polyA biased only" = "#0072B2",
    "riboD biased only" = "#E69F00",
    "riboD unbiased & polyA biased" = "#8391b5",
    "riboD unbiased & riboD biased" = "#e2a86c",
    "polyA biased & riboD biased" = "#bfad80",
    "riboD unbiased, polyA biased, riboD biased" = "#b7986a"
  )) +
  labs(
    title    = "Percentage of upregulated genes in each region - riboD unbiased vs biases",
    subtitle = "Each bar shows the percentage of genes in each \nvenn diagram region for a given disease comparison",
    x        = "Disease comparison",
    y        = "Percentage",
    fill     = "Venn diagram region"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", 
                                    size = 15,
                                    margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )

up_riboDunbiased_vs_biases_plot_bydisease
```

### **Down genes in PolyAunbiased v biases**

Taking the venn diagram numbers (proportions of genes in each region) from DESeq2_results_analysis for each disease comparisons (SS:aRMS, WT:NB, ALL:AML), visualizing how the regions of the VDs are similar/different (upregulated genes)

```{r down_polyAunbiased_vs_biases}
down_polyAunbiased_vs_biases <- data.frame(
  region  = rep(c("polyA unbiased only", 
                  "polyA biased only", 
                  "riboD biased only",
                  "polyA unbiased & polyA biased",  
                  "polyA unbiased & riboD biased",  
                  "polyA biased & riboD biased",  
                  "polyA unbiased, polyA biased, riboD biased"), times = 3),
  disease_comparison = rep(c("up in SS relative to aRMS", "up in WT relative to NB", "up in AML relative to ALL"), each = 7),
  percentage   = c(
    # down in SS relative to aRMS
    1.72, 36.2, 13.7, 20.1, 7.92, 0.448, 19.9,
    # down in WT relative to NB
    2.84, 35.2, 11.6, 20, 7.69, 0.472, 22.2,
    # down in AML relative to ALL
    3.17, 40.9, 15.3, 11.3, 11, 0.603, 17.7
  )
)
```

```{r}
# Fix region order for the x-axis
down_polyAunbiased_vs_biases$region <- factor(down_polyAunbiased_vs_biases$region,
                      levels = c("polyA unbiased only", 
                                 "polyA biased only", 
                                 "riboD biased only", 
                                 "polyA unbiased & polyA biased",  
                                 "polyA unbiased & riboD biased",  
                                 "polyA biased & riboD biased", 
                                 "polyA unbiased, polyA biased, riboD biased"))
```

```{r down_polyAunbiased_vs_biases_plot}
down_polyAunbiased_vs_biases_plot <- ggplot(down_polyAunbiased_vs_biases, aes(x = region, y = percentage, fill = disease_comparison)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_dodge(width = 0.75),
#            vjust = -0.4, 
            size = 2.8, 
            angle = 90,
            color = "grey30") +
  scale_fill_manual(values = c(
    "up in SS relative to aRMS" = "#f0eca3",
    "up in WT relative to NB" = "#cda5e6",
    "up in AML relative to ALL" = "#a1d6a1"
  )) +
  labs(
    title    = "Venn Diagram Region Percentages  - downregulated polyA unbiased vs biases",
    subtitle = "Each group of bars shows one intersection region across the three disease contexts",
    x        = "Venn Diagram Region",
    y        = "Percentage",
    fill     = "Disease comparison"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )
down_polyAunbiased_vs_biases_plot
```

```{r down_polyAunbiased_vs_biases_plot_bydisease}
down_polyAunbiased_vs_biases_plot_bydisease <- ggplot(down_polyAunbiased_vs_biases, aes(x = disease_comparison, y = percentage, fill = region)) +
  geom_col(position = position_stack(), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_stack(vjust = 0.5),
#            vjust = -0.4,
            size = 2.8,
            color = "black") +
  scale_fill_manual(values = c(
    "polyA unbiased only" = "#BB5566",
    "polyA biased only" = "#0072B2",
    "riboD biased only" = "#E69F00",
    "polyA unbiased & polyA biased" = "#8391b5",
    "polyA unbiased & riboD biased" = "#e2a86c",
    "polyA biased & riboD biased" = "#bfad80",
    "polyA unbiased, polyA biased, riboD biased" = "#b7986a"
  )) +
  labs(
    title    = "Percentage of downregulated genes in each region - polyA unbiased vs biases",
    subtitle = "Each bar shows the percentage of genes in each \nvenn diagram region for a given disease comparison",
    x        = "Disease comparison",
    y        = "Percentage",
    fill     = "Venn diagram region"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", 
                                    size = 15,
                                    margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )

down_polyAunbiased_vs_biases_plot_bydisease
```

### **Down genes in riboD unbiased v biases**

```{r down_riboDunbiased_vs_biases}
down_riboDunbiased_vs_biases <- data.frame(
  region  = rep(c("riboD unbiased only", 
                  "polyA biased only", 
                  "riboD biased only",
                  "riboD unbiased & polyA biased",
                  "riboD unbiased & riboD biased",  
                  "polyA biased & riboD biased",  
                  "riboD unbiased, polyA biased, riboD biased"), times = 3),
  disease_comparison = rep(c("up in SS relative to aRMS", "up in WT relative to NB", "up in AML relative to ALL"), each = 7),
  percentage   = c(
    # down in SS relative to aRMS
    0.498, 42, 18.6, 15, 3.25, 2.3, 18.3,
    # down in WT relative to NB
    1.02, 42.5, 14.9, 13.7, 4.76, 2.44, 20.6,
    # down in AML relative to ALL
    1.78, 40.5, 23.2, 12.5, 3.46, 1.55, 17.1
  )
)
```

```{r}
# Fix region order for the x-axis
down_riboDunbiased_vs_biases$region <- factor(down_riboDunbiased_vs_biases$region,
                      levels = c("riboD unbiased only", 
                                 "polyA biased only", 
                                 "riboD biased only", 
                                 "riboD unbiased & polyA biased",  
                                 "riboD unbiased & riboD biased",  
                                 "polyA biased & riboD biased", 
                                 "riboD unbiased, polyA biased, riboD biased"))
```

```{r down_riboDunbiased_vs_biases_plot}
down_riboDunbiased_vs_biases_plot <- ggplot(down_riboDunbiased_vs_biases, aes(x = region, y = percentage, fill = disease_comparison)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_dodge(width = 0.75),
#            vjust = -0.4, 
            size = 2.8, 
            angle = 90,
            color = "grey30") +
  scale_fill_manual(values = c(
    "up in SS relative to aRMS" = "#f0eca3",
    "up in WT relative to NB" = "#cda5e6",
    "up in AML relative to ALL" = "#a1d6a1"
  )) +
  labs(
    title    = "Venn Diagram Region Percentages - downregulated riboD unbiased vs biases",
    subtitle = "Each group of bars shows one intersection region across the three disease contexts",
    x        = "Venn Diagram Region",
    y        = "Percentage",
    fill     = "Disease comparison"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )

down_riboDunbiased_vs_biases_plot
```

```{r down_riboDunbiased_vs_biases_plot_bydisease}
down_riboDunbiased_vs_biases_plot_bydisease <- ggplot(down_riboDunbiased_vs_biases, aes(x = disease_comparison, y = percentage, fill = region)) +
  geom_col(position = position_stack(), width = 0.7) +
  geom_text(aes(label = percentage),
            position = position_stack(vjust = 0.5),
#            vjust = -0.4,
            size = 2.8,
            color = "black") +
  scale_fill_manual(values = c(
    "riboD unbiased only" = "#BB5566",
    "polyA biased only" = "#0072B2",
    "riboD biased only" = "#E69F00",
    "riboD unbiased & polyA biased" = "#8391b5",
    "riboD unbiased & riboD biased" = "#e2a86c",
    "polyA biased & riboD biased" = "#bfad80",
    "riboD unbiased, polyA biased, riboD biased" = "#b7986a"
  )) +
  labs(
    title    = "Percentage of downregulated genes in each region - riboD unbiased vs biases",
    subtitle = "Each bar shows the percentage of genes in each \nvenn diagram region for a given disease comparison",
    x        = "Disease comparison",
    y        = "Percentage",
    fill     = "Venn diagram region"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", 
                                    size = 15,
                                    margin = margin(b = 4)),
    plot.subtitle    = element_text(color = "grey45", size = 11, margin = margin(b = 12)),
    axis.text.x      = element_text(angle = 50, hjust = 1),
    panel.grid.major.x = element_blank(), #removes the major vertical grid lines from the plot panel
    legend.position  = "right",
    legend.title     = element_text(face = "bold")
  )
down_riboDunbiased_vs_biases_plot_bydisease
```

### Misc Analysis

Interested in knowing how many DEGs are unique across all 4 comparisons (polyA unbiased, riboD unbiased, polyA biased, and riboD biased).

#### aRMS v SS - Up DEGS across all 4 comparisons

```{r aRMS_SS_up_genes}
aRMS_SS_up_list <- list(armsPolyA_ssPolyA_great1, armsRiboD_ssRiboD_great1, armsPolyA_ssRiboD_great1, armsRiboD_ssPolyA_great1)

aRMS_SS_up_genes <- aRMS_SS_up_list %>%
  bind_rows() %>%
  summarize(distinct_genes = unique(Gene))
print(aRMS_SS_up_genes %>% nrow())
```

```{r aRMS_SS_up_uniques}
# genes unique to each comparison across all 4 comparisons

#polyA_unbiased
aRMS_SS_up_polyAunbiased_unique_v_all <- armsPolyA_ssPolyA_great1 %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene")
print(armsPolyA_ssPolyA_great1 %>% nrow())
print(aRMS_SS_up_polyAunbiased_unique_v_all %>% nrow())

#riboD_unbiased
aRMS_SS_up_riboDunbiased_unique_v_all <- armsRiboD_ssRiboD_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene")
print(armsRiboD_ssRiboD_great1 %>% nrow())
print(aRMS_SS_up_riboDunbiased_unique_v_all %>% nrow())

#polyA_biased
aRMS_SS_up_polyAbiased_unique_v_all <- armsRiboD_ssPolyA_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene") 
print(armsRiboD_ssPolyA_great1 %>% nrow())
print(aRMS_SS_up_polyAbiased_unique_v_all %>% nrow())

#riboD_biased
aRMS_SS_up_riboDbiased_unique_v_all <- armsPolyA_ssRiboD_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene") 
print(armsPolyA_ssRiboD_great1 %>% nrow())
print(aRMS_SS_up_riboDbiased_unique_v_all %>% nrow())
```

4-part venn diagram

```{r aRMS_SS_up_VD}
g1_aRMS_SS_up = armsPolyA_ssPolyA_great1$Gene
g2_aRMS_SS_up = armsRiboD_ssRiboD_great1$Gene
g3_aRMS_SS_up = armsRiboD_ssPolyA_great1$Gene
g4_aRMS_SS_up = armsPolyA_ssRiboD_great1$Gene

grid.newpage()
draw.quad.venn(
  area1 = length(g1_aRMS_SS_up), 
  area2 = length(g2_aRMS_SS_up), 
  area3 = length(g3_aRMS_SS_up), 
  area4 = length(g4_aRMS_SS_up),
  n12 = length(intersect(g1_aRMS_SS_up, g2_aRMS_SS_up)), 
  n13 = length(intersect(g1_aRMS_SS_up, g3_aRMS_SS_up)), 
  n14 = length(intersect(g1_aRMS_SS_up, g4_aRMS_SS_up)),
  n23 = length(intersect(g2_aRMS_SS_up, g3_aRMS_SS_up)), 
  n24 = length(intersect(g2_aRMS_SS_up, g4_aRMS_SS_up)), 
  n34 = length(intersect(g3_aRMS_SS_up, g4_aRMS_SS_up)),
  n123 = length(intersect(intersect(g1_aRMS_SS_up, g2_aRMS_SS_up), g3_aRMS_SS_up)),
  n124 = length(intersect(intersect(g1_aRMS_SS_up, g2_aRMS_SS_up), g4_aRMS_SS_up)),
  n134 = length(intersect(intersect(g1_aRMS_SS_up, g3_aRMS_SS_up), g4_aRMS_SS_up)),
  n234 = length(intersect(intersect(g2_aRMS_SS_up, g3_aRMS_SS_up), g4_aRMS_SS_up)),
  n1234 = length(intersect(intersect(intersect(g1_aRMS_SS_up, g2_aRMS_SS_up), g3_aRMS_SS_up), g4_aRMS_SS_up)),
  category = c("PolyA unbiased", "RiboD unbiased", "PolyA biased", "RiboD biased"),
  fill = c("#BB5566", "#D55E00", "#0072B2", "#E69F00"),
  print.mode = c("raw", "percent")
)
```

#### aRMS v SS - Down DEGS across all 4 comparisons

```{r aRMS_SS_down_genes}
aRMS_SS_down_list <- list(armsPolyA_ssPolyA_less1, armsRiboD_ssRiboD_less1, armsPolyA_ssRiboD_less1, armsRiboD_ssPolyA_less1)

aRMS_SS_down_genes <- aRMS_SS_down_list %>%
  bind_rows() %>%
  summarize(distinct_genes = unique(Gene))
print(aRMS_SS_down_genes %>% nrow())
```

```{r aRMS_SS_down_uniques}
# genes unique to each comparison across all 4 comparisons

#polyA_unbiased
aRMS_SS_down_polyAunbiased_unique_v_all <- armsPolyA_ssPolyA_less1 %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene")
print(armsPolyA_ssPolyA_less1 %>% nrow())
print(aRMS_SS_down_polyAunbiased_unique_v_all %>% nrow())

#riboD_unbiased
aRMS_SS_down_riboDunbiased_unique_v_all <- armsRiboD_ssRiboD_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene")
print(armsRiboD_ssRiboD_less1 %>% nrow())
print(aRMS_SS_down_riboDunbiased_unique_v_all %>% nrow())

#polyA_biased
aRMS_SS_down_polyAbiased_unique_v_all <- armsRiboD_ssPolyA_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene") 
print(armsRiboD_ssPolyA_less1 %>% nrow())
print(aRMS_SS_down_polyAbiased_unique_v_all %>% nrow())

#riboD_biased
aRMS_SS_down_riboDbiased_unique_v_all <- armsPolyA_ssRiboD_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene") 
print(armsPolyA_ssRiboD_less1 %>% nrow())
print(aRMS_SS_down_riboDbiased_unique_v_all %>% nrow())
```

```{r aRMS_SS_down_VD}
g1_aRMS_SS_down = armsPolyA_ssPolyA_less1$Gene
g2_aRMS_SS_down = armsRiboD_ssRiboD_less1$Gene
g3_aRMS_SS_down = armsRiboD_ssPolyA_less1$Gene
g4_aRMS_SS_down = armsPolyA_ssRiboD_less1$Gene


grid.newpage()
draw.quad.venn(
  area1 = length(g1_aRMS_SS_down), 
  area2 = length(g2_aRMS_SS_down), 
  area3 = length(g3_aRMS_SS_down), 
  area4 = length(g4_aRMS_SS_down),
  n12 = length(intersect(g1_aRMS_SS_down, g2_aRMS_SS_down)), 
  n13 = length(intersect(g1_aRMS_SS_down, g3_aRMS_SS_down)), 
  n14 = length(intersect(g1_aRMS_SS_down, g4_aRMS_SS_down)),
  n23 = length(intersect(g2_aRMS_SS_down, g3_aRMS_SS_down)), 
  n24 = length(intersect(g2_aRMS_SS_down, g4_aRMS_SS_down)), 
  n34 = length(intersect(g3_aRMS_SS_down, g4_aRMS_SS_down)),
  n123 = length(intersect(intersect(g1_aRMS_SS_down, g2_aRMS_SS_down), g3_aRMS_SS_down)),
  n124 = length(intersect(intersect(g1_aRMS_SS_down, g2_aRMS_SS_down), g4_aRMS_SS_down)),
  n134 = length(intersect(intersect(g1_aRMS_SS_down, g3_aRMS_SS_down), g4_aRMS_SS_down)),
  n234 = length(intersect(intersect(g2_aRMS_SS_down, g3_aRMS_SS_down), g4_aRMS_SS_down)),
  n1234 = length(intersect(intersect(intersect(g1_aRMS_SS_down, g2_aRMS_SS_down), g3_aRMS_SS_down), g4_aRMS_SS_down)),
  category = c("PolyA unbiased", "RiboD unbiased", "PolyA biased", "RiboD biased"),
  fill = c("#BB5566", "#D55E00", "#0072B2", "#E69F00"),
  print.mode = c("raw", "percent")
)
```

#### NB v WT - Up DEGS across all 4 comparisons

```{r NB_WT_up_genes}
NB_WT_up_list <- list(nbPolyA_wtPolyA_great1, nbRiboD_wtRiboD_great1, nbPolyA_wtRiboD_great1, nbRiboD_wtPolyA_great1)

NB_WT_up_genes <- NB_WT_up_list %>%
  bind_rows() %>%
  summarize(distinct_genes = unique(Gene))
print(NB_WT_up_genes %>% nrow())
```

```{r NB_WT_up_uniques}
# genes unique to each comparison across all 4 comparisons

#polyA_unbiased
NB_WT_up_polyAunbiased_unique_v_all <- nbPolyA_wtPolyA_great1 %>%
  anti_join(nbRiboD_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_great1, by = "Gene")
print(nbPolyA_wtPolyA_great1 %>% nrow())
print(NB_WT_up_polyAunbiased_unique_v_all %>% nrow())

#riboD_unbiased
NB_WT_up_riboDunbiased_unique_v_all <- nbRiboD_wtRiboD_great1 %>%
  anti_join(nbPolyA_wtPolyA_great1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_great1, by = "Gene")
print(nbRiboD_wtRiboD_great1 %>% nrow())
print(NB_WT_up_riboDunbiased_unique_v_all %>% nrow())

#polyA_biased
NB_WT_up_polyAbiased_unique_v_all <- nbRiboD_wtPolyA_great1 %>%
  anti_join(nbPolyA_wtPolyA_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_great1, by = "Gene") 
print(nbRiboD_wtPolyA_great1 %>% nrow())
print(NB_WT_up_polyAbiased_unique_v_all %>% nrow())

#riboD_biased
NB_WT_up_riboDbiased_unique_v_all <- nbPolyA_wtRiboD_great1 %>%
  anti_join(nbPolyA_wtPolyA_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtRiboD_great1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_great1, by = "Gene") 
print(nbPolyA_wtRiboD_great1 %>% nrow())
print(NB_WT_up_riboDbiased_unique_v_all %>% nrow())
```

```{r NB_WT_up_VD}
g1_NB_WT_up = nbPolyA_wtPolyA_great1$Gene
g2_NB_WT_up = nbRiboD_wtRiboD_great1$Gene
g3_NB_WT_up = nbRiboD_wtPolyA_great1$Gene
g4_NB_WT_up = nbPolyA_wtRiboD_great1$Gene


grid.newpage()
draw.quad.venn(
  area1 = length(g1_NB_WT_up), 
  area2 = length(g2_NB_WT_up), 
  area3 = length(g3_NB_WT_up), 
  area4 = length(g4_NB_WT_up),
  n12 = length(intersect(g1_NB_WT_up, g2_NB_WT_up)), 
  n13 = length(intersect(g1_NB_WT_up, g3_NB_WT_up)), 
  n14 = length(intersect(g1_NB_WT_up, g4_NB_WT_up)),
  n23 = length(intersect(g2_NB_WT_up, g3_NB_WT_up)), 
  n24 = length(intersect(g2_NB_WT_up, g4_NB_WT_up)), 
  n34 = length(intersect(g3_NB_WT_up, g4_NB_WT_up)),
  n123 = length(intersect(intersect(g1_NB_WT_up, g2_NB_WT_up), g3_NB_WT_up)),
  n124 = length(intersect(intersect(g1_NB_WT_up, g2_NB_WT_up), g4_NB_WT_up)),
  n134 = length(intersect(intersect(g1_NB_WT_up, g3_NB_WT_up), g4_NB_WT_up)),
  n234 = length(intersect(intersect(g2_NB_WT_up, g3_NB_WT_up), g4_NB_WT_up)),
  n1234 = length(intersect(intersect(intersect(g1_NB_WT_up, g2_NB_WT_up), g3_NB_WT_up), g4_NB_WT_up)),
  category = c("PolyA unbiased", "RiboD unbiased", "PolyA biased", "RiboD biased"),
  fill = c("#BB5566", "#D55E00", "#0072B2", "#E69F00"),
  print.mode = c("raw", "percent")
)
```

#### NB v WT - Down DEGS across all 4 comparisons

```{r NB_WT_down_genes}
NB_WT_down_list <- list(nbPolyA_wtPolyA_less1, nbRiboD_wtRiboD_less1, nbPolyA_wtRiboD_less1, nbRiboD_wtPolyA_less1)

NB_WT_down_genes <- NB_WT_down_list %>%
  bind_rows() %>%
  summarize(distinct_genes = unique(Gene))
print(NB_WT_down_genes %>% nrow())
```

```{r NB_WT_down_uniques}
# genes unique to each comparison across all 4 comparisons

#polyA_unbiased
NB_WT_down_polyAunbiased_unique_v_all <- nbPolyA_wtPolyA_less1 %>%
  anti_join(nbRiboD_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_less1, by = "Gene")
print(nbPolyA_wtPolyA_less1 %>% nrow())
print(NB_WT_down_polyAunbiased_unique_v_all %>% nrow())

#riboD_unbiased
NB_WT_down_riboDunbiased_unique_v_all <- nbRiboD_wtRiboD_less1 %>%
  anti_join(nbPolyA_wtPolyA_less1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_less1, by = "Gene")
print(nbRiboD_wtRiboD_less1 %>% nrow())
print(NB_WT_down_riboDunbiased_unique_v_all %>% nrow())

#polyA_biased
NB_WT_down_polyAbiased_unique_v_all <- nbRiboD_wtPolyA_less1 %>%
  anti_join(nbPolyA_wtPolyA_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbPolyA_wtRiboD_less1, by = "Gene") 
print(nbRiboD_wtPolyA_less1 %>% nrow())
print(NB_WT_down_polyAbiased_unique_v_all %>% nrow())

#riboD_biased
NB_WT_down_riboDbiased_unique_v_all <- nbPolyA_wtRiboD_less1 %>%
  anti_join(nbPolyA_wtPolyA_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtRiboD_less1, by = "Gene") %>%
  anti_join(nbRiboD_wtPolyA_less1, by = "Gene") 
print(nbPolyA_wtRiboD_less1 %>% nrow())
print(NB_WT_down_riboDbiased_unique_v_all %>% nrow())
```

```{r NB_WT_down_VD}
g1_NB_WT_down = nbPolyA_wtPolyA_less1$Gene
g2_NB_WT_down = nbRiboD_wtRiboD_less1$Gene
g3_NB_WT_down = nbRiboD_wtPolyA_less1$Gene
g4_NB_WT_down = nbPolyA_wtRiboD_less1$Gene


grid.newpage()
draw.quad.venn(
  area1 = length(g1_NB_WT_down), 
  area2 = length(g2_NB_WT_down), 
  area3 = length(g3_NB_WT_down), 
  area4 = length(g4_NB_WT_down),
  n12 = length(intersect(g1_NB_WT_down, g2_NB_WT_down)), 
  n13 = length(intersect(g1_NB_WT_down, g3_NB_WT_down)), 
  n14 = length(intersect(g1_NB_WT_down, g4_NB_WT_down)),
  n23 = length(intersect(g2_NB_WT_down, g3_NB_WT_down)), 
  n24 = length(intersect(g2_NB_WT_down, g4_NB_WT_down)), 
  n34 = length(intersect(g3_NB_WT_down, g4_NB_WT_down)),
  n123 = length(intersect(intersect(g1_NB_WT_down, g2_NB_WT_down), g3_NB_WT_down)),
  n124 = length(intersect(intersect(g1_NB_WT_down, g2_NB_WT_down), g4_NB_WT_down)),
  n134 = length(intersect(intersect(g1_NB_WT_down, g3_NB_WT_down), g4_NB_WT_down)),
  n234 = length(intersect(intersect(g2_NB_WT_down, g3_NB_WT_down), g4_NB_WT_down)),
  n1234 = length(intersect(intersect(intersect(g1_NB_WT_down, g2_NB_WT_down), g3_NB_WT_down), g4_NB_WT_down)),
  category = c("PolyA unbiased", "RiboD unbiased", "PolyA biased", "RiboD biased"),
  fill = c("#BB5566", "#D55E00", "#0072B2", "#E69F00"),
  print.mode = c("raw", "percent")
)
```

#### ALL v AML - Up DEGS across all 4 comparisons

```{r ALL_AML_up_genes}
ALL_AML_up_list <- list(allPolyA_amlPolyA_great1, allRiboD_amlRiboD_great1, allPolyA_amlRiboD_great1, allRiboD_amlPolyA_great1)

ALL_AML_up_genes <- ALL_AML_up_list %>%
  bind_rows() %>%
  summarize(distinct_genes = unique(Gene))
print(ALL_AML_up_genes %>% nrow())
```

```{r ALL_AML_up_uniques}
# genes unique to each comparison across all 4 comparisons

#polyA_unbiased
ALL_AML_up_polyAunbiased_unique_v_all <- allPolyA_amlPolyA_great1 %>%
  anti_join(allRiboD_amlRiboD_great1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_great1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_great1, by = "Gene")
print(allPolyA_amlPolyA_great1 %>% nrow())
print(ALL_AML_up_polyAunbiased_unique_v_all %>% nrow())

#riboD_unbiased
ALL_AML_up_riboDunbiased_unique_v_all <- allRiboD_amlRiboD_great1 %>%
  anti_join(allPolyA_amlPolyA_great1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_great1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_great1, by = "Gene")
print(allRiboD_amlRiboD_great1 %>% nrow())
print(ALL_AML_up_riboDunbiased_unique_v_all %>% nrow())

#polyA_biased
ALL_AML_up_polyAbiased_unique_v_all <- allRiboD_amlPolyA_great1 %>%
  anti_join(allPolyA_amlPolyA_great1, by = "Gene") %>%
  anti_join(allRiboD_amlRiboD_great1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_great1, by = "Gene") 
print(allRiboD_amlPolyA_great1 %>% nrow())
print(ALL_AML_up_polyAbiased_unique_v_all %>% nrow())

#riboD_biased
ALL_AML_up_riboDbiased_unique_v_all <- allPolyA_amlRiboD_great1 %>%
  anti_join(allPolyA_amlPolyA_great1, by = "Gene") %>%
  anti_join(allRiboD_amlRiboD_great1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_great1, by = "Gene") 
print(allPolyA_amlRiboD_great1 %>% nrow())
print(ALL_AML_up_riboDbiased_unique_v_all %>% nrow())
```

```{r ALL_AML_up_VD}
g1_all_aml_up = allPolyA_amlPolyA_great1$Gene
g2_all_aml_up = allRiboD_amlRiboD_great1$Gene
g3_all_aml_up = allRiboD_amlPolyA_great1$Gene
g4_all_aml_up = allPolyA_amlRiboD_great1$Gene


grid.newpage()
draw.quad.venn(
  area1 = length(g1_all_aml_up), 
  area2 = length(g2_all_aml_up), 
  area3 = length(g3_all_aml_up), 
  area4 = length(g4_all_aml_up),
  n12 = length(intersect(g1_all_aml_up, g2_all_aml_up)), 
  n13 = length(intersect(g1_all_aml_up, g3_all_aml_up)), 
  n14 = length(intersect(g1_all_aml_up, g4_all_aml_up)),
  n23 = length(intersect(g2_all_aml_up, g3_all_aml_up)), 
  n24 = length(intersect(g2_all_aml_up, g4_all_aml_up)), 
  n34 = length(intersect(g3_all_aml_up, g4_all_aml_up)),
  n123 = length(intersect(intersect(g1_all_aml_up, g2_all_aml_up), g3_all_aml_up)),
  n124 = length(intersect(intersect(g1_all_aml_up, g2_all_aml_up), g4_all_aml_up)),
  n134 = length(intersect(intersect(g1_all_aml_up, g3_all_aml_up), g4_all_aml_up)),
  n234 = length(intersect(intersect(g2_all_aml_up, g3_all_aml_up), g4_all_aml_up)),
  n1234 = length(intersect(intersect(intersect(g1_all_aml_up, g2_all_aml_up), g3_all_aml_up), g4_all_aml_up)),
  category = c("PolyA unbiased", "RiboD unbiased", "PolyA biased", "RiboD biased"),
  fill = c("#BB5566", "#D55E00", "#0072B2", "#E69F00"),
  print.mode = c("raw", "percent")
)
```

#### ALL v AML - Down DEGS across all 4 comparisons

```{r ALL_AML_down_genes}
ALL_AML_down_list <- list(allPolyA_amlPolyA_less1, allRiboD_amlRiboD_less1, allPolyA_amlRiboD_less1, allRiboD_amlPolyA_less1)

ALL_AML_down_genes <- ALL_AML_down_list %>%
  bind_rows() %>%
  summarize(distinct_genes = unique(Gene))
print(ALL_AML_down_genes %>% nrow())
```

```{r ALL_AML_down_uniques}
# genes unique to each comparison across all 4 comparisons

#polyA_unbiased
ALL_AML_down_polyAunbiased_unique_v_all <- allPolyA_amlPolyA_less1 %>%
  anti_join(allRiboD_amlRiboD_less1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_less1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_less1, by = "Gene")
print(allPolyA_amlPolyA_less1 %>% nrow())
print(ALL_AML_down_polyAunbiased_unique_v_all %>% nrow())

#riboD_unbiased
ALL_AML_down_riboDunbiased_unique_v_all <- allRiboD_amlRiboD_less1 %>%
  anti_join(allPolyA_amlPolyA_less1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_less1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_less1, by = "Gene")
print(allRiboD_amlRiboD_less1 %>% nrow())
print(ALL_AML_down_riboDunbiased_unique_v_all %>% nrow())

#polyA_biased
ALL_AML_down_polyAbiased_unique_v_all <- allRiboD_amlPolyA_less1 %>%
  anti_join(allPolyA_amlPolyA_less1, by = "Gene") %>%
  anti_join(allRiboD_amlRiboD_less1, by = "Gene") %>%
  anti_join(allPolyA_amlRiboD_less1, by = "Gene") 
print(allRiboD_amlPolyA_less1 %>% nrow())
print(ALL_AML_down_polyAbiased_unique_v_all %>% nrow())

#riboD_biased
ALL_AML_down_riboDbiased_unique_v_all <- allPolyA_amlRiboD_less1 %>%
  anti_join(allPolyA_amlPolyA_less1, by = "Gene") %>%
  anti_join(allRiboD_amlRiboD_less1, by = "Gene") %>%
  anti_join(allRiboD_amlPolyA_less1, by = "Gene") 
print(allPolyA_amlRiboD_less1 %>% nrow())
print(ALL_AML_down_riboDbiased_unique_v_all %>% nrow())
```

```{r ALL_AML_down_VD}
g1_all_aml_down = allPolyA_amlPolyA_less1$Gene
g2_all_aml_down = allRiboD_amlRiboD_less1$Gene
g3_all_aml_down = allRiboD_amlPolyA_less1$Gene
g4_all_aml_down = allPolyA_amlRiboD_less1$Gene


grid.newpage()
draw.quad.venn(
  area1 = length(g1_all_aml_down), 
  area2 = length(g2_all_aml_down), 
  area3 = length(g3_all_aml_down), 
  area4 = length(g4_all_aml_down),
  n12 = length(intersect(g1_all_aml_down, g2_all_aml_down)), 
  n13 = length(intersect(g1_all_aml_down, g3_all_aml_down)), 
  n14 = length(intersect(g1_all_aml_down, g4_all_aml_down)),
  n23 = length(intersect(g2_all_aml_down, g3_all_aml_down)), 
  n24 = length(intersect(g2_all_aml_down, g4_all_aml_down)), 
  n34 = length(intersect(g3_all_aml_down, g4_all_aml_down)),
  n123 = length(intersect(intersect(g1_all_aml_down, g2_all_aml_down), g3_all_aml_down)),
  n124 = length(intersect(intersect(g1_all_aml_down, g2_all_aml_down), g4_all_aml_down)),
  n134 = length(intersect(intersect(g1_all_aml_down, g3_all_aml_down), g4_all_aml_down)),
  n234 = length(intersect(intersect(g2_all_aml_down, g3_all_aml_down), g4_all_aml_down)),
  n1234 = length(intersect(intersect(intersect(g1_all_aml_down, g2_all_aml_down), g3_all_aml_down), g4_all_aml_down)),
  category = c("PolyA unbiased", "RiboD unbiased", "PolyA biased", "RiboD biased"),
  fill = c("#BB5566", "#D55E00", "#0072B2", "#E69F00"),
  print.mode = c("raw", "percent")
)
```

## Session Info

```{r session info}
sessioninfo::session_info()
```
