---
title: "NB_WT_DESeq2"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## NB v WT

This notebook will be analyzing the expression profiles of Wilms tumor (WT) and neuroblastoma (NB) tumors from the v25 riboD and polyA compendia.

The goal is to determine the extent that genes get called as differentially expressed as a result of the transcript enrichment method of the library preparation process.

I’ll be using a subsetted list of tumors for each disease in each library prep method, based on the correlation analysis and custom tumor map analysis I did previously.

The comparisons we’ll be making (WT = Wilms tumor, NB = neuroblastoma):

- NB polyA vs WT polyA

- NB riboD vs WT riboD

- NB polyA vs WT riboD

- NB riboD vs WT polyA

```{r install packages}
library(tidyverse)
library(DESeq2)
```

```{r import data}
# sample metadata
# This includes the top 15 samples from WT polyA, WT riboD, NB polyA, NB riboD
# The top 15 samples came from the correlation analysis, where this tsv file was generated
filtered_WT_NB_list <- read_tsv("../../input_data/filtered_WT_NB_list.tsv", show_col_types = FALSE)

expression_WT_NB <- read_tsv("../../input_data/expression_WT_NB.tsv.gz", show_col_types = FALSE)

gene_names <- read_tsv("../../input_data/EnsGeneID_Hugo_Observed_Conversions.txt", show_col_types = FALSE)
```

That expression file contains all expression information for all WT and NB samples across the compendia (501 samples). We want only the expression information for the samples we’re interested in.

```{r}
# selecting only the 15 most correlated samples for each category,
# based on the metadata file
interested_expression <- expression_WT_NB %>%
  select(all_of(c("Gene", filtered_WT_NB_list$term)))
```

Filter expression file to keep only genes w/ a count greater than 10 in at least 10 samples.

```{r}
# Explanation:
# across(where(is.numeric)) is selecting all the numeric columns
# i.e., all the sample columns. This deselects the Gene column, since that is a character column

# rowSums() counts how many samples meet the threshold for each gene,
# i.e., have a value of at least 10

# then keeps only the genes that meet that threshold in at least 10 samples
updated_expression <- interested_expression %>%
  filter(rowSums(across(where(is.numeric)) >= 10) >= 10)
```

### NB polyA vs WT polyA

```{r}
# metadata
WT_polyA_NB_polyA_meta <- filtered_WT_NB_list %>%
  filter(disease_and_prep == "WT_polyA" | disease_and_prep == "NB_polyA")
# the pipe | means we're taking all samples called WT_riboD OR NB_polyA in column disease_and_prep

# expression
WT_polyA_NB_polyA_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", WT_polyA_NB_polyA_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_WT_polyA_NB_polyA_updated_expr <- WT_polyA_NB_polyA_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged
fix_WT_polyA_NB_polyA_meta <- WT_polyA_NB_polyA_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names of metadata are same as column names of expression
all(rownames(fix_WT_polyA_NB_polyA_meta) %in% colnames(fix_WT_polyA_NB_polyA_updated_expr))
# check if order is the same
all(rownames(fix_WT_polyA_NB_polyA_meta) == colnames(fix_WT_polyA_NB_polyA_updated_expr))
```

```{r}
# format data
round_WT_polyA_NB_polyA_updated_expr = round(fix_WT_polyA_NB_polyA_updated_expr)

WT_polyA_NB_polyA_updated_expr_matrix <- as.matrix(round_WT_polyA_NB_polyA_updated_expr)

fix_WT_polyA_NB_polyA_meta$disease_and_prep <- as.factor(fix_WT_polyA_NB_polyA_meta$disease_and_prep)
```

```{r}
# set reference levels
# we want to make sure the references are consistent across the comparisons
levels(fix_WT_polyA_NB_polyA_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_WT_polyA_NB_polyA_updated <- DESeqDataSetFromMatrix(countData = WT_polyA_NB_polyA_updated_expr_matrix,
                              colData = fix_WT_polyA_NB_polyA_meta,
                              design = ~disease_and_prep)
dds_WT_polyA_NB_polyA_updated
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
WT_polyA_NB_polyA_updated_genes_to_keep <- rowSums(counts(dds_WT_polyA_NB_polyA_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_WT_pA_NB_pA_updated <- dds_WT_polyA_NB_polyA_updated[which(WT_polyA_NB_polyA_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_WT_pA_NB_pA_updated_object <- DESeq(dds_WT_pA_NB_pA_updated)
```

```{r}
# create results table
result_dds_WT_pA_NB_pA_updated <- results(dds_WT_pA_NB_pA_updated_object)
# summarize results table
summary(result_dds_WT_pA_NB_pA_updated, alpha = 0.05)
```

```{r}
# convert results to dataframe
result_dds_WT_pA_NB_pA_updated_df <- result_dds_WT_pA_NB_pA_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
WT_polyA_NB_polyA_LFC1great_padj <- result_dds_WT_pA_NB_pA_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
WT_polyA_NB_polyA_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
WT_polyA_NB_polyA_LFC1less_padj <- result_dds_WT_pA_NB_pA_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
WT_polyA_NB_polyA_LFC1less_padj
```

```{r}
# convert EnsemblID to HugoID for gene name readability
hugo_results_WT_polyA_NB_polyA_updated <- result_dds_WT_pA_NB_pA_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_WT_polyA_NB_polyA_updated, path = "../../output_data/NB_WT/hugo_results_WT_polyA_NB_polyA_updated.tsv.gz")
```

### NB riboD vs WT riboD

```{r}
# metadata
WT_riboD_NB_riboD_meta <- filtered_WT_NB_list %>%
  filter(disease_and_prep == "WT_riboD" | disease_and_prep == "NB_riboD")

# expression
WT_riboD_NB_riboD_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", WT_riboD_NB_riboD_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_WT_riboD_NB_riboD_updated_expr <- WT_riboD_NB_riboD_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id")
fix_WT_riboD_NB_riboD_meta <- WT_riboD_NB_riboD_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_WT_riboD_NB_riboD_meta) %in% colnames(fix_WT_riboD_NB_riboD_updated_expr))

# check in order is the same
all(rownames(fix_WT_riboD_NB_riboD_meta) == colnames(fix_WT_riboD_NB_riboD_updated_expr))
```

```{r}
# format data
round_WT_riboD_NB_riboD_updated_expr = round(fix_WT_riboD_NB_riboD_updated_expr)

WT_riboD_NB_riboD_updated_expr_matrix <- as.matrix(round_WT_riboD_NB_riboD_updated_expr)

fix_WT_riboD_NB_riboD_meta$disease_and_prep <- as.factor(fix_WT_riboD_NB_riboD_meta$disease_and_prep)
```

```{r}
# set reference levels
# we want to make sure the references are consistent across the comparisons
levels(fix_WT_riboD_NB_riboD_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_WT_riboD_NB_riboD_updated <- DESeqDataSetFromMatrix(countData = WT_riboD_NB_riboD_updated_expr_matrix,
                              colData = fix_WT_riboD_NB_riboD_meta,
                              design = ~disease_and_prep)
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
WT_riboD_NB_riboD_updated_genes_to_keep <- rowSums(counts(dds_WT_riboD_NB_riboD_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_WT_riboD_NB_riboD_updated <- dds_WT_riboD_NB_riboD_updated[which(WT_riboD_NB_riboD_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_WT_riboD_NB_riboD_updated_object <- DESeq(dds_WT_riboD_NB_riboD_updated)
```

```{r}
# create results table
res_WT_riboD_NB_riboD_updated <- results(dds_WT_riboD_NB_riboD_updated_object)

# summarize results table
summary(res_WT_riboD_NB_riboD_updated, alpha = 0.05)
```

```{r}
# first we convert to df
res_WT_riboD_NB_riboD_updated_df <- res_WT_riboD_NB_riboD_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
WT_riboD_NB_riboD_LFC1great_padj <- res_WT_riboD_NB_riboD_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
WT_riboD_NB_riboD_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
WT_riboD_NB_riboD_LFC1less_padj <- res_WT_riboD_NB_riboD_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
WT_riboD_NB_riboD_LFC1less_padj
```

```{r}
hugo_results_WT_riboD_NB_riboD_updated <- res_WT_riboD_NB_riboD_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_WT_riboD_NB_riboD_updated, path = "../../output_data/NB_WT/hugo_results_WT_riboD_NB_riboD_updated.tsv.gz")
```

### **WT polyA v. NB riboD**

```{r}
# metadata
WT_polyA_NB_riboD_meta <- filtered_WT_NB_list %>%
  filter(disease_and_prep == "WT_polyA" | disease_and_prep == "NB_riboD")
# the pipe | means we're taking all samples called WT_polyA OR NB_riboD in column disease_and_prep

# expression
WT_polyA_NB_riboD_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", WT_polyA_NB_riboD_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_WT_polyA_NB_riboD_updated_expr <- WT_polyA_NB_riboD_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_WT_polyA_NB_riboD_meta <- WT_polyA_NB_riboD_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_WT_polyA_NB_riboD_meta) %in% colnames(fix_WT_polyA_NB_riboD_updated_expr))

# check in order is the same
all(rownames(fix_WT_polyA_NB_riboD_meta) == colnames(fix_WT_polyA_NB_riboD_updated_expr))
```

```{r}
# format data
round_WT_polyA_NB_riboD_updated_expr = round(fix_WT_polyA_NB_riboD_updated_expr)

WT_polyA_NB_riboD_updated_expr_matrix <- as.matrix(round_WT_polyA_NB_riboD_updated_expr)

fix_WT_polyA_NB_riboD_meta$disease_and_prep <- as.factor(fix_WT_polyA_NB_riboD_meta$disease_and_prep)

# check reference levels
levels(fix_WT_polyA_NB_riboD_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_WT_polyA_NB_riboD_updated <- DESeqDataSetFromMatrix(countData = WT_polyA_NB_riboD_updated_expr_matrix,
                              colData = fix_WT_polyA_NB_riboD_meta,
                              design = ~disease_and_prep)

dds_WT_polyA_NB_riboD_updated
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
WT_polyA_NB_riboD_updated_genes_to_keep <- rowSums(counts(dds_WT_polyA_NB_riboD_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_WT_polyA_NB_riboD_updated <- dds_WT_polyA_NB_riboD_updated[which(WT_polyA_NB_riboD_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_WT_polyA_NB_riboD_updated_object <- DESeq(dds_WT_polyA_NB_riboD_updated)
```

```{r}
# create results table
result_dds_WT_polyA_NB_riboD_updated <- results(dds_WT_polyA_NB_riboD_updated_object)
# summarize results table
summary(result_dds_WT_polyA_NB_riboD_updated, alpha = 0.05)
```

```{r}
# convert results to dataframe
result_dds_WT_polyA_NB_riboD_updated_df <- result_dds_WT_polyA_NB_riboD_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
WT_polyA_NB_riboD_LFC1great_padj <- result_dds_WT_polyA_NB_riboD_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
WT_polyA_NB_riboD_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
WT_polyA_NB_riboD_LFC1less_padj <- result_dds_WT_polyA_NB_riboD_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
WT_polyA_NB_riboD_LFC1less_padj
```

```{r}
hugo_results_WT_polyA_NB_riboD_updated <- result_dds_WT_polyA_NB_riboD_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_WT_polyA_NB_riboD_updated, path = "../../output_data/NB_WT/hugo_results_WT_polyA_NB_riboD_updated.tsv.gz")
```

### **WT riboD v. NB polyA**

```{r}
# metadata
WT_riboD_NB_polyA_meta <- filtered_WT_NB_list %>%
  filter(disease_and_prep == "WT_riboD" | disease_and_prep == "NB_polyA")

# expression
WT_riboD_NB_polyA_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", WT_riboD_NB_polyA_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_WT_riboD_NB_polyA_updated_expr <- WT_riboD_NB_polyA_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_WT_riboD_NB_polyA_meta <- WT_riboD_NB_polyA_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_WT_riboD_NB_polyA_meta) %in% colnames(fix_WT_riboD_NB_polyA_updated_expr))

# check if order is the same
all(rownames(fix_WT_riboD_NB_polyA_meta) == colnames(fix_WT_riboD_NB_polyA_updated_expr))
```

```{r}
# format data
round_WT_riboD_NB_polyA_updated_expr = round(fix_WT_riboD_NB_polyA_updated_expr)

WT_riboD_NB_polyA_updated_expr_matrix <- as.matrix(round_WT_riboD_NB_polyA_updated_expr)

fix_WT_riboD_NB_polyA_meta$disease_and_prep <- as.factor(fix_WT_riboD_NB_polyA_meta$disease_and_prep)
```

```{r}
# check reference levels
levels(fix_WT_riboD_NB_polyA_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_WT_riboD_NB_polyA_updated <- DESeqDataSetFromMatrix(countData = WT_riboD_NB_polyA_updated_expr_matrix,
                              colData = fix_WT_riboD_NB_polyA_meta,
                              design = ~disease_and_prep)
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
WT_riboD_NB_polyA_updated_genes_to_keep <- rowSums(counts(dds_WT_riboD_NB_polyA_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_WT_riboD_NB_polyA_updated <- dds_WT_riboD_NB_polyA_updated[which(WT_riboD_NB_polyA_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_WT_riboD_NB_polyA_updated_object <- DESeq(dds_WT_riboD_NB_polyA_updated)
```

```{r}
# create results table
res_WT_riboD_NB_polyA_updated <- results(dds_WT_riboD_NB_polyA_updated_object)

# summarize results table
summary(res_WT_riboD_NB_polyA_updated, alpha = 0.05)
```

```{r}
# first we convert to df
results_WT_riboD_NB_polyA_updated_df <- res_WT_riboD_NB_polyA_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
WT_riboD_NB_polyA_LFC1great_padj <- results_WT_riboD_NB_polyA_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
WT_riboD_NB_polyA_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
WT_riboD_NB_polyA_LFC1less_padj <- results_WT_riboD_NB_polyA_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
WT_riboD_NB_polyA_LFC1less_padj
```

```{r}
hugo_results_WT_riboD_NB_polyA_updated <- results_WT_riboD_NB_polyA_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_WT_riboD_NB_polyA_updated, path = "../../output_data/NB_WT/hugo_results_WT_riboD_NB_polyA_updated.tsv.gz")
```

### **Session Info**

```{r}
sessioninfo::session_info()
```
