---
title: "ALL_AML_DESeq2_analysis"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## **ALL v AML - DESeq2 Results Analysis**

```{r load libraries}
library(tidyverse)
library(VennDiagram)
library(ggplot2)
```

```{r}
AMLPolyA_ALLPolyA <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_polyA_ALL_polyA_updated.tsv.gz")
AMLRiboD_ALLRiboD <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_riboD_ALL_riboD_updated.tsv.gz")
AMLPolyA_ALLRiboD <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_polyA_ALL_riboD_updated.tsv.gz")
AMLRiboD_ALLPolyA <- read_tsv("../../output_data/ALL_AML/hugo_results_AML_riboD_ALL_polyA_updated.tsv.gz")
```

## **L2FC \>/\<= 1 or -1**

Filter for statistically significant DEGs

```{r}
# Filter for genes with a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05

AMLPolyA_ALLPolyA_great1 <- AMLPolyA_ALLPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "AMLPolyA_ALLPolyA_great1") %>%
  relocate(comparison)

AMLRiboD_ALLRiboD_great1 <- AMLRiboD_ALLRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "AMLRiboD_ALLRiboD_great1") %>%
  relocate(comparison)

AMLPolyA_ALLRiboD_great1 <- AMLPolyA_ALLRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "AMLPolyA_ALLRiboD_great1") %>%
  relocate(comparison)

AMLRiboD_ALLPolyA_great1 <- AMLRiboD_ALLPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "AMLRiboD_ALLPolyA_great1") %>%
  relocate(comparison)


# Filter for genes with a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05

AMLPolyA_ALLPolyA_less1 <- AMLPolyA_ALLPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "AMLPolyA_ALLPolyA_less1") %>%
  relocate(comparison)

AMLRiboD_ALLRiboD_less1 <- AMLRiboD_ALLRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "AMLRiboD_ALLRiboD_less1") %>%
  relocate(comparison)

AMLPolyA_ALLRiboD_less1 <- AMLPolyA_ALLRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "AMLPolyA_ALLRiboD_less1") %>%
  relocate(comparison)

AMLRiboD_ALLPolyA_less1 <- AMLRiboD_ALLPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "AMLRiboD_ALLPolyA_less1") %>%
  relocate(comparison)
```

### **PolyA unbiased against biased**

#### **Up DEGs**

```{r ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased}
ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased <- list(
  "Up in AML polyA relative to ALL polyA (polyA unbiased)" = AMLPolyA_ALLPolyA_great1$Gene,
  "Up in AML polyA relative to ALL riboD (polyA biased)" = AMLPolyA_ALLRiboD_great1$Gene,
  "Up in AML riboD relative to ALL polyA (riboD biased)" = AMLRiboD_ALLPolyA_great1$Gene
)

write_rds(ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased, "../../output_data/ALL_AML/ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased.rds")
```

```{r ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased_VD}
ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Up in AML polyA relative to \n ALL polyA (polyA unbiased)",
    "Up in AML polyA relative to \n ALL riboD (polyA biased)",
    "Up in AML riboD relative to \n ALL polyA (riboD biased)"
    ),
  filename = NULL,
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in AML relative to ALL (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
ALL_AML_up_polyAunbiased_polyAbiased_riboDbiased_VD
```

#### **Down DEGs**

```{r ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased}
ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased <- list(
  "Down in AML polyA relative to ALL polyA (polyA unbiased)" = AMLPolyA_ALLPolyA_less1$Gene,
  "Down in AML polyA relative to ALL riboD (polyA biased)" = AMLPolyA_ALLRiboD_less1$Gene,
  "Down in AML riboD relative to ALL polyA (riboD biased)" = AMLRiboD_ALLPolyA_less1$Gene
)
write_rds(ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased, "../../output_data/ALL_AML/ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased.rds")
```

```{r ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased_VD}
ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Down in AML polyA relative to \n ALL polyA (polyA unbiased)",
    "Down in AML polyA relative to \n ALL riboD (polyA biased)",
    "Down in AML riboD relative to \n ALL polyA (riboD biased)"
    ),
  filename = NULL,
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in AML relative to ALL (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
ALL_AML_down_polyAunbiased_polyAbiased_riboDbiased_VD
```

### **RiboD unbiased against biased**

#### **Up DEGs**

```{r ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased}
ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased <- list(
  "Up in AML riboD relative to ALL riboD (riboD unbiased)" = AMLRiboD_ALLRiboD_great1$Gene,
  "Up in AML polyA relative to ALL riboD (polyA biased)" = AMLPolyA_ALLRiboD_great1$Gene,
  "Up in AML riboD relative to ALL polyA (riboD biased)" = AMLRiboD_ALLPolyA_great1$Gene
)
write_rds(ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased, "../../output_data/ALL_AML/ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased.rds")
```

```{r ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased_VD}
ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Up in AML riboD relative to \n ALL riboD (riboD unbiased)",
    "Up in AML polyA relative to \n ALL riboD (polyA biased)",
    "Up in AML riboD relative to \n ALL polyA (riboD biased)"
    ),
  filename = NULL,
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in AML relative to ALL (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
ALL_AML_up_riboDunbiased_polyAbiased_riboDbiased_VD
```

#### **Down DEGs**

```{r ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased}
ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased <- list(
  "Down in AML riboD relative to ALL riboD (riboD unbiased)" = AMLRiboD_ALLRiboD_less1$Gene,
  "Down in AML polyA relative to ALL riboD (polyA biased)" = AMLPolyA_ALLRiboD_less1$Gene,
  "Down in AML riboD relative to ALL polyA (riboD biased)" = AMLRiboD_ALLPolyA_less1$Gene
)

write_rds(ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased, "../../output_data/ALL_AML/ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased.rds")
```

```{r ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased_VD}
ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Down in AML riboD relative to \n ALL riboD (riboD unbiased)",
    "Down in AML polyA relative to \n ALL riboD (polyA biased)",
    "Down in AML riboD relative to \n ALL polyA (riboD biased)"
    ),
  filename = NULL,
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in AML relative to ALL (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
ALL_AML_down_riboDunbiased_polyAbiased_riboDbiased_VD
```

### **Histograms of LFC of genes in intersections**

Starting with upregulated genes in the polyAunbiased:polyAbiased:riboDbiased comparison.

Looking at genes that are unique to polyAbiased

```{r ALL_AML_up_polyAunbiased_uniquetopolyAbiased}
ALL_AML_up_polyAunbiased_uniquetopolyAbiased <- AMLPolyA_ALLRiboD_great1 %>%
  anti_join(AMLPolyA_ALLPolyA_great1, by = "Gene") %>%
  anti_join(AMLRiboD_ALLPolyA_great1, by = "Gene")
print(nrow(ALL_AML_up_polyAunbiased_uniquetopolyAbiased))
```

```{r ALL_AML_up_polyAunbiased_uniquetopolyAbiased_avgLFC}
ALL_AML_up_polyAunbiased_uniquetopolyAbiased_avgLFC <- ALL_AML_up_polyAunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_up_polyAunbiased_uniquetopolyAbiased_avgLFC
```

```{r ALL_AML_up_polyAunbiased_uniquetopolyAbiased_hist}
ALL_AML_up_polyAunbiased_uniquetopolyAbiased_hist <- ggplot(ALL_AML_up_polyAunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in AML polyA relative to ALL riboD",
       subtitle = "polyA biased, compared to polyA unbiased. Median LFC = 1.45",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))

ALL_AML_up_polyAunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboDbiased

```{r ALL_AML_up_polyAunbiased_uniquetoriboDbiased}
ALL_AML_up_polyAunbiased_uniquetoriboDbiased <- AMLRiboD_ALLPolyA_great1 %>%
  anti_join(AMLPolyA_ALLPolyA_great1, by = "Gene") %>%
  anti_join(AMLPolyA_ALLRiboD_great1, by = "Gene")
print(nrow(ALL_AML_up_polyAunbiased_uniquetoriboDbiased))
```

```{r ALL_AML_up_polyAunbiased_uniquetoriboDbiased_avgLFC}
ALL_AML_up_polyAunbiased_uniquetoriboDbiased_avgLFC <- ALL_AML_up_polyAunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_up_polyAunbiased_uniquetoriboDbiased_avgLFC
```

```{r ALL_AML_up_polyAunbiased_uniquetoriboDbiased_hist}
ALL_AML_up_polyAunbiased_uniquetoriboDbiased_hist <- ggplot(ALL_AML_up_polyAunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,1000)) +
#    scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in AML riboD relative to ALL polyA",
       subtitle = "riboD biased, compared to polyA unbiased. Median LFC = 1.77",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_up_polyAunbiased_uniquetoriboDbiased_hist
```

Upregulated genes in the riboD unbiased:polyAbiased:riboDbiased comparison

Looking at genes that are unique to the polyA biased

```{r ALL_AML_up_riboDunbiased_uniquetopolyAbiased}
ALL_AML_up_riboDunbiased_uniquetopolyAbiased <- AMLPolyA_ALLRiboD_great1 %>%
  anti_join(AMLRiboD_ALLRiboD_great1, by = "Gene") %>%
  anti_join(AMLRiboD_ALLPolyA_great1, by = "Gene")
print(nrow(ALL_AML_up_riboDunbiased_uniquetopolyAbiased))
```

```{r ALL_AML_up_riboDunbiased_uniquetopolyAbiased_avgLFC}
ALL_AML_up_riboDunbiased_uniquetopolyAbiased_avgLFC <- ALL_AML_up_riboDunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_up_riboDunbiased_uniquetopolyAbiased_avgLFC
```

```{r ALL_AML_up_riboDunbiased_uniquetopolyAbiased_hist}
ALL_AML_up_riboDunbiased_uniquetopolyAbiased_hist <- ggplot(ALL_AML_up_riboDunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in AML polyA relative to ALL riboD",
       subtitle = "polyA biased, compared to riboD unbiased. Median LFC = 1.41",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_up_riboDunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboDbiased

```{r ALL_AML_up_riboDunbiased_uniquetoriboDbiased}
ALL_AML_up_riboDunbiased_uniquetoriboDbiased <- AMLRiboD_ALLPolyA_great1 %>%
  anti_join(AMLRiboD_ALLRiboD_great1, by = "Gene") %>%
  anti_join(AMLPolyA_ALLRiboD_great1, by = "Gene")
print(nrow(ALL_AML_up_riboDunbiased_uniquetoriboDbiased))
```

```{r ALL_AML_up_riboDunbiased_uniquetoriboDbiased_avgLFC}
ALL_AML_up_riboDunbiased_uniquetoriboDbiased_avgLFC <- ALL_AML_up_riboDunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_up_riboDunbiased_uniquetoriboDbiased_avgLFC
```

```{r ALL_AML_up_riboDunbiased_uniquetoriboDbiased_hist}
ALL_AML_up_riboDunbiased_uniquetoriboDbiased_hist <- ggplot(ALL_AML_up_riboDunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800, 1000, 1200), limits = c(0,1200)) +
#    scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in AML riboD relative to ALL polyA",
       subtitle = "riboD biased, compared to riboD unbiased. Median LFC = 2.17",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_up_riboDunbiased_uniquetoriboDbiased_hist
```

Looking at downregulated genes in the polyA unbiased:polyA biased:riboD biased comparison

Looking at genes that are unique to the polyA biased

```{r ALL_AML_down_polyAunbiased_uniquetopolyAbiased}
ALL_AML_down_polyAunbiased_uniquetopolyAbiased <- AMLPolyA_ALLRiboD_less1 %>%
  anti_join(AMLPolyA_ALLPolyA_less1, by = "Gene") %>%
  anti_join(AMLRiboD_ALLPolyA_less1, by = "Gene")
print(nrow(ALL_AML_down_polyAunbiased_uniquetopolyAbiased))
```

```{r ALL_AML_down_polyAunbiased_uniquetopolyAbiased_avgLFC}
ALL_AML_down_polyAunbiased_uniquetopolyAbiased_avgLFC <- ALL_AML_down_polyAunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_down_polyAunbiased_uniquetopolyAbiased_avgLFC
```

```{r ALL_AML_down_polyAunbiased_uniquetopolyAbiased_hist}
ALL_AML_down_polyAunbiased_uniquetopolyAbiased_hist <- ggplot(ALL_AML_down_polyAunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -1, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in AML polyA relative to ALL riboD",
       subtitle = "polyA biased, compared to polyA unbiased. Median LFC = -1.7",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

ALL_AML_down_polyAunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to the riboD biased

```{r ALL_AML_down_polyAunbiased_uniquetoriboDbiased}
ALL_AML_down_polyAunbiased_uniquetoriboDbiased <- AMLRiboD_ALLPolyA_less1 %>%
  anti_join(AMLPolyA_ALLPolyA_less1, by = "Gene") %>%
  anti_join(AMLPolyA_ALLRiboD_less1, by = "Gene")
print(nrow(ALL_AML_down_polyAunbiased_uniquetoriboDbiased))
```

```{r ALL_AML_down_polyAunbiased_uniquetoriboDbiased_avgLFC}
ALL_AML_down_polyAunbiased_uniquetoriboDbiased_avgLFC <- ALL_AML_down_polyAunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_down_polyAunbiased_uniquetoriboDbiased_avgLFC
```

```{r ALL_AML_down_polyAunbiased_uniquetoriboDbiased_hist}
ALL_AML_down_polyAunbiased_uniquetoriboDbiased_hist <- ggplot(ALL_AML_down_polyAunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, compared to polyA_unbiased. Median LFC = -1.31",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

ALL_AML_down_polyAunbiased_uniquetoriboDbiased_hist
```

Looking at downregulated genes in the riboD unbiased:polyA biased:riboD biased comparison

Looking at genes that are unique to the polyA biased

```{r ALL_AML_down_riboDunbiased_uniquetopolyAbiased}
ALL_AML_down_riboDunbiased_uniquetopolyAbiased <- AMLPolyA_ALLRiboD_less1 %>%
  anti_join(AMLRiboD_ALLRiboD_less1, by = "Gene") %>%
  anti_join(AMLRiboD_ALLPolyA_less1, by = "Gene")
print(nrow(ALL_AML_down_riboDunbiased_uniquetopolyAbiased))
```

```{r ALL_AML_down_riboDunbiased_uniquetopolyAbiased_avgLFC}
ALL_AML_down_riboDunbiased_uniquetopolyAbiased_avgLFC <- ALL_AML_down_riboDunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_down_riboDunbiased_uniquetopolyAbiased_avgLFC
```

```{r ALL_AML_down_riboDunbiased_uniquetopolyAbiased_hist}
ALL_AML_down_riboDunbiased_uniquetopolyAbiased_hist <- ggplot(ALL_AML_down_riboDunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -1, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in AML polyA relative to ALL riboD",
       subtitle = "polyA biased, compared to riboD unbiased. Median LFC = -1.69",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_down_riboDunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to the riboD biased

```{r ALL_AML_down_riboDunbiased_uniquetoriboDbiased}
ALL_AML_down_riboDunbiased_uniquetoriboDbiased <- AMLRiboD_ALLPolyA_less1 %>%
  anti_join(AMLRiboD_ALLRiboD_less1, by = "Gene") %>%
  anti_join(AMLPolyA_ALLRiboD_less1, by = "Gene")
print(nrow(ALL_AML_down_riboDunbiased_uniquetoriboDbiased))
```

```{r ALL_AML_down_riboDunbiased_uniquetoriboDbiased_avgLFC}
ALL_AML_down_riboDunbiased_uniquetoriboDbiased_avgLFC <- ALL_AML_down_riboDunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
ALL_AML_down_riboDunbiased_uniquetoriboDbiased_avgLFC
```

```{r ALL_AML_down_riboDunbiased_uniquetoriboDbiased_hist}
ALL_AML_down_riboDunbiased_uniquetoriboDbiased_hist <- ggplot(ALL_AML_down_riboDunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in AML riboD relative to ALL polyA",
       subtitle = "riboD biased, compared to riboD unbiased. Median LFC = -1.42",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_down_riboDunbiased_uniquetoriboDbiased_hist
```

Genes in common between the polyAunbiased:polyAbiased:riboDbiased venn diagram and the riboDunbiased:polyAbiased:riboDbiased venn diagram

upregulated polyA biased

```{r ALL_AML_up_polyAbiased}
ALL_AML_up_polyAbiased <- inner_join(ALL_AML_up_polyAunbiased_uniquetopolyAbiased, ALL_AML_up_riboDunbiased_uniquetopolyAbiased, by = "Gene")

write_rds(ALL_AML_up_polyAbiased, "../../output_data/ALL_AML/ALL_AML_up_polyAbiased.rds")

print(nrow(ALL_AML_up_polyAbiased))
```

```{r ALL_AML_up_polyAbiased_avgLFC}
ALL_AML_up_polyAbiased_avgLFC <- ALL_AML_up_polyAbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
ALL_AML_up_polyAbiased_avgLFC
```

```{r ALL_AML_up_polyAbiased_hist}
ALL_AML_up_polyAbiased_hist <- ggplot(ALL_AML_up_polyAbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in AML polyA relative to ALL riboD",
       subtitle = "polyA biased, Median LFC = 1.34",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_up_polyAbiased_hist
```

upregulated riboD biased

```{r ALL_AML_up_riboDbiased}
ALL_AML_up_riboDbiased <- inner_join(ALL_AML_up_polyAunbiased_uniquetoriboDbiased, ALL_AML_up_riboDunbiased_uniquetoriboDbiased, by = "Gene")

write_rds(ALL_AML_up_riboDbiased, "../../output_data/ALL_AML/ALL_AML_up_riboDbiased.rds")

print(nrow(ALL_AML_up_riboDbiased))
```

```{r ALL_AML_up_riboDbiased_avgLFC}
ALL_AML_up_riboDbiased_avgLFC <- ALL_AML_up_riboDbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
ALL_AML_up_riboDbiased_avgLFC
```

```{r ALL_AML_up_riboDbiased_hist}
ALL_AML_up_riboDbiased_hist <- ggplot(ALL_AML_up_riboDbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in AML riboD relative to ALL polyA",
       subtitle = "riboDbiased, Median LFC = 1.75",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_up_riboDbiased_hist
```

Genes in common between the polyAunbiased:polyAbiased:riboDbiased venn diagram and the riboDunbiased:polyAbiased:riboDbiased venn diagram

downregulated polyA biased

```{r ALL_AML_down_polyAbiased}
ALL_AML_down_polyAbiased <- inner_join(ALL_AML_down_polyAunbiased_uniquetopolyAbiased, ALL_AML_down_riboDunbiased_uniquetopolyAbiased, by = "Gene")

write_rds(ALL_AML_down_polyAbiased, "../../output_data/ALL_AML/ALL_AML_down_polyAbiased.rds")

print(nrow(ALL_AML_down_polyAbiased))
```

```{r ALL_AML_down_polyAbiased_avgLFC}
ALL_AML_down_polyAbiased_avgLFC <- ALL_AML_down_polyAbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
ALL_AML_down_polyAbiased_avgLFC
```

```{r ALL_AML_down_polyAbiased_hist}
ALL_AML_down_polyAbiased_hist <- ggplot(ALL_AML_down_polyAbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in AML polyA relative to ALL riboD",
       subtitle = "polyA biased, Median LFC = -1.63",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_down_polyAbiased_hist
```

downregulated riboD biased

```{r ALL_AML_down_riboDbiased}
ALL_AML_down_riboDbiased <- inner_join(ALL_AML_down_polyAunbiased_uniquetoriboDbiased, ALL_AML_down_riboDunbiased_uniquetoriboDbiased, by = "Gene")

write_rds(ALL_AML_down_riboDbiased, "../../output_data/ALL_AML/ALL_AML_down_riboDbiased.rds")

print(nrow(ALL_AML_down_riboDbiased))
```

```{r ALL_AML_down_riboDbiased_avgLFC}
ALL_AML_down_riboDbiased_avgLFC <- ALL_AML_down_riboDbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
ALL_AML_down_riboDbiased_avgLFC
```

```{r ALL_AML_down_riboDbiased_hist}
ALL_AML_down_riboDbiased_hist <- ggplot(ALL_AML_down_riboDbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in AML riboD relative to ALL polyA",
       subtitle = "riboD biased, Mean LFC = -1.30",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
ALL_AML_down_riboDbiased_hist
```

### **Session Info**

```{r session_info}
sessioninfo::session_info()
```
