---
title: "misc-analysis"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## 

Investigating the median ratio genes from each disease cohort in all diseases

```{r load library}
library(tidyverse)
library(cowplot)
library(ggbeeswarm)
```

```{r load files}
rsem_log2TPM1_THR13 <- read_tsv("../input_data/matched_THR13_0970/rsem_ensembl_log2TPM1_THR13_0970_S12-S17.tsv.gz", show_col_types = FALSE)

# sample ID files
# polyA
SS_polyA_list <- read_tsv("../input_data/sample_selection/SS_polyA_list.tsv", show_col_types = FALSE)
aRMS_polyA_list <- read_tsv("../input_data/sample_selection/aRMS_polyA_list.tsv", show_col_types = FALSE)
WT_polyA_list <- read_tsv("../input_data/sample_selection/WT_polyA_list.tsv", show_col_types = FALSE)
NB_polyA_list <- read_tsv("../input_data/sample_selection/NB_polyA_list.tsv", show_col_types = FALSE)
ALL_polyA_list <- read_tsv("../input_data/sample_selection/ALL_polyA_list.tsv", show_col_types = FALSE)
AML_polyA_list <- read_tsv("../input_data/sample_selection/AML_polyA_list.tsv", show_col_types = FALSE)

# riboD
SS_riboD_list <- read_tsv("../input_data/sample_selection/SS_riboD_list.tsv", show_col_types = FALSE)
aRMS_riboD_list <- read_tsv("../input_data/sample_selection/aRMS_riboD_list.tsv", show_col_types = FALSE)
WT_riboD_list <- read_tsv("../input_data/sample_selection/WT_riboD_list.tsv", show_col_types = FALSE)
NB_riboD_list <- read_tsv("../input_data/sample_selection/NB_riboD_list.tsv", show_col_types = FALSE)
ALL_riboD_list <- read_tsv("../input_data/sample_selection/ALL_riboD_list.tsv", show_col_types = FALSE)
AML_riboD_list <- read_tsv("../input_data/sample_selection/AML_riboD_list.tsv", show_col_types = FALSE)

# log2(TPM+1) expression files
SS_polyA_log2tpm1 <- read_tsv("../input_data/sample_selection/SS_polyA_log2tpm1.tsv", show_col_types = FALSE)
SS_riboD_log2tpm1 <- read_tsv("../input_data/sample_selection/SS_riboD_log2tpm1.tsv", show_col_types = FALSE)

aRMS_polyA_log2tpm1 <- read_tsv("../input_data/sample_selection/aRMS_polyA_log2tpm1.tsv", show_col_types = FALSE)
aRMS_riboD_log2tpm1 <- read_tsv("../input_data/sample_selection/aRMS_riboD_log2tpm1.tsv", show_col_types = FALSE)

WT_polyA_log2tpm1 <- read_tsv("../input_data/sample_selection/WT_polyA_log2tpm1.tsv", show_col_types = FALSE)
WT_riboD_log2tpm1 <- read_tsv("../input_data/sample_selection/WT_riboD_log2tpm1.tsv", show_col_types = FALSE)

NB_polyA_log2tpm1 <- read_tsv("../input_data/sample_selection/NB_polyA_log2tpm1.tsv", show_col_types = FALSE)
NB_riboD_log2tpm1 <- read_tsv("../input_data/sample_selection/NB_riboD_log2tpm1.tsv", show_col_types = FALSE)

ALL_polyA_log2tpm1 <- read_tsv("../input_data/sample_selection/ALL_polyA_log2tpm1.tsv", show_col_types = FALSE)
ALL_riboD_log2tpm1 <- read_tsv("../input_data/sample_selection/ALL_riboD_log2tpm1.tsv", show_col_types = FALSE)

AML_polyA_log2tpm1 <- read_tsv("../input_data/sample_selection/AML_polyA_log2tpm1.tsv", show_col_types = FALSE)
AML_riboD_log2tpm1 <- read_tsv("../input_data/sample_selection/AML_riboD_log2tpm1.tsv", show_col_types = FALSE)


# to convert EnsemblIDs to HugoIDs
gene_names <- read.table("../input_data/EnsGeneID_Hugo_Observed_Conversions.txt",
header = TRUE, sep = "\t", stringsAsFactors = FALSE )
```

```{r rsem_log2TPM1_THR13_prepname}
rsem_log2TPM1_THR13_prepname <- rsem_log2TPM1_THR13 %>%
  mutate(
    prepname = case_when(
    full_sample == "THR13_0970_S12" ~ "PolyA_1",
    full_sample == "THR13_0970_S13" ~ "PolyA_2",
    full_sample == "THR13_0970_S14" ~ "PolyA_3",
    full_sample == "THR13_0970_S15" ~ "RiboD_1",
    full_sample == "THR13_0970_S16" ~ "RiboD_2",
    full_sample == "THR13_0970_S17" ~ "RiboD_3"
  ))

THR13_polyA_log2tpm1 <- rsem_log2TPM1_THR13_prepname %>%
  filter(lib_prep == "polyA") %>%
  rename(Gene = ensembl_gene_ID) %>%
  select(full_sample, Gene, log2TPM1) %>%
  pivot_wider(names_from = full_sample, values_from = log2TPM1)

THR13_riboD_log2tpm1 <- rsem_log2TPM1_THR13_prepname %>%
  filter(lib_prep == "riboD") %>%
  rename(Gene = ensembl_gene_ID) %>%
  select(full_sample, Gene, log2TPM1) %>%
  pivot_wider(names_from = full_sample, values_from = log2TPM1)
```

```{r ratio_genes}
# Create a dataframe that specifies all the median ratio gene (as calculated in Fig1E)
ratio_genes <- data.frame(HugoID = c("MTMR2", "TM2D1", "SYPL1", "CCDC64", "XPO5", "EML3", "NUDT5"))
```

Function for filtering for the ratio genes

```{r filter_ratio}
filter_ratio <- function(df, ratio_genes) {
  df %>% filter(Gene %in% ratio_genes$HugoID) }
```

Function for calculating median expression across samples

```{r compute_gene_medians}
compute_gene_medians_polyA <- function(polyA_df, disease_name, Compendia) {
  polyA_medians <- polyA_df %>%
    rowwise() %>%
    mutate(polyA_medians = median(c_across(-Gene), na.rm = TRUE)) %>%
    ungroup() %>%
    select(Gene, polyA_medians)
}
compute_gene_medians_riboD <- function(riboD_df, disease_name, Compendia) {
   riboD_medians <- riboD_df %>%
    rowwise() %>%
    mutate(riboD_medians = median(c_across(-Gene), na.rm = TRUE)) %>%
    ungroup() %>%
    select(Gene, riboD_medians)
}
```

Function for converting EnsemblID to HugoID for readability

```{r convert_to_hugo}
convert_to_hugo <- function(df) {
  df %>%
    left_join(gene_names, by = c("Gene" = "EnsGeneID")) %>%
    mutate(Gene = ifelse(!is.na(HugoID), HugoID, Gene)) %>%
    select(-HugoID)
}
```

Applying gene medians function

```{r apply gene_medians}
aRMS_polyA_medians <- compute_gene_medians_polyA(aRMS_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "aRMS", lib_prep = "PolyA") %>%
  convert_to_hugo()

aRMS_riboD_medians <- compute_gene_medians_riboD(aRMS_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "aRMS", lib_prep = "RiboD") %>%
  convert_to_hugo()

SS_polyA_medians <- compute_gene_medians_polyA(SS_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "SS", lib_prep = "PolyA") %>%
  convert_to_hugo()

SS_riboD_medians <- compute_gene_medians_riboD(SS_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "SS", lib_prep = "RiboD") %>%
  convert_to_hugo()

WT_polyA_medians <- compute_gene_medians_polyA(WT_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "WT", lib_prep = "PolyA") %>%
  convert_to_hugo()

WT_riboD_medians <- compute_gene_medians_riboD(WT_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "WT", lib_prep = "RiboD") %>%
  convert_to_hugo()

NB_polyA_medians <- compute_gene_medians_polyA(NB_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "NB", lib_prep = "PolyA") %>%
  convert_to_hugo()

NB_riboD_medians <- compute_gene_medians_riboD(NB_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "NB", lib_prep = "RiboD") %>%
  convert_to_hugo()

ALL_polyA_medians <- compute_gene_medians_polyA(ALL_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "ALL", lib_prep = "PolyA") %>%
  convert_to_hugo()

ALL_riboD_medians <- compute_gene_medians_riboD(ALL_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "ALL", lib_prep = "RiboD") %>%
  convert_to_hugo()

AML_polyA_medians <- compute_gene_medians_polyA(AML_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "AML", lib_prep = "PolyA") %>%
  convert_to_hugo()

AML_riboD_medians <- compute_gene_medians_riboD(AML_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "AML", lib_prep = "RiboD") %>%
  convert_to_hugo()

DIPGIV_polyA_medians <- compute_gene_medians_polyA(THR13_polyA_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "DIPGIV", lib_prep = "PolyA") %>%
  convert_to_hugo()

DIPGIV_riboD_medians <- compute_gene_medians_riboD(THR13_riboD_log2tpm1) %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "DIPGIV", lib_prep = "RiboD") %>%
  convert_to_hugo()
```

Filtering for median ratio genes

```{r medians_ratio}
aRMS_polyA_medians_ratio <- filter_ratio(aRMS_polyA_medians, ratio_genes)
aRMS_riboD_medians_ratio <- filter_ratio(aRMS_riboD_medians, ratio_genes)

SS_polyA_medians_ratio <- filter_ratio(SS_polyA_medians, ratio_genes)
SS_riboD_medians_ratio <- filter_ratio(SS_riboD_medians, ratio_genes)

WT_polyA_medians_ratio <- filter_ratio(WT_polyA_medians, ratio_genes)
WT_riboD_medians_ratio <- filter_ratio(WT_riboD_medians, ratio_genes)

NB_polyA_medians_ratio <- filter_ratio(NB_polyA_medians, ratio_genes)
NB_riboD_medians_ratio <- filter_ratio(NB_riboD_medians, ratio_genes)

ALL_polyA_medians_ratio <- filter_ratio(ALL_polyA_medians, ratio_genes)
ALL_riboD_medians_ratio <- filter_ratio(ALL_riboD_medians, ratio_genes)

AML_polyA_medians_ratio <- filter_ratio(AML_polyA_medians, ratio_genes)
AML_riboD_medians_ratio <- filter_ratio(AML_riboD_medians, ratio_genes)

DIPGIV_polyA_medians_ratio <- filter_ratio(DIPGIV_polyA_medians, ratio_genes)
DIPGIV_riboD_medians_ratio <- filter_ratio(DIPGIV_riboD_medians, ratio_genes)
```

```{r combined_medians_ratio}
combined_medians_ratio <- bind_rows(
  aRMS_polyA_medians_ratio,
  aRMS_riboD_medians_ratio,
  SS_polyA_medians_ratio,
  SS_riboD_medians_ratio,
  WT_polyA_medians_ratio,
  WT_riboD_medians_ratio,
  NB_polyA_medians_ratio,
  NB_riboD_medians_ratio,
  ALL_polyA_medians_ratio,
  ALL_riboD_medians_ratio,
  AML_polyA_medians_ratio,
  AML_riboD_medians_ratio,
  DIPGIV_polyA_medians_ratio,
  DIPGIV_riboD_medians_ratio
)
```

```{r scale_color_compendia}
# Define consistent color-blind–safe palette
scale_color_compendia <- function() {
  scale_color_manual(values = c(
    "RiboD" = "#E69F00",  # yellow
    "PolyA" = "#0072B2"   # blue
  ))
}
```

Generate plots

```{r med_ratio_plot}
# Loop through diseases and generate plots
med_ratio_plot <- map(unique(combined_medians_ratio$Disease), function(d) {

 df <- combined_medians_ratio %>% 
  filter(Disease == d)

df <- combined_medians_ratio %>% 
  filter(Disease == d) %>%
  mutate(Gene = factor(Gene, levels = df %>% 
                         group_by(Gene) %>% 
                         summarize(median_expr = median(Expression), .groups = "drop") %>% 
                         arrange(median_expr) %>% 
                         pull(Gene)))

  ggplot(df, aes(x = Gene, y = Expression, color = lib_prep)) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.4) +
    # 1. vertical lines for each gene 
    geom_vline(aes(xintercept = as.numeric(Gene)), color = "grey85", linewidth = 0.3) + 
    # 2. points that stay centered on those lines 
    geom_point(size = 1.4, alpha = 0.75) +
    scale_x_discrete(expand = expansion(mult = c(0.01, 0.01))) +
    scale_y_continuous(limits = c(0, 11), expand = expansion(mult = c(0, 0.05))) +
    scale_color_compendia() +
    labs(
      title = paste(d),
      x = "Median Ratio Genes",
      # y = "Expression log2(TPM+1)",
      color = "Library Prep"
    ) +
    ylab(bquote(Median~log[2](TPM+1))) +
    theme(
      # axis.text.x = element_blank(),
      # axis.title.x = element_blank(), 
      # axis.ticks.x = element_blank(),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
      panel.grid.major.x = element_blank(),
      axis.text.y = element_text(angle = 0, hjust = 1, size = 12),
      axis.title.y = element_text(angle = 90, hjust = 0.5, size = 15), 
      plot.title = element_text(hjust = 0.5, face = "bold", size = 20),
      legend.position = "bottom"
    )
})

# Assign names for easy access
names(med_ratio_plot) <- unique(combined_medians_ratio$Disease)

# View plots
med_ratio_plot
```

Next, I'll try plotting each data point instead of the median expression

```{r}
aRMS_polyA_log2tpm1_hugo <- convert_to_hugo(aRMS_polyA_log2tpm1)
aRMS_riboD_log2tpm1_hugo <- convert_to_hugo(aRMS_riboD_log2tpm1)

SS_polyA_log2tpm1_hugo <- convert_to_hugo(SS_polyA_log2tpm1)
SS_riboD_log2tpm1_hugo <- convert_to_hugo(SS_riboD_log2tpm1)

WT_polyA_log2tpm1_hugo <- convert_to_hugo(WT_polyA_log2tpm1)
WT_riboD_log2tpm1_hugo <- convert_to_hugo(WT_riboD_log2tpm1)

NB_polyA_log2tpm1_hugo <- convert_to_hugo(NB_polyA_log2tpm1)
NB_riboD_log2tpm1_hugo <- convert_to_hugo(NB_riboD_log2tpm1)

ALL_polyA_log2tpm1_hugo <- convert_to_hugo(ALL_polyA_log2tpm1)
ALL_riboD_log2tpm1_hugo <- convert_to_hugo(ALL_riboD_log2tpm1)

AML_polyA_log2tpm1_hugo <- convert_to_hugo(AML_polyA_log2tpm1)
AML_riboD_log2tpm1_hugo <- convert_to_hugo(AML_riboD_log2tpm1)

DIPGIV_polyA_log2tpm1_hugo <- convert_to_hugo(THR13_polyA_log2tpm1)
DIPGIV_riboD_log2tpm1_hugo <- convert_to_hugo(THR13_riboD_log2tpm1)
```

```{r log2tpm1_ratio}
aRMS_polyA_log2tpm1_ratio <- filter_ratio(aRMS_polyA_log2tpm1_hugo, ratio_genes)
aRMS_riboD_log2tpm1_ratio <- filter_ratio(aRMS_riboD_log2tpm1_hugo, ratio_genes)

SS_polyA_log2tpm1_ratio <- filter_ratio(SS_polyA_log2tpm1_hugo, ratio_genes)
SS_riboD_log2tpm1_ratio <- filter_ratio(SS_riboD_log2tpm1_hugo, ratio_genes)

WT_polyA_log2tpm1_ratio <- filter_ratio(WT_polyA_log2tpm1_hugo, ratio_genes)
WT_riboD_log2tpm1_ratio <- filter_ratio(WT_riboD_log2tpm1_hugo, ratio_genes)

NB_polyA_log2tpm1_ratio <- filter_ratio(NB_polyA_log2tpm1_hugo, ratio_genes)
NB_riboD_log2tpm1_ratio <- filter_ratio(NB_riboD_log2tpm1_hugo, ratio_genes)

ALL_polyA_log2tpm1_ratio <- filter_ratio(ALL_polyA_log2tpm1_hugo, ratio_genes)
ALL_riboD_log2tpm1_ratio <- filter_ratio(ALL_riboD_log2tpm1_hugo, ratio_genes)

AML_polyA_log2tpm1_ratio <- filter_ratio(AML_polyA_log2tpm1_hugo, ratio_genes)
AML_riboD_log2tpm1_ratio <- filter_ratio(AML_riboD_log2tpm1_hugo, ratio_genes)

DIPGIV_polyA_log2tpm1_ratio <- filter_ratio(DIPGIV_polyA_log2tpm1_hugo, ratio_genes)
DIPGIV_riboD_log2tpm1_ratio <- filter_ratio(DIPGIV_riboD_log2tpm1_hugo, ratio_genes)
```

```{r log2tpm1_ratio_long}
aRMS_polyA_log2tpm1_ratio_long <- aRMS_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "aRMS", lib_prep = "PolyA")

aRMS_riboD_log2tpm1_ratio_long <- aRMS_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "aRMS", lib_prep = "RiboD")

SS_polyA_log2tpm1_ratio_long <- SS_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "SS", lib_prep = "PolyA")

SS_riboD_log2tpm1_ratio_long <- SS_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "SS", lib_prep = "RiboD")

WT_polyA_log2tpm1_ratio_long <- WT_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "WT", lib_prep = "PolyA")

WT_riboD_log2tpm1_ratio_long <- WT_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "WT", lib_prep = "RiboD")

NB_polyA_log2tpm1_ratio_long <- NB_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "NB", lib_prep = "PolyA")

NB_riboD_log2tpm1_ratio_long <- NB_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "NB", lib_prep = "RiboD")

ALL_polyA_log2tpm1_ratio_long <- ALL_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "ALL", lib_prep = "PolyA")

ALL_riboD_log2tpm1_ratio_long <- ALL_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "ALL", lib_prep = "RiboD")

AML_polyA_log2tpm1_ratio_long <- AML_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "AML", lib_prep = "PolyA")

AML_riboD_log2tpm1_ratio_long <- AML_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "AML", lib_prep = "RiboD")

DIPGIV_polyA_log2tpm1_ratio_long <- DIPGIV_polyA_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "DIPGIV", lib_prep = "PolyA")

DIPGIV_riboD_log2tpm1_ratio_long <- DIPGIV_riboD_log2tpm1_ratio %>%
  pivot_longer(-Gene, values_to = "Expression") %>%
  mutate(Disease = "DIPGIV", lib_prep = "RiboD")
```

```{r combined_log2tpm1_ratio}
combined_log2tpm1_ratio <- bind_rows(
  aRMS_polyA_log2tpm1_ratio_long,
  aRMS_riboD_log2tpm1_ratio_long,
  SS_polyA_log2tpm1_ratio_long,
  SS_riboD_log2tpm1_ratio_long,
  WT_polyA_log2tpm1_ratio_long,
  WT_riboD_log2tpm1_ratio_long,
  NB_polyA_log2tpm1_ratio_long,
  NB_riboD_log2tpm1_ratio_long,
  ALL_polyA_log2tpm1_ratio_long,
  ALL_riboD_log2tpm1_ratio_long,
  AML_polyA_log2tpm1_ratio_long,
  AML_riboD_log2tpm1_ratio_long,
  DIPGIV_polyA_log2tpm1_ratio_long,
  DIPGIV_riboD_log2tpm1_ratio_long
)
```

```{r log2tpm1_ratio_plot}
# Loop through diseases and generate plots
log2tpm1_ratio_plot <- map(unique(combined_log2tpm1_ratio$Disease), function(d) {

 df <- combined_log2tpm1_ratio %>% 
  filter(Disease == d)

df <- combined_log2tpm1_ratio %>% 
  filter(Disease == d) %>%
  mutate(Gene = factor(Gene, levels = df %>% 
                         group_by(Gene) %>% 
                         summarize(median_expr = median(Expression), .groups = "drop") %>% 
                         arrange(median_expr) %>% 
                         pull(Gene)))

  ggplot(df, aes(x = Gene, y = Expression, color = lib_prep)) +
    geom_quasirandom(
      dodge.width = 0.7,
      width = 0.15,
      alpha = .8,
      size = 1.2
  ) +
  #   geom_quasirandom(
  #     dodge.width = 0.7,
  #     width = 0.15,
  #     alpha = .8,
  #     size = 2
  # ) +

    # geom_hline(yintercept = 1, linetype = "dashed", color = "grey40", linewidth = 0.4) +
    # 1. vertical lines for each gene 
    # geom_vline(aes(xintercept = as.numeric(Gene)), color = "grey85", linewidth = 0.3) + 
    # # 2. points that stay centered on those lines 
    # geom_point(size = 1.4, alpha = 0.75) +
    # scale_x_discrete(expand = expansion(mult = c(0.01, 0.01))) +
    scale_y_continuous(limits = c(0, 11), expand = expansion(mult = c(0, 0.05))) +
    scale_color_compendia() +
    labs(
      title = paste(d),
      x = "Expression of Median Ratio Genes",
      # y = "Expression log2(TPM+1)",
      color = "Library Prep"
    ) +
    ylab(bquote(log[2](TPM+1))) +
    theme(
      # axis.text.x = element_blank(),
      # axis.title.x = element_blank(), 
      # axis.ticks.x = element_blank(),
      axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 8),
      panel.grid.major.x = element_blank(),
      axis.text.y = element_text(angle = 0, hjust = 1, size = 12),
      axis.title.y = element_text(angle = 90, hjust = 0.5, size = 15), 
      plot.title = element_text(hjust = 0.5, face = "bold", size = 20),
      legend.position = "bottom"
    )
})

# Assign names for easy access
names(log2tpm1_ratio_plot) <- unique(combined_log2tpm1_ratio$Disease)

# View plots
log2tpm1_ratio_plot
```

```{r sessioninfo}
sessioninfo::session_info()
```
