---
title: "THR13_0970_S12-S17_correlogram"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## THR13_0970 matched PolyA / RiboD

Correlogram showing similarity between datasets

```{r load library}
library(tidyverse)
library(corrr)
library(GGally)
library(corrplot)
```

```{r load in expr file}
rsem_log2TPM1_THR13 <- read_tsv("../input_data/matched_THR13_0970/rsem_ensembl_log2TPM1_THR13_0970_S12-S17.tsv.gz")
```

```{r wide}
rsem_log2TPM1_THR13_prepname <- rsem_log2TPM1_THR13 %>%
  mutate(
    prepname = case_when(
    full_sample == "THR13_0970_S12" ~ "PolyA_1",
    full_sample == "THR13_0970_S13" ~ "PolyA_2",
    full_sample == "THR13_0970_S14" ~ "PolyA_3",
    full_sample == "THR13_0970_S15" ~ "RiboD_1",
    full_sample == "THR13_0970_S16" ~ "RiboD_2",
    full_sample == "THR13_0970_S17" ~ "RiboD_3"
  ))

# pivot to wide format
rsem_wide <- rsem_log2TPM1_THR13_prepname %>%
  select(prepname, ensembl_gene_ID, log2TPM1) %>%
  pivot_wider(names_from = prepname, values_from = log2TPM1)
```

```{r corrr}
# convert to matrix
rsem_matrix <- rsem_wide %>%
  column_to_rownames("ensembl_gene_ID") %>%
  as.matrix()

set.seed(123) # for reproducibility
# Compute Spearman correlation 
rsem_correlation <- cor(rsem_matrix, method = "spearman", use = "pairwise.complete.obs") #pairwise.complete.obs = the correlation between each pair of variables is computed using all complete pairs of those particular variables
# ^ I think this means NAs will not be counted in calculation
```

```{r prep_map}
# prep_map <- c(
#   "THR13_0970_S12" = "PolyA",
#   "THR13_0970_S13" = "PolyA",
#   "THR13_0970_S14" = "PolyA",
#   "THR13_0970_S15" = "RiboD",
#   "THR13_0970_S16" = "RiboD",
#   "THR13_0970_S17" = "RiboD"
# )
# 
# # Rename columns to include prep method, e.g. "S12 (PolyA)"
# colnames(rsem_matrix) <- paste0(
#   colnames(rsem_matrix),
#   " (", prep_map[colnames(rsem_matrix)], ")"
# )
```

```{r cor_color_fn}
# integrate density plot with coefficient value and color by coefficient value

cor_color_fn <- function(data, mapping, ...) {
  corr <- eval_data_col(data, mapping$x) %>%
    cor(eval_data_col(data, mapping$y), method = "spearman", use = "pairwise.complete.obs")
  
  ggally_cor(data, mapping, method = "spearman", colour = I("black")) +
    theme_bw() +
    theme(panel.background = element_rect(
      fill = scales::col_numeric(c("#ec944d", "white", "#084c8b"), c(0.8,1))(corr)
    ),
    panel.grid = element_blank())
}
```

```{r rsem_ggpairs_plot}
# rsem_ggpairs_plot <- ggpairs(
#   rsem_matrix,
#   # upper = list(continuous = wrap(cor_color_fn, method = "spearman")),
#   lower = list(continuous = "blank"),
#   upper = list(continuous = wrap(cor_color_fn, method = "spearman")),
#   diag = list(continuous = wrap("densityDiag"))
#   # diag = list(continuous = "blank")
# ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   theme(
#     strip.text.y = element_text(angle = 0, size = 9),
#     strip.text.x = element_text(angle = 45, size = 9)
#   ) +
#   theme(panel.grid = element_blank())
# rsem_ggpairs_plot
```

```{r}
# ggsave(
#   "rsem_ggpairs_plot.png",
#   rsem_ggpairs_plot,
#   width = 10, height = 12, units = "in", dpi = 300
# )
```

```{r rsem_ggpairs_plot_scatter}
# # with scatterplot
# rsem_ggpairs_plot_scatter <- ggpairs(
#   rsem_matrix,
#   lower = list(continuous = wrap("points", alpha = 0.3, size = 0.5)),
#   upper = list(continuous = wrap(cor_color_fn, method = "spearman")),
#   diag = list(continuous = wrap("densityDiag"))
#   # diag = list(continuous = "blank")
# ) +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
#   theme(
#     strip.text.y = element_text(angle = 0, size = 9),
#     strip.text.x = element_text(angle = 45, size = 9)
#   ) +
#   theme(panel.grid = element_blank())
# rsem_ggpairs_plot_scatter
```

```{r}
# ggsave(
#   "rsem_ggpairs_plot_scatter.png",
#   rsem_ggpairs_plot_scatter,
#   width = 10, height = 12, units = "in", dpi = 300
# )
```

I want the diagonal density plot to only visualize genes that are expressed in that sample, rather than every gene.

```{r diag_filtered_density}
# function for diagonal density
diag_filtered_density <- function(data, mapping, ...) {
  # Pull out just the column being plotted on this diagonal panel
  x <- eval_data_col(data, mapping$x)
  
  # Keep only expressed genes (log2TPM > 0)
  x_filtered <- x[x > 0]
  
  ggplot(data.frame(x = x_filtered), aes(x = x)) +
    geom_density(fill = "grey70", color = "black") +
    theme_bw() +
    theme(panel.grid = element_blank())
}
```

```{r rsem_ggpairs_plot_scatter_diagfilter}
# with scatterplot
rsem_ggpairs_plot_scatter_diagfilter <- ggpairs(
  rsem_matrix,
  lower = list(continuous = wrap("points", alpha = 0.3, size = 0.5)),
  upper = list(continuous = wrap(cor_color_fn, method = "spearman")),
  diag = list(continuous = wrap(diag_filtered_density))
  # diag = list(continuous = "blank")
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.text.y = element_text(angle = 0, size = 9),
    strip.text.x = element_text(angle = 45, size = 9)
  ) +
  theme(panel.grid = element_blank())
rsem_ggpairs_plot_scatter_diagfilter
```

```{r}
# ggsave(
#   "rsem_ggpairs_plot_scatter_diagfilter.png",
#   rsem_ggpairs_plot_scatter_diagfilter,
#   width = 10, height = 12, units = "in", dpi = 300
# )
```

Making all scatterplot axes the same

```{r scatter_fixed_axes}
# calculating the max and min expression values
global_min <- min(rsem_matrix, na.rm = TRUE)
global_max <- max(rsem_matrix, na.rm = TRUE)

# function for scatterplot with fixed axes
scatter_fixed_axes <- function(data, mapping, ...) {
  ggplot(data, mapping) +
    geom_point(alpha = 0.3, size = 0.5) +
    xlim(global_min, global_max) +
    ylim(global_min, global_max) +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "green4") # with y=x slope intercept line
}
```

```{r rsem_ggpairs_plot_fixedaxes}
# with scatterplot
rsem_ggpairs_plot_fixedaxes <- ggpairs(
  rsem_matrix,
  lower = list(continuous = wrap(scatter_fixed_axes)),
  upper = list(continuous = wrap(cor_color_fn, method = "spearman")),
  diag = list(continuous = wrap(diag_filtered_density))
  # diag = list(continuous = "blank")
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.text.y = element_text(angle = 0, size = 9),
    strip.text.x = element_text(angle = 45, size = 9)
  ) +
  theme(panel.grid = element_blank())
rsem_ggpairs_plot_fixedaxes
```

```{r}
# ggsave(
#   "rsem_ggpairs_plot_fixedaxes.png",
#   rsem_ggpairs_plot_fixedaxes,
#   width = 10, height = 10, units = "in", dpi = 300
# )
```

Plotting number of expressed genes instead of a density plot

```{r diag_n_expressed}
diag_n_expressed <- function(data, mapping, ...) {
  x <- eval_data_col(data, mapping$x)
  n_expr <- sum(x > 0, na.rm = TRUE)
  
  ggplot() +
    annotate("text", x = 0.5, y = 0.5, label = paste0("N =\n", n_expr), size = 3) +
    xlim(0, 1) + ylim(0, 1) +
    theme_bw() +
    theme(panel.grid = element_blank(), axis.text = element_blank(),
          axis.ticks = element_blank(), axis.title = element_blank())
}
```

```{r rsem_ggpairs_plot_nexpressed}
# n expressed
rsem_ggpairs_plot_nexpressed <- ggpairs(
  rsem_matrix,
  lower = list(continuous = wrap(scatter_fixed_axes)),
  upper = list(continuous = wrap(cor_color_fn, method = "spearman")),
  diag = list(continuous = wrap(diag_n_expressed))
  # diag = list(continuous = "blank")
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.text.y = element_text(angle = 0, size = 9),
    strip.text.x = element_text(angle = 45, size = 9)
  ) +
  theme(panel.grid = element_blank())
rsem_ggpairs_plot_nexpressed
```

```{r}
# ggsave(
#   "rsem_ggpairs_plot_nexpressed.png",
#   rsem_ggpairs_plot_nexpressed,
#   width = 10, height = 10, units = "in", dpi = 300
# )
```

#### Applying a variance filter to correlogram

First, remove all genes where, for that gene, more than 80% of samples have an expression of 0 (in this case, remove genes where 5 out of 6 datasets have expression of 0).

From the remaining genes, sort them by variance (or standard deviation, either way) of expression within the matrix -- then remove the 20% of genes that are the least variant. (So if we had 30,000 genes after removing those that failed the expression filter, we would remove the bottom 6,000.)

```{r DIPGIV_expr_filter_zero_variance}
DIPGIV_expr_filter_zero <- rsem_wide %>%
  filter(rowSums(across(c(everything())) == 0) < 5)

nrow(rsem_wide)
nrow(DIPGIV_expr_filter_zero)

DIPGIV_expr_filter_zero_variance <- DIPGIV_expr_filter_zero %>%
  rowwise() %>%
  mutate(row_sd = sd(c_across(-ensembl_gene_ID), na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(row_sd) %>%
  slice_tail(prop = 0.8)

nrow(DIPGIV_expr_filter_zero_variance)
```

```{r DIPGIV_expr_filter_zero_variance_plot}

# integrate density plot with coefficient value and color by coefficient value

cor_color_fn2 <- function(data, mapping, ...) {
  corr <- eval_data_col(data, mapping$x) %>%
    cor(eval_data_col(data, mapping$y), method = "spearman", use = "pairwise.complete.obs")
  
  ggally_cor(data, mapping, method = "spearman", colour = I("black")) +
    theme_bw() +
    theme(panel.background = element_rect(
      fill = scales::col_numeric(c("#ec944d", "white", "#084c8b"), c(0.9,1))(corr)
    ),
    panel.grid = element_blank())
}

# with scatterplot
DIPGIV_expr_filter_zero_variance_plot <- DIPGIV_expr_filter_zero_variance %>%
  column_to_rownames("ensembl_gene_ID") %>%
  select(-row_sd) %>%
  ggpairs(
  DIPGIV_expr_filter_zero_variance,
  lower = list(continuous = wrap(scatter_fixed_axes)),
  upper = list(continuous = wrap(cor_color_fn2, method = "spearman")),
  diag = list(continuous = wrap(diag_filtered_density))
  # diag = list(continuous = "blank")
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.text.y = element_text(angle = 0, size = 9),
    strip.text.x = element_text(angle = 45, size = 9)
  ) +
  theme(panel.grid = element_blank())
DIPGIV_expr_filter_zero_variance_plot
```

```{r}
# ggsave(
#   "DIPGIV_expr_filter_zero_variance_plot.png",
#   DIPGIV_expr_filter_zero_variance_plot,
#   width = 10, height = 10, units = "in", dpi = 300
# )
```

Some genes have very low but not quite zero expression (i.e. ENSG00000000005.5 has values of 0.02856915, 0.01435529, 0.05658353, 0.16349873, 0.04264434, and 0.16349873). 

Going to remove genes where 5 or more out of 6 samples have expression \< 1

```{r DIPGIV_expr_filter_one_variance}
DIPGIV_expr_filter_one <- rsem_wide %>%
  filter(rowSums(across(c(everything())) < 1) < 5)

nrow(rsem_wide)
nrow(DIPGIV_expr_filter_one)

DIPGIV_expr_filter_one_variance <- DIPGIV_expr_filter_one %>%
  rowwise() %>%
  mutate(row_sd = sd(c_across(-ensembl_gene_ID), na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(row_sd) %>%
  slice_tail(prop = 0.8)

nrow(DIPGIV_expr_filter_one_variance)
```

```{r DIPGIV_expr_filter_one_variance_plot}
# integrate density plot with coefficient value and color by coefficient value

cor_color_fn3 <- function(data, mapping, ...) {
  corr <- eval_data_col(data, mapping$x) %>%
    cor(eval_data_col(data, mapping$y), method = "spearman", use = "pairwise.complete.obs")

  ggally_cor(data, mapping, method = "spearman", colour = I("black")) +
    theme_bw() +
    theme(panel.background = element_rect(
      fill = scales::col_numeric(c("#ec944d", "white", "#084c8b"), c(0.88,1))(corr)
    ),
    panel.grid = element_blank())
}

# with scatterplot
DIPGIV_expr_filter_one_variance_plot <- DIPGIV_expr_filter_one_variance %>%
  column_to_rownames("ensembl_gene_ID") %>%
  select(-row_sd) %>%
  ggpairs(
  DIPGIV_expr_filter_one_variance,
  lower = list(continuous = wrap(scatter_fixed_axes)),
  upper = list(continuous = wrap(cor_color_fn3, method = "spearman")),
  diag = list(continuous = wrap(diag_filtered_density))
  # diag = list(continuous = "blank")
) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(
    strip.text.y = element_text(angle = 0, size = 9),
    strip.text.x = element_text(angle = 45, size = 9)
  ) +
  theme(panel.grid = element_blank())
DIPGIV_expr_filter_one_variance_plot
```

```{r}
# ggsave(
#   "DIPGIV_expr_filter_one_variance_plot.png",
#   DIPGIV_expr_filter_one_variance_plot,
#   width = 10, height = 10, units = "in", dpi = 300
# )
```

Session Info

```{r sessioninfo}
sessioninfo::session_info()
```
