---
title: "NB_WT_DESeq2_analysis"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## **NB v WT - DESeq2 Results Analysis**

```{r load libraries}
library(tidyverse)
library(VennDiagram)
library(ggplot2)
```

```{r load in files}
WTPolyA_NBPolyA <- read_tsv("../../output_data/NB_WT/hugo_results_WT_polyA_NB_polyA_updated.tsv.gz")
WTRiboD_NBRiboD <- read_tsv("../../output_data/NB_WT/hugo_results_WT_riboD_NB_riboD_updated.tsv.gz")
WTPolyA_NBRiboD <- read_tsv("../../output_data/NB_WT/hugo_results_WT_polyA_NB_riboD_updated.tsv.gz")
WTRiboD_NBPolyA <- read_tsv("../../output_data/NB_WT/hugo_results_WT_riboD_NB_polyA_updated.tsv.gz")
```

## **L2FC \>/\<= 1 or -1**

Filter for statistically significant DEGs

```{r}
# Filter for genes with a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05

WTPolyA_NBPolyA_great1 <- WTPolyA_NBPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "WTPolyA_NBPolyA_great1") %>%
  relocate(comparison)

WTRiboD_NBRiboD_great1 <- WTRiboD_NBRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "WTRiboD_NBRiboD_great1") %>%
  relocate(comparison)

WTPolyA_NBRiboD_great1 <- WTPolyA_NBRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "WTPolyA_NBRiboD_great1") %>%
  relocate(comparison)

WTRiboD_NBPolyA_great1 <- WTRiboD_NBPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "WTRiboD_NBPolyA_great1") %>%
  relocate(comparison)

# Filter for genes with a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05

WTPolyA_NBPolyA_less1 <- WTPolyA_NBPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "WTPolyA_NBPolyA_less1") %>%
  relocate(comparison)

WTRiboD_NBRiboD_less1 <- WTRiboD_NBRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "WTRiboD_NBRiboD_less1") %>%
  relocate(comparison)

WTPolyA_NBRiboD_less1 <- WTPolyA_NBRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "WTPolyA_NBRiboD_less1") %>%
  relocate(comparison)

WTRiboD_NBPolyA_less1 <- WTRiboD_NBPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "WTRiboD_NBPolyA_less1") %>%
  relocate(comparison)
```

### **PolyA unbiased against biased**

#### **Up DEGs**

```{r WT_NB_up_polyAunbiased_polyAbiased_riboDbiased}
WT_NB_up_polyAunbiased_polyAbiased_riboDbiased <- list(
  "Up in WT polyA relative to NB polyA (polyA unbiased)" = WTPolyA_NBPolyA_great1$Gene,
  "Up in WT polyA relative to NB riboD (polyA biased)" = WTPolyA_NBRiboD_great1$Gene,
    "Up in WT riboD relative to NB polyA (riboD biased)" = WTRiboD_NBPolyA_great1$Gene
)

write_rds(WT_NB_up_polyAunbiased_polyAbiased_riboDbiased,
  "../../output_data/NB_WT/WT_NB_up_polyAunbiased_polyAbiased_riboDbiased.rds")
```

```{r WT_NB_up_polyAunbiased_polyAbiased_riboDbiased_VD}
WT_NB_up_polyAunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = WT_NB_up_polyAunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Up in WT polyA relative to \n NB polyA (polyA unbiased)",
    "Up in WT polyA relative to \n NB riboD (polyA biased)",
    "Up in WT riboD relative to \n NB polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in WT relative to NB (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)

WT_NB_up_polyAunbiased_polyAbiased_riboDbiased_VD
```

#### Down DEGs

```{r WT_NB_down_polyAunbiased_polyAbiased_riboDbiased}
WT_NB_down_polyAunbiased_polyAbiased_riboDbiased <- list(
  "Down in WT polyA relative to NB polyA (polyA unbiased)" = WTPolyA_NBPolyA_less1$Gene,
  "Down in WT polyA relative to NB riboD (polyA biased)" = WTPolyA_NBRiboD_less1$Gene,
  "Down in WT riboD relative to NB polyA (riboD biased)" = WTRiboD_NBPolyA_less1$Gene
)

write_rds(
  WT_NB_down_polyAunbiased_polyAbiased_riboDbiased,
  "../../output_data/NB_WT/WT_NB_down_polyAunbiased_polyAbiased_riboDbiased.rds")
```

```{r WT_NB_down_polyAunbiased_polyAbiased_riboDbiased_VD}
WT_NB_down_polyAunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = WT_NB_down_polyAunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Down in WT polyA relative to \n NB polyA (polyA unbiased)",
    "Down in WT polyA relative to \n NB riboD (polyA biased)",
    "Down in WT riboD relative to \n NB polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in WT relative to NB (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
WT_NB_down_polyAunbiased_polyAbiased_riboDbiased_VD
```

### RiboD unbiased against polyA biased and riboD biased

#### Up DEGs

```{r WT_NB_up_riboDunbiased_polyAbiased_riboDbiased}
WT_NB_up_riboDunbiased_polyAbiased_riboDbiased <- list(
  "Up in WT riboD relative to NB riboD (riboD unbiased)" = WTRiboD_NBRiboD_great1$Gene,
  "Up in WT polyA relative to NB riboD (polyA biased)" = WTPolyA_NBRiboD_great1$Gene,
    "Up in WT riboD relative to NB polyA (riboD biased)" = WTRiboD_NBPolyA_great1$Gene
)

write_rds(
  WT_NB_up_riboDunbiased_polyAbiased_riboDbiased,
  "../../output_data/NB_WT/WT_NB_up_riboDunbiased_polyAbiased_riboDbiased.rds")
```

```{r WT_NB_up_riboDunbiased_polyAbiased_riboDbiased_VD}
WT_NB_up_riboDunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = WT_NB_up_riboDunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Up in WT riboD relative to \n NB riboD (riboD unbiased)",
    "Up in WT polyA relative to \n NB riboD (polyA biased)",
    "Up in WT riboD relative to \n NB polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in WT relative to NB (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
WT_NB_up_riboDunbiased_polyAbiased_riboDbiased_VD
```

### Down DEGs

```{r WT_NB_down_riboDunbiased_polyAbiased_riboDbiased}
WT_NB_down_riboDunbiased_polyAbiased_riboDbiased <- list(
  "Down in WT riboD relative to NB riboD (riboD unbiased)" = WTRiboD_NBRiboD_less1$Gene,
  "Down in WT polyA relative to NB riboD (polyA biased)" = WTPolyA_NBRiboD_less1$Gene,
   "Down in WT riboD relative to NB polyA (riboD biased)" = WTRiboD_NBPolyA_less1$Gene
)

write_rds(WT_NB_down_riboDunbiased_polyAbiased_riboDbiased, "../../output_data/NB_WT/WT_NB_down_riboDunbiased_polyAbiased_riboDbiased_upset.rds")
```

```{r WT_NB_down_riboDunbiased_polyAbiased_riboDbiased_VD}
WT_NB_down_riboDunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = WT_NB_down_riboDunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Down in WT riboD relative to \n NB riboD (riboD unbiased)",
    "Down in WT polyA relative to \n NB riboD (polyA biased)",
    "Down in WT riboD relative to \n NB polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in WT relative to NB (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
WT_NB_down_riboDunbiased_polyAbiased_riboDbiased_VD
```

### Histograms of LFC of genes in intersections

#### Up PolyA unbiased against polyA biased and riboD biased

Starting with upregulated genes in the polyA unbiased:polyA biased:riboD biased comparison

Looking at genes that are unique to the polyAbiased

```{r WT_NB_up_polyAunbiased_uniquetopolyAbiased}
WT_NB_up_polyAunbiased_uniquetopolyAbiased <- WTPolyA_NBRiboD_great1 %>%
  anti_join(WTPolyA_NBPolyA_great1, by = "Gene") %>%
  anti_join(WTRiboD_NBPolyA_great1, by = "Gene")
print(WT_NB_up_polyAunbiased_uniquetopolyAbiased %>% nrow())
```

What is the average LFC?

```{r WT_NB_up_polyAunbiased_uniquetopolyAbiased_avgLFC}
WT_NB_up_polyAunbiased_uniquetopolyAbiased_avgLFC <- WT_NB_up_polyAunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_up_polyAunbiased_uniquetopolyAbiased_avgLFC
```

```{r WT_NB_up_polyAunbiased_uniquetopolyAbiased_hist}
WT_NB_up_polyAunbiased_uniquetopolyAbiased_hist <- ggplot(WT_NB_up_polyAunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
  scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,NA)) +
  scale_x_continuous(breaks = seq(2, 30, by = 2)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in WT polyA relative to NB riboD",
       subtitle = "polyA biased, compared to polyA unbiased, Median LFC = 1.48",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_up_polyAunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboD biased

```{r WT_NB_up_polyAunbiased_uniquetoriboDbiased}
WT_NB_up_polyAunbiased_uniquetoriboDbiased <- WTRiboD_NBPolyA_great1 %>%
  anti_join(WTPolyA_NBPolyA_great1, by = "Gene") %>%
  anti_join(WTPolyA_NBRiboD_great1, by = "Gene")
print(WT_NB_up_polyAunbiased_uniquetoriboDbiased %>% nrow())
```

```{r WT_NB_polyAunbiased_uniquetoriboDbiased_avgLFC}
WT_NB_polyAunbiased_uniquetoriboDbiased_avgLFC <- WT_NB_up_polyAunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_polyAunbiased_uniquetoriboDbiased_avgLFC
```

```{r WT_NB_up_polyAunbiased_uniquetoriboDbiased_hist}
WT_NB_up_polyAunbiased_uniquetoriboDbiased_hist <- ggplot(WT_NB_up_polyAunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
  scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,NA)) +
    scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in WT riboD relative to NB polyA",
       subtitle = "riboD biased, compared to polyA unbiased, Median LFC = 2.69",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_up_polyAunbiased_uniquetoriboDbiased_hist
```

#### Down PolyA unbiased against polyA biased and riboD biased

Looking at genes that are unique to polyA biased

```{r WT_NB_down_polyAunbiased_uniquetopolyAbiased}
WT_NB_down_polyAunbiased_uniquetopolyAbiased <- WTPolyA_NBRiboD_less1 %>%
  anti_join(WTPolyA_NBPolyA_less1, by = "Gene") %>%
  anti_join(WTRiboD_NBPolyA_less1, by = "Gene")
print(WT_NB_down_polyAunbiased_uniquetopolyAbiased %>% nrow())
```

```{r WT_NB_down_polyAunbiased_uniquetopolyAbiased_avgLFC}
WT_NB_down_polyAunbiased_uniquetopolyAbiased_avgLFC <- WT_NB_down_polyAunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_down_polyAunbiased_uniquetopolyAbiased_avgLFC
```

```{r WT_NB_down_polyAunbiased_uniquetopolyAbiased_hist}
WT_NB_down_polyAunbiased_uniquetopolyAbiased_hist <- ggplot(WT_NB_down_polyAunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
  # scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,NA)) +
    # scale_x_continuous(
    #   breaks = c(0,-1,-2,-4,-6,-8,-10,-12,-14,-16,-18,-20,-22,-24,-26)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in WT PolyA relative to NB RiboD",
       subtitle = "polyA biased, compared to polyA unbiased, Median LFC = -3.15",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_down_polyAunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboD biased

```{r WT_NB_down_polyAunbiased_uniquetoriboDbiased}
WT_NB_down_polyAunbiased_uniquetoriboDbiased <- WTRiboD_NBPolyA_less1 %>%
  anti_join(WTPolyA_NBPolyA_less1, by = "Gene") %>%
  anti_join(WTPolyA_NBRiboD_less1, by = "Gene")
print(WT_NB_down_polyAunbiased_uniquetoriboDbiased %>% nrow())
```

```{r WT_NB_down_polyAunbiased_uniquetoriboDbiased_avgLFC}
WT_NB_down_polyAunbiased_uniquetoriboDbiased_avgLFC <- WT_NB_down_polyAunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_down_polyAunbiased_uniquetoriboDbiased_avgLFC
```

```{r WT_NB_down_polyAunbiased_uniquetoriboDbiased_hist}
WT_NB_down_polyAunbiased_uniquetoriboDbiased_hist <- ggplot(WT_NB_down_polyAunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
  # scale_y_continuous(breaks = c(200, 400, 600), limits = c(0,800)) +
  #   scale_x_continuous(breaks = c(0,-1,-2,-4,-6,-8,-10,-12,-14,-16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in WT riboD relative to NB polyA",
       subtitle = "riboD biased, compared to polyA unbiased, Median LFC = -1.39",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_down_polyAunbiased_uniquetoriboDbiased_hist
```

#### Up RiboD unbiased against polyA biased and riboD biased

Starting with upregulated genes in the riboD unbiased:polyAbiased:riboDbiased comparison

Looking at genes that are unique to the polyAbiased

```{r WT_NB_up_riboDunbiased_uniquetopolyAbiased}
WT_NB_up_riboDunbiased_uniquetopolyAbiased <- WTPolyA_NBRiboD_great1 %>%
  anti_join(WTRiboD_NBRiboD_great1, by = "Gene") %>%
  anti_join(WTRiboD_NBPolyA_great1, by = "Gene")
print(WT_NB_up_riboDunbiased_uniquetopolyAbiased %>% nrow())
```

```{r WT_NB_up_riboDunbiased_uniquetopolyAbiased_avgLFC}
WT_NB_up_riboDunbiased_uniquetopolyAbiased_avgLFC <- WT_NB_up_riboDunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_up_riboDunbiased_uniquetopolyAbiased_avgLFC
```

```{r WT_NB_up_riboDunbiased_uniquetopolyAbiased_hist}
WT_NB_up_riboDunbiased_uniquetopolyAbiased_hist <- ggplot(WT_NB_up_riboDunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
  # scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,NA)) +
  # scale_x_continuous(breaks = seq(2, 30, by = 2)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in WT polyA relative to NB riboD",
       subtitle = "polyA biased, compared to riboD unbiased, Median LFC = 1.51",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))

WT_NB_up_riboDunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboD biased

```{r WT_NB_up_riboDunbiased_uniquetoriboDbiased}
WT_NB_up_riboDunbiased_uniquetoriboDbiased <- WTRiboD_NBPolyA_great1 %>%
  anti_join(WTRiboD_NBRiboD_great1, by = "Gene") %>%
  anti_join(WTPolyA_NBRiboD_great1, by = "Gene")
print(WT_NB_up_riboDunbiased_uniquetoriboDbiased %>% nrow())
```

```{r WT_NB_up_riboDunbiased_uniquetoriboDbiased_avgLFC}
WT_NB_up_riboDunbiased_uniquetoriboDbiased_avgLFC <- WT_NB_up_riboDunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_up_riboDunbiased_uniquetoriboDbiased_avgLFC
```

```{r WT_NB_up_riboDunbiased_uniquetoriboDbiased_hist}
WT_NB_up_riboDunbiased_uniquetoriboDbiased_hist <- ggplot(WT_NB_up_riboDunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
  # scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,NA)) +
  #   scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in WT riboD relative to NB polyA",
       subtitle = "riboD biased, compared to riboD unbiased, Median LFC = 2.69",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_up_riboDunbiased_uniquetoriboDbiased_hist
```

#### Down RiboD unbiased against polyA biased and riboD biased

Looking at genes that are unique to polyA biased

```{r WT_NB_down_riboDunbiased_uniquetopolyAbiased}
WT_NB_down_riboDunbiased_uniquetopolyAbiased <- WTPolyA_NBRiboD_less1 %>%
  anti_join(WTRiboD_NBRiboD_less1, by = "Gene") %>%
  anti_join(WTRiboD_NBPolyA_less1, by = "Gene")
print(WT_NB_down_riboDunbiased_uniquetopolyAbiased %>% nrow())
```

```{r WT_NB_down_riboDunbiased_uniquetopolyAbiased_avgLFC}
WT_NB_down_riboDunbiased_uniquetopolyAbiased_avgLFC <- WT_NB_down_riboDunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_down_riboDunbiased_uniquetopolyAbiased_avgLFC
```

```{r WT_NB_down_riboDunbiased_uniquetopolyAbiased_hist}
WT_NB_down_riboDunbiased_uniquetopolyAbiased_hist <- ggplot(WT_NB_down_riboDunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
  # scale_y_continuous(breaks = c(100,200), limits = c(0,NA)) +
  #   scale_x_continuous(breaks =  seq(0, -30, by = -2)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in WT PolyA relative to NB RiboD",
       subtitle = "polyA biased compared to riboD unbiased, Median LFC = -2.94",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_down_riboDunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboD biased

```{r WT_NB_down_riboDunbiased_uniquetoriboDbiased}
WT_NB_down_riboDunbiased_uniquetoriboDbiased <- WTRiboD_NBPolyA_less1 %>%
  anti_join(WTRiboD_NBRiboD_less1, by = "Gene") %>%
  anti_join(WTPolyA_NBRiboD_less1, by = "Gene")
print(WT_NB_down_riboDunbiased_uniquetoriboDbiased %>% nrow())
```

```{r}
WT_NB_down_riboDunbiased_uniquetoriboDbiased_avgLFC <- WT_NB_down_riboDunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
WT_NB_down_riboDunbiased_uniquetoriboDbiased_avgLFC
```

```{r WT_NB_down_riboDunbiased_uniquetoriboDbiased_hist}
WT_NB_down_riboDunbiased_uniquetoriboDbiased_hist <- ggplot(WT_NB_down_riboDunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
  # scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,1000)) +
  #   scale_x_continuous(breaks = c(0,-1,-2,-4,-6,-8,-10,-12,-14,-16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in WT riboD relative to NB polyA",
       subtitle = "riboD biased, compared to riboD unbiased, Median LFC = -1.41",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_down_riboDunbiased_uniquetoriboDbiased_hist
```

Genes in common between the polyAunbiased:polyAbiased:riboDbiased venn diagram and the riboDunbiased:polyAbiased:riboDbiased venn diagram

upregulated polyA biased

```{r WT_NB_up_polyAbiased}
WT_NB_up_polyAbiased <- inner_join(WT_NB_up_polyAunbiased_uniquetopolyAbiased, WT_NB_up_riboDunbiased_uniquetopolyAbiased, by = "Gene")

write_rds(WT_NB_up_polyAbiased, "../../output_data/NB_WT/WT_NB_up_polyAbiased.rds")

print(WT_NB_up_polyAbiased %>% nrow())
```

```{r WT_NB_up_polyAbiased_avgLFC}
WT_NB_up_polyAbiased_avgLFC <- WT_NB_up_polyAbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
WT_NB_up_polyAbiased_avgLFC
```

```{r WT_NB_up_polyAbiased_hist}
WT_NB_up_polyAbiased_hist <- ggplot(WT_NB_up_polyAbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in WT polyA relative to NB riboD",
       subtitle = "polyA biased, Median LFC = 1.44",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_up_polyAbiased_hist
```

upregulated riboD biased

```{r WT_NB_up_riboDbiased}
WT_NB_up_riboDbiased <- inner_join(WT_NB_up_polyAunbiased_uniquetoriboDbiased, WT_NB_up_riboDunbiased_uniquetoriboDbiased, by = "Gene")

write_rds(WT_NB_up_riboDbiased, "../../output_data/NB_WT/WT_NB_up_riboDbiased.rds")

print(WT_NB_up_riboDbiased %>% nrow())
```

```{r WT_NB_up_riboDbiased_avgLFC}
WT_NB_up_riboDbiased_avgLFC <- WT_NB_up_riboDbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
WT_NB_up_riboDbiased_avgLFC
```

```{r WT_NB_up_riboDbiased_hist}
WT_NB_up_riboDbiased_hist <- ggplot(WT_NB_up_riboDbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in WT riboD relative to NB polyA",
       subtitle = "riboD biased, Median LFC = 2.59",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_up_riboDbiased_hist
```

downregulated polyA biased

```{r WT_NB_down_polyAbiased}
WT_NB_down_polyAbiased <- inner_join(WT_NB_down_polyAunbiased_uniquetopolyAbiased, WT_NB_down_riboDunbiased_uniquetopolyAbiased, by = "Gene")

write_rds(WT_NB_down_polyAbiased, "../../output_data/NB_WT/WT_NB_down_polyAbiased.rds")

print(WT_NB_down_polyAbiased %>% nrow())
```

```{r WT_NB_down_polyAbiased_avgLFC}
WT_NB_down_polyAbiased_avgLFC <- WT_NB_down_polyAbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
WT_NB_down_polyAbiased_avgLFC
```

```{r WT_NB_down_polyAbiased_hist}
WT_NB_down_polyAbiased_hist <- ggplot(WT_NB_down_polyAbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in WT polyA relative to NB riboD",
       subtitle = "polyA biased, Median LFC = -3.06",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_down_polyAbiased_hist
```

downregulated riboD biased

```{r WT_NB_down_riboDbiased}
WT_NB_down_riboDbiased <- inner_join(WT_NB_down_polyAunbiased_uniquetoriboDbiased, WT_NB_down_riboDunbiased_uniquetoriboDbiased, by = "Gene")

write_rds(WT_NB_down_riboDbiased, "../../output_data/NB_WT/WT_NB_down_riboDbiased.rds")

print(WT_NB_down_riboDbiased %>% nrow())
```

```{r WT_NB_down_riboDbiased_avgLFC}
WT_NB_down_riboDbiased_avgLFC <- WT_NB_down_riboDbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
WT_NB_down_riboDbiased_avgLFC
```

```{r WT_NB_down_riboDbiased_hist}
WT_NB_down_riboDbiased_hist <- ggplot(WT_NB_down_riboDbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in WT riboD relative to NB polyA",
       subtitle = "riboD biased, Median LFC = -1.33",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
            axis.text = element_text(size = 10),
        plot.subtitle = element_text(size = 8, face = "bold"))
WT_NB_down_riboDbiased_hist
```

### Save Session Info

```{r}
sessioninfo::session_info()
```
