---
title: "aRMS_SS_DESeq2_analysis"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## **aRMS v SS - DESeq2 Results Analysis**

```{r load libraries}
library(tidyverse)
library(VennDiagram)
library(ggplot2)
```

```{r load in files}
armsPolyA_ssPolyA <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_polyA_SS_polyA_updated.tsv.gz")
armsRiboD_ssRiboD <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_riboD_SS_riboD_updated.tsv.gz")
armsPolyA_ssRiboD <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_polyA_SS_riboD_updated.tsv.gz")
armsRiboD_ssPolyA <- read_tsv("../../output_data/aRMS_SS/hugo_results_aRMS_riboD_SS_polyA_updated.tsv.gz")
```

## **L2FC \>/\<= 1 or -1**

Filter for statistically significant DEGs

```{r}
# Filter for genes with a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05

armsPolyA_ssPolyA_great1 <- armsPolyA_ssPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "armsPolyA_ssPolyA_great1") %>%
  relocate(comparison)

armsRiboD_ssRiboD_great1 <- armsRiboD_ssRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "armsRiboD_ssRiboD_great1") %>%
  relocate(comparison)

armsPolyA_ssRiboD_great1 <- armsPolyA_ssRiboD %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "armsPolyA_ssRiboD_great1") %>%
  relocate(comparison)

armsRiboD_ssPolyA_great1 <- armsRiboD_ssPolyA %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  mutate(comparison = "armsRiboD_ssPolyA_great1") %>%
  relocate(comparison)

# Filter for genes with a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05

armsPolyA_ssPolyA_less1 <- armsPolyA_ssPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "armsPolyA_ssPolyA_less1") %>%
  relocate(comparison)

armsRiboD_ssRiboD_less1 <- armsRiboD_ssRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "armsRiboD_ssRiboD_less1") %>%
  relocate(comparison)

armsPolyA_ssRiboD_less1 <- armsPolyA_ssRiboD %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "armsPolyA_ssRiboD_less1") %>%
  relocate(comparison)

armsRiboD_ssPolyA_less1 <- armsRiboD_ssPolyA %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  mutate(comparison = "armsRiboD_ssPolyA_less1") %>%
  relocate(comparison)
```

### **PolyA unbiased against biased**

#### **Up DEGs**

```{r aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased}
aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased <- list(
  "Up in SS polyA relative to aRMS polyA (polyA unbiased)" = armsPolyA_ssPolyA_great1$Gene,
  "Up in SS polyA relative to aRMS riboD (polyA biased)" = armsRiboD_ssPolyA_great1$Gene,
  "Up in SS riboD relative to aRMS polyA (riboD biased)" = armsPolyA_ssRiboD_great1$Gene
)

write_rds(aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased, "../../output_data/aRMS_SS/aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased.rds")
```

```{r aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased_VD}
aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Up in SS polyA relative to \n aRMS polyA (polyA unbiased)",
    "Up in SS polyA relative to \n aRMS riboD (polyA biased)",
    "Up in SS riboD relative to \n aRMS polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
#  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in SS relative to aRMS (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
aRMS_SS_up_polyAunbiased_polyAbiased_riboDbiased_VD
```

#### **Down DEGs**

```{r aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased}
aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased <- list(
  "Down in SS polyA relative to aRMS polyA (polyA unbiased)" = armsPolyA_ssPolyA_less1$Gene,
  "Down in SS polyA relative to aRMS riboD (polyA biased)" = armsRiboD_ssPolyA_less1$Gene,
  "Down in SS riboD relative to aRMS polyA (riboD biased)" = armsPolyA_ssRiboD_less1$Gene
)

write_rds(aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased, "../../output_data/aRMS_SS/aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased.rds")
```

```{r aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased_VD}
aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Down in SS polyA relative to \n aRMS polyA (polyA unbiased)",
    "Down in SS polyA relative to \n aRMS riboD (polyA biased)",
    "Down in SS riboD relative to \n aRMS polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in SS relative to aRMS (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
aRMS_SS_down_polyAunbiased_polyAbiased_riboDbiased_VD
```

### **RiboD unbiased against biased**

#### **Up DEGs**

```{r aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased}
aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased <- list(
  "Up in SS riboD relative to aRMS riboD (riboD unbiased)" = armsRiboD_ssRiboD_great1$Gene,
  "Up in SS polyA relative to aRMS riboD (polyA biased)" = armsRiboD_ssPolyA_great1$Gene,
  "Up in SS riboD relative to aRMS polyA (riboD biased)" = armsPolyA_ssRiboD_great1$Gene
)

write_rds(aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased, "../../output_data/aRMS_SS/aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased.rds")
```

```{r aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased_VD}
aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Up in SS riboD relative to \n aRMS riboD (riboD unbiased)",
    "Up in SS polyA relative to \n aRMS riboD (polyA biased)",
    "Up in SS riboD relative to \n aRMS polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes upregulated in SS relative to aRMS (L2FC >= 1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
aRMS_SS_up_riboDunbiased_polyAbiased_riboDbiased_VD
```

#### **Down DEGs**

```{r aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased}
aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased <- list(
  "Down in SS riboD relative to aRMS riboD (riboD unbiased)" = armsRiboD_ssRiboD_less1$Gene,
  "Down in SS polyA relative to aRMS riboD (polyA biased)" = armsRiboD_ssPolyA_less1$Gene,
  "Down in SS riboD relative to aRMS polyA (riboD biased)" = armsPolyA_ssRiboD_less1$Gene
)

write_rds(aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased, "../../output_data/aRMS_SS/aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased.rds")
```

```{r aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased_VD}
aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased_VD <- venn.diagram(
  x = aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased,
  category.names = c(
    "Down in SS riboD relative to \n aRMS riboD (riboD unbiased)",
    "Down in SS polyA relative to \n aRMS riboD (polyA biased)",
    "Down in SS riboD relative to \n aRMS polyA (riboD biased)"
    ),
  filename = NULL, # Save as a PNG file
  output = TRUE,
  print.mode = c("raw", "percent"),
  # Customize appearance (optional)
  fill = c("#BB5566", "#0072B2", "#E69F00"),
  cat.col = c("#BB5566", "#0072B2", "#E69F00"),
  cex = 1, # Font size for counts
  cat.cex = 0.65, # Font size for category names
  cat.dist = c(0.05, 0.05, 0.05),
  height = 2000,
  width = 2000,
  cat.default.pos = "outer",
  cat.pos = c(-12, 12, 175), 
  main = "Genes downregulated in SS relative to aRMS (L2FC <= -1 and  p-adj < 0.05)",
  main.cex = 0.75, # Font size for main title
  disable.logging = TRUE
)
aRMS_SS_down_riboDunbiased_polyAbiased_riboDbiased_VD
```

### **Histograms of LFC of genes in intersections**

Starting with upregulated genes in the polyAunbiased:polyAbiased:riboDbiased comparison.

Looking at genes that are unique to polyAbiased

```{r aRMS_SS_up_polyAunbiased_uniquetopolyAbiased}
aRMS_SS_up_polyAunbiased_uniquetopolyAbiased <- armsRiboD_ssPolyA_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene")
print(aRMS_SS_up_polyAunbiased_uniquetopolyAbiased %>% nrow())
```

What is the average LFC?

```{r aRMS_SS_up_polyAunbiased_uniquetopolyAbiased_avgLFC}
aRMS_SS_up_polyAunbiased_uniquetopolyAbiased_avgLFC <- aRMS_SS_up_polyAunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_up_polyAunbiased_uniquetopolyAbiased_avgLFC
```

Histogram of LFC

```{r aRMS_SS_up_polyAunbiased_uniquetopolyAbiased_hist}
aRMS_SS_up_polyAunbiased_uniquetopolyAbiased_hist <- ggplot(aRMS_SS_up_polyAunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in SS polyA relative to aRMS riboD",
       subtitle = "polyAbiased, compared to polyAunbiased. Median LFC = 1.37",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_up_polyAunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboDbiased

```{r aRMS_SS_up_polyAunbiased_uniquetoriboDbiased}
aRMS_SS_up_polyAunbiased_uniquetoriboDbiased <- armsPolyA_ssRiboD_great1 %>%
  anti_join(armsPolyA_ssPolyA_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene")
print(aRMS_SS_up_polyAunbiased_uniquetoriboDbiased %>% nrow())
```

What’s the average LFC?

```{r aRMS_SS_up_polyAunbiased_uniquetoriboDbiased_avgLFC}
aRMS_SS_up_polyAunbiased_uniquetoriboDbiased_avgLFC <- aRMS_SS_up_polyAunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_up_polyAunbiased_uniquetoriboDbiased_avgLFC
```

```{r aRMS_SS_up_polyAunbiased_uniquetoriboDbiased_hist}
aRMS_SS_up_polyAunbiased_uniquetoriboDbiased_hist <- ggplot(aRMS_SS_up_polyAunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800, 1000), limits = c(0,1000)) +
#    scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, compared to polyA_unbiased. Median LFC = 2.81",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_up_polyAunbiased_uniquetoriboDbiased_hist
```

Upregulated genes in the riboD unbiased:polyAbiased:riboDbiased comparison

Looking at genes that are unique to the polyA biased

```{r aRMS_SS_up_riboDunbiased_uniquetopolyAbiased}
aRMS_SS_up_riboDunbiased_uniquetopolyAbiased <- armsRiboD_ssPolyA_great1 %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_great1, by = "Gene")
print(aRMS_SS_up_riboDunbiased_uniquetopolyAbiased %>% nrow())
```

```{r aRMS_SS_up_riboDunbiased_uniquetopolyAbiased_avgLFC}
aRMS_SS_up_riboDunbiased_uniquetopolyAbiased_avgLFC <- aRMS_SS_up_riboDunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_up_riboDunbiased_uniquetopolyAbiased_avgLFC
```

```{r aRMS_SS_up_riboDunbiased_uniquetopolyAbiased_hist}
aRMS_SS_up_riboDunbiased_uniquetopolyAbiased_hist <- ggplot(aRMS_SS_up_riboDunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in SS polyA relative to aRMS riboD",
       subtitle = "polyA biased, compared to RiboD_unbiased. Median LFC = 1.39",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_up_riboDunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to riboDbiased

```{r aRMS_SS_up_riboDunbiased_uniquetoriboDbiased}
aRMS_SS_up_riboDunbiased_uniquetoriboDbiased <- armsPolyA_ssRiboD_great1 %>%
  anti_join(armsRiboD_ssRiboD_great1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_great1, by = "Gene")
print(aRMS_SS_up_riboDunbiased_uniquetoriboDbiased %>% nrow())
```

```{r aRMS_SS_up_riboDunbiased_uniquetoriboDbiased_avgLFC}
aRMS_SS_up_riboDunbiased_uniquetoriboDbiased_avgLFC <- aRMS_SS_up_riboDunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_up_riboDunbiased_uniquetoriboDbiased_avgLFC
```

```{r aRMS_SS_up_riboDunbiased_uniquetoriboDbiased_hist}
aRMS_SS_up_riboDunbiased_uniquetoriboDbiased_hist <- ggplot(aRMS_SS_up_riboDunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30,
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800, 1000, 1200), limits = c(0,1200)) +
#    scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, compared to riboD_unbiased. Median LFC = 2.65",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_up_riboDunbiased_uniquetoriboDbiased_hist
```

Looking at downregulated genes in the polyA unbiased:polyA biased:riboD biased comparison

Looking at genes that are unique to the polyA biased

```{r aRMS_SS_down_polyAunbiased_uniquetopolyAbiased}
aRMS_SS_down_polyAunbiased_uniquetopolyAbiased <- armsRiboD_ssPolyA_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene")
print(aRMS_SS_down_polyAunbiased_uniquetopolyAbiased %>% nrow())
```

```{r aRMS_SS_down_polyAunbiased_uniquetopolyAbiased_avgLFC}
aRMS_SS_down_polyAunbiased_uniquetopolyAbiased_avgLFC <- aRMS_SS_down_polyAunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_down_polyAunbiased_uniquetopolyAbiased_avgLFC
```

```{r aRMS_SS_down_polyAunbiased_uniquetopolyAbiased_hist}
aRMS_SS_down_polyAunbiased_uniquetopolyAbiased_hist <- ggplot(aRMS_SS_down_polyAunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -1, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS polyA relative to aRMS riboD",
       subtitle = "polyA biased, compared to polyA_unbiased. Median LFC = -3.25",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_down_polyAunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to the riboD biased

```{r aRMS_SS_down_polyAunbiased_uniquetoriboDbiased}
aRMS_SS_down_polyAunbiased_uniquetoriboDbiased <- armsPolyA_ssRiboD_less1 %>%
  anti_join(armsPolyA_ssPolyA_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene")
print(aRMS_SS_down_polyAunbiased_uniquetoriboDbiased %>% nrow())
```

```{r aRMS_SS_down_polyAunbiased_uniquetoriboDbiased_avgLFC}
aRMS_SS_down_polyAunbiased_uniquetoriboDbiased_avgLFC <- aRMS_SS_down_polyAunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_down_polyAunbiased_uniquetoriboDbiased_avgLFC
```

```{r aRMS_SS_down_polyAunbiased_uniquetoriboDbiased_hist}
aRMS_SS_down_polyAunbiased_uniquetoriboDbiased_hist <- ggplot(aRMS_SS_down_polyAunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, compared to polyA_unbiased. Median LFC = -1.38",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_down_polyAunbiased_uniquetoriboDbiased_hist
```

Looking at downregulated genes in the riboD unbiased:polyA biased:riboD biased comparison

Looking at genes that are unique to the polyA biased

```{r aRMS_SS_down_riboDunbiased_uniquetopolyAbiased}
aRMS_SS_down_riboDunbiased_uniquetopolyAbiased <- armsRiboD_ssPolyA_less1 %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsPolyA_ssRiboD_less1, by = "Gene")
print(aRMS_SS_down_riboDunbiased_uniquetopolyAbiased %>% nrow())
```

```{r aRMS_SS_down_riboDunbiased_uniquetopolyAbiased_avgLFC}
aRMS_SS_down_riboDunbiased_uniquetopolyAbiased_avgLFC <- aRMS_SS_down_riboDunbiased_uniquetopolyAbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_down_riboDunbiased_uniquetopolyAbiased_avgLFC
```

```{r aRMS_SS_down_riboDunbiased_uniquetopolyAbiased_hist}
aRMS_SS_down_riboDunbiased_uniquetopolyAbiased_hist <- ggplot(aRMS_SS_down_riboDunbiased_uniquetopolyAbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -1, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS polyA relative to aRMS riboD",
       subtitle = "polyA biased, compared to riboD_truth. Median LFC = -2.85",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_down_riboDunbiased_uniquetopolyAbiased_hist
```

Looking at genes that are unique to the riboD biased

```{r aRMS_SS_down_riboDunbiased_uniquetoriboDbiased}
aRMS_SS_down_riboDunbiased_uniquetoriboDbiased <- armsPolyA_ssRiboD_less1 %>%
  anti_join(armsRiboD_ssRiboD_less1, by = "Gene") %>%
  anti_join(armsRiboD_ssPolyA_less1, by = "Gene")
print(aRMS_SS_down_riboDunbiased_uniquetoriboDbiased %>% nrow())
```

```{r aRMS_SS_down_riboDunbiased_uniquetoriboDbiased_avgLFC}
aRMS_SS_down_riboDunbiased_uniquetoriboDbiased_avgLFC <- aRMS_SS_down_riboDunbiased_uniquetoriboDbiased %>%
  summarise(
    min_value = min(log2FoldChange),
    max_value = max(log2FoldChange),
    median_value = median(log2FoldChange),
    mean_value = mean(log2FoldChange),
    stdev = sd(log2FoldChange)
  )
aRMS_SS_down_riboDunbiased_uniquetoriboDbiased_avgLFC
```

```{r aRMS_SS_down_riboDunbiased_uniquetoriboDbiased_hist}
aRMS_SS_down_riboDunbiased_uniquetoriboDbiased_hist <- ggplot(aRMS_SS_down_riboDunbiased_uniquetoriboDbiased, aes(x = log2FoldChange)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(0, -2, -4, -6, -8, -10, -12, -14, -16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, compared to riboD_truth. Median LFC = -1.47",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_down_riboDunbiased_uniquetoriboDbiased_hist
```

Genes in common between the polyAunbiased:polyAbiased:riboDbiased venn diagram and the riboDunbiased:polyAbiased:riboDbiased venn diagram

upregulated polyA biased

```{r aRMS_SS_up_polyAbiased}
aRMS_SS_up_polyAbiased <- inner_join(aRMS_SS_up_polyAunbiased_uniquetopolyAbiased, aRMS_SS_up_riboDunbiased_uniquetopolyAbiased, by = "Gene")

write_rds(aRMS_SS_up_polyAbiased, "../../output_data/aRMS_SS/aRMS_SS_up_polyAbiased.rds")

print(aRMS_SS_up_polyAbiased %>% nrow())
```

```{r aRMS_SS_up_polyAbiased_avgLFC}
aRMS_SS_up_polyAbiased_avgLFC <- aRMS_SS_up_polyAbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
aRMS_SS_up_polyAbiased_avgLFC
```

```{r aRMS_SS_up_polyAbiased_hist}
aRMS_SS_up_polyAbiased_hist <- ggplot(aRMS_SS_up_polyAbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in SS polyA relative to aRMS riboD",
       subtitle = "polyA biased, Median LFC = 1.32",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

# ggsave(filename = "../../plots/DESeq2_results_analysis/histograms/aRMS_SS_up_polyAbiased_hist.png", plot = aRMS_SS_up_polyAbiased_hist)
# 
# ggsave(filename = "../../plots/DESeq2_results_analysis/histograms/aRMS_SS_up_polyAbiased_hist.svg", plot = aRMS_SS_up_polyAbiased_hist)
# 
# saveRDS(aRMS_SS_up_polyAbiased_hist, file = "../../plots/DESeq2_results_analysis/histograms/aRMS_SS_up_polyAbiased_hist.rds")

aRMS_SS_up_polyAbiased_hist
```

upregulated riboD biased

```{r aRMS_SS_up_riboDbiased}
aRMS_SS_up_riboDbiased <- inner_join(aRMS_SS_up_polyAunbiased_uniquetoriboDbiased, aRMS_SS_up_riboDunbiased_uniquetoriboDbiased, by = "Gene")

write_rds(aRMS_SS_up_riboDbiased, "../../output_data/aRMS_SS/aRMS_SS_up_riboDbiased.rds")

print(aRMS_SS_up_riboDbiased %>% nrow())
```

```{r aRMS_SS_up_riboDbiased_avgLFC}
aRMS_SS_up_riboDbiased_avgLFC <- aRMS_SS_up_riboDbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
aRMS_SS_up_riboDbiased_avgLFC
```

```{r aRMS_SS_up_riboDbiased_hist}
aRMS_SS_up_riboDbiased_hist <- ggplot(aRMS_SS_up_riboDbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely upregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, Median LFC = 2.76",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

# ggsave(filename = "../../plots/DESeq2_results_analysis/histograms/aRMS_SS_up_riboDbiased_hist.png", plot = aRMS_SS_up_riboDbiased_hist)
# 
# ggsave(filename = "../../plots/DESeq2_results_analysis/histograms/aRMS_SS_up_riboDbiased_hist.svg", plot = aRMS_SS_up_riboDbiased_hist)
# 
# saveRDS(aRMS_SS_up_riboDbiased_hist, file = "../../plots/DESeq2_results_analysis/histograms/aRMS_SS_up_riboDbiased_hist.rds")

aRMS_SS_up_riboDbiased_hist
```

Genes in common between the polyAunbiased:polyAbiased:riboDbiased venn diagram and the riboDunbiased:polyAbiased:riboDbiased venn diagram

downregulated polyA biased

```{r aRMS_SS_down_polyAbiased}
aRMS_SS_down_polyAbiased <- inner_join(aRMS_SS_down_polyAunbiased_uniquetopolyAbiased, aRMS_SS_down_riboDunbiased_uniquetopolyAbiased, by = "Gene")

write_rds(aRMS_SS_down_polyAbiased, "../../output_data/aRMS_SS/aRMS_SS_down_polyAbiased.rds")

print(aRMS_SS_down_polyAbiased %>% nrow())
```

```{r aRMS_SS_down_polyAbiased_avgLFC}
aRMS_SS_down_polyAbiased_avgLFC <- aRMS_SS_down_polyAbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
aRMS_SS_down_polyAbiased_avgLFC
```

```{r aRMS_SS_down_polyAbiased_hist}
aRMS_SS_down_polyAbiased_hist <- ggplot(aRMS_SS_down_polyAbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#0072B2",
    color = "#0072B2",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS polyA relative to aRMS riboD",
       subtitle = "polyA biased, Median LFC = -2.90",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_down_polyAbiased_hist
```

downregulated riboD biased

```{r aRMS_SS_down_riboDbiased}
aRMS_SS_down_riboDbiased <- inner_join(aRMS_SS_down_polyAunbiased_uniquetoriboDbiased, aRMS_SS_down_riboDunbiased_uniquetoriboDbiased, by = "Gene")

write_rds(aRMS_SS_down_riboDbiased, "../../output_data/aRMS_SS/aRMS_SS_down_riboDbiased.rds")

print(aRMS_SS_down_riboDbiased %>% nrow())
```

```{r aRMS_SS_down_riboDbiased_avgLFC}
aRMS_SS_down_riboDbiased_avgLFC <- aRMS_SS_down_riboDbiased %>%
  summarise(
    min_value = min(log2FoldChange.x),
    max_value = max(log2FoldChange.x),
    median_value = median(log2FoldChange.x),
    mean_value = mean(log2FoldChange.x),
    stdev = sd(log2FoldChange.x)
  )
aRMS_SS_down_riboDbiased_avgLFC
```

```{r aRMS_SS_down_riboDbiased_hist}
aRMS_SS_down_riboDbiased_hist <- ggplot(aRMS_SS_down_riboDbiased, aes(x = log2FoldChange.x)) +
  geom_histogram(
    bins = 30, # 30 is default
    fill = "#E69F00",
    color = "#E69F00",
    alpha = 0.7
    ) +
#  scale_y_continuous(breaks = c(200, 400, 600, 800), limits = c(0,1000)) +
#  scale_x_continuous(breaks = c(2, 4, 6, 8, 10, 12, 14, 16)) +
  labs(title = "Distribution of log2FoldChange values of \n genes uniquely downregulated in SS riboD relative to aRMS polyA",
       subtitle = "riboD biased, Median LFC = -1.36",
       y = "Number of genes") +
  theme(plot.title = element_text(hjust = 0.5, size = 13, face = "bold"),
            axis.text = element_text(size = 13),
        plot.subtitle = element_text(size = 8, face = "bold"))

aRMS_SS_down_riboDbiased_hist
```

### Session Info

```{r}
sessioninfo::session_info()
```
