---
title: "Fig1E"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## Fig 1E

Boxplot showing expression level of HIST1H1B, a nonpolyadenylated histone subunit gene.

Boxplot showing expression of average median ratio gene, calculated as the median expression level in PolyA samples divided by the median expression level in RiboD samples

```{r load libraries}
library(tidyverse)
```

```{r load in files}
# expression files
SS_aRMS_log2tpm1 <- read_tsv("../../input_data/SS_aRMS_log2TPM1_ensembl_TEBA.tsv.gz")
WT_NB_log2tpm1 <- read_tsv("../../input_data/WT_NB_log2TPM1_ensembl_TEBA.tsv.gz")
ALL_AML_log2tpm1 <- read_tsv("../../input_data/ALL_AML_log2TPM1_ensembl_TEBA.tsv.gz")

# sample ID files, from correlation_analysis
filtered_SS_aRMS_list <- read_tsv("../../input_data/filtered_SS_aRMS_list.tsv")
filtered_WT_NB_list <- read_tsv("../../input_data/filtered_WT_NB_list.tsv")
filtered_ALL_AML_list <- read_tsv("../../input_data/filtered_ALL_AML_list.tsv")

# to convert EnsemblIDs to HugoIDs
gene_names <- read.table("../../input_data/EnsGeneID_Hugo_Observed_Conversions.txt",
header = TRUE, sep = "\t", stringsAsFactors = FALSE )
```

```{r separating samples into disease + prep method}
# SS polyA
SS_polyA_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("SS_polyA")) %>% 
  select(term, disease_and_prep)

SS_polyA_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(SS_polyA_list$term))

# SS riboD
SS_riboD_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("SS_riboD")) %>% 
  select(term, disease_and_prep)

SS_riboD_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(SS_riboD_list$term))

# aRMS polyA
aRMS_polyA_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("aRMS_polyA")) %>% 
  select(term, disease_and_prep)

aRMS_polyA_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(aRMS_polyA_list$term))

# aRMS riboD
aRMS_riboD_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("aRMS_riboD")) %>% 
  select(term, disease_and_prep)

aRMS_riboD_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(aRMS_riboD_list$term))

# WT polyA
WT_polyA_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("WT_polyA")) %>% 
  select(term, disease_and_prep)

WT_polyA_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(WT_polyA_list$term))

# WT riboD 
WT_riboD_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("WT_riboD")) %>% 
  select(term, disease_and_prep)

WT_riboD_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(WT_riboD_list$term))

# NB polyA
NB_polyA_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("NB_polyA")) %>% 
  select(term, disease_and_prep)

NB_polyA_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(NB_polyA_list$term))

# NB riboD
NB_riboD_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("NB_riboD")) %>% 
  select(term, disease_and_prep)

NB_riboD_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(NB_riboD_list$term))

# ALL polyA
ALL_polyA_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("ALL_polyA")) %>% 
  select(term, disease_and_prep)

ALL_polyA_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(ALL_polyA_list$term))

# ALL riboD
ALL_riboD_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("ALL_riboD")) %>% 
  select(term, disease_and_prep)

ALL_riboD_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(ALL_riboD_list$term))


# AML polyA
AML_polyA_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("AML_polyA")) %>% 
  select(term, disease_and_prep)

AML_polyA_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(AML_polyA_list$term))

# AML riboD
AML_riboD_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("AML_riboD")) %>% 
  select(term, disease_and_prep)

AML_riboD_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(AML_riboD_list$term))
```

Convert EnsemblIDs to HugoIDs

```{r}
convert_to_hugo <- function(df) {
  df %>%
    left_join(gene_names, by = c("Gene" = "EnsGeneID")) %>%
    mutate(Gene = ifelse(!is.na(HugoID), HugoID, Gene)) %>%
    select(-HugoID)
}
```

Finding median expression genes

```{r median expression gene - SS}
# Find gene medians in polyA SS
SS_polyA_medians <- SS_polyA_expr %>%
  rowwise() %>%
  mutate(SS_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, SS_polyA_median)

# Find gene medians in riboD SS
SS_riboD_medians <- SS_riboD_expr %>%
  rowwise() %>%
  mutate(SS_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, SS_riboD_median)

# Join and compute ratio (polyA median/riboD median)
SS_gene_medians <- SS_polyA_medians %>%
  inner_join(SS_riboD_medians, by = "Gene") %>%
  mutate(SS_median_ratio = SS_polyA_median / SS_riboD_median)

# Filter out genes with no expression
SS_gene_medians_filtered <- SS_gene_medians %>%
  filter(SS_polyA_median > 0 & SS_riboD_median > 0)

# Recompute median ratio value
SS_median_ratio_val <- median(SS_gene_medians_filtered$SS_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
SS_most_average_gene <- SS_gene_medians_filtered %>%
  mutate(dist = abs(SS_median_ratio - SS_median_ratio_val)) %>%
  arrange(dist) %>%
  slice(1)  

SS_most_average_gene_hugo <- SS_most_average_gene %>%
  convert_to_hugo()
SS_most_average_gene_hugo
```

```{r median expression gene - aRMS}
# Find gene medians in polyA aRMS
aRMS_polyA_medians <- aRMS_polyA_expr %>%
  rowwise() %>%
  mutate(aRMS_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, aRMS_polyA_median)

# Find gene medians in riboD aRMS
aRMS_riboD_medians <- aRMS_riboD_expr %>%
  rowwise() %>%
  mutate(aRMS_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, aRMS_riboD_median)

# Join and compute ratio (polyA median/riboD median)
aRMS_gene_medians <- aRMS_polyA_medians %>%
  inner_join(aRMS_riboD_medians, by = "Gene") %>%
  mutate(aRMS_median_ratio = aRMS_polyA_median / aRMS_riboD_median)

# Filter out genes with no or near-zero expression
aRMS_gene_medians_filtered <- aRMS_gene_medians %>%
  filter(aRMS_polyA_median > 0 & aRMS_riboD_median > 0)

# Recompute median ratio value
aRMS_median_ratio_val <- median(aRMS_gene_medians_filtered$aRMS_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
aRMS_most_average_gene <- aRMS_gene_medians_filtered %>%
  mutate(dist = abs(aRMS_median_ratio - aRMS_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)   # Isolates gene closest to ave median ration

aRMS_most_average_gene_hugo <- aRMS_most_average_gene %>%
  convert_to_hugo()
aRMS_most_average_gene_hugo
```

```{r median expression gene - WT}
# Find gene medians in polyA WT
WT_polyA_medians <- WT_polyA_expr %>%
  rowwise() %>%
  mutate(WT_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, WT_polyA_median)

# Find gene medians in riboD WT
WT_riboD_medians <- WT_riboD_expr %>%
  rowwise() %>%
  mutate(WT_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, WT_riboD_median)

# Join and compute ratio (polyA median/riboD median)
WT_gene_medians <- WT_polyA_medians %>%
  inner_join(WT_riboD_medians, by = "Gene") %>%
  mutate(WT_median_ratio = WT_polyA_median / WT_riboD_median)

# Filter out genes with no or near-zero expression
WT_gene_medians_filtered <- WT_gene_medians %>%
  filter(WT_polyA_median > 0 & WT_riboD_median > 0)

# Recompute median ratio value
WT_median_ratio_val <- median(WT_gene_medians_filtered$WT_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
WT_most_average_gene <- WT_gene_medians_filtered %>%
  mutate(dist = abs(WT_median_ratio - WT_median_ratio_val)) %>%
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

WT_most_average_gene_hugo <- WT_most_average_gene %>%
  convert_to_hugo()
WT_most_average_gene_hugo
```

```{r median expression gene - NB}
# Find gene medians in polyA NB
NB_polyA_medians <- NB_polyA_expr %>%
  rowwise() %>%
  mutate(NB_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, NB_polyA_median)

# Find gene medians in riboD NB
NB_riboD_medians <- NB_riboD_expr %>%
  rowwise() %>%
  mutate(NB_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, NB_riboD_median)

# Join and compute ratio (polyA median/riboD median)
NB_gene_medians <- NB_polyA_medians %>%
  inner_join(NB_riboD_medians, by = "Gene") %>%
  mutate(NB_median_ratio = NB_polyA_median / NB_riboD_median)

# Filter out genes with no or near-zero expression
NB_gene_medians_filtered <- NB_gene_medians %>%
  filter(NB_polyA_median > 0 & NB_riboD_median > 0)

# Recompute median ratio value
NB_median_ratio_val <- median(NB_gene_medians_filtered$NB_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
NB_most_average_gene <- NB_gene_medians_filtered %>%
  mutate(dist = abs(NB_median_ratio - NB_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

NB_most_average_gene_hugo <- NB_most_average_gene %>%
  convert_to_hugo()
NB_most_average_gene_hugo
```

```{r median expression gene - ALL}
# Find gene medians in polyA ALL
ALL_polyA_medians <- ALL_polyA_expr %>%
  rowwise() %>%
  mutate(ALL_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, ALL_polyA_median)

# Find gene medians in riboD ALL
ALL_riboD_medians <- ALL_riboD_expr %>%
  rowwise() %>%
  mutate(ALL_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, ALL_riboD_median)

# Join and compute ratio (polyA median/riboD median)
ALL_gene_medians <- ALL_polyA_medians %>%
  inner_join(ALL_riboD_medians, by = "Gene") %>%
  mutate(ALL_median_ratio = ALL_polyA_median / ALL_riboD_median)

# Filter out genes with no or near-zero expression
ALL_gene_medians_filtered <- ALL_gene_medians %>%
  filter(ALL_polyA_median > 0 & ALL_riboD_median > 0)

# Recompute median ratio value
ALL_median_ratio_val <- median(ALL_gene_medians_filtered$ALL_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
ALL_most_average_gene <- ALL_gene_medians_filtered %>%
  mutate(dist = abs(ALL_median_ratio - ALL_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

ALL_most_average_gene_hugo <- ALL_most_average_gene %>%
  convert_to_hugo()
ALL_most_average_gene_hugo
```

```{r median expression gene - AML}
# Find gene medians in polyA AML
AML_polyA_medians <- AML_polyA_expr %>%
  rowwise() %>%
  mutate(AML_polyA_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, AML_polyA_median)

# Find gene medians in riboD AML
AML_riboD_medians <- AML_riboD_expr %>%
  rowwise() %>%
  mutate(AML_riboD_median = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
  ungroup() %>%
  select(Gene, AML_riboD_median)

# Join and compute ratio (polyA median/riboD median)
AML_gene_medians <- AML_polyA_medians %>%
  inner_join(AML_riboD_medians, by = "Gene") %>%
  mutate(AML_median_ratio = AML_polyA_median / AML_riboD_median)

# Filter out genes with no or near-zero expression
AML_gene_medians_filtered <- AML_gene_medians %>%
  filter(AML_polyA_median > 1 & AML_riboD_median > 0.9)

# Recompute median ratio value
AML_median_ratio_val <- median(AML_gene_medians_filtered$AML_median_ratio, na.rm = TRUE)

# Find gene closest to that median ratio
AML_most_average_gene <- AML_gene_medians_filtered %>%
  mutate(dist = abs(AML_median_ratio - AML_median_ratio_val)) %>%  # computes how far each gene’s ratio is from overall median
  arrange(dist) %>%
  slice(1)         # Isolates gene closest to ave median ration

AML_most_average_gene_hugo <- AML_most_average_gene %>%
  convert_to_hugo()
AML_most_average_gene_hugo
```

Converting expression to hugoIDs

```{r convert to HugoIDs}
# SS
SS_polyA_expression_hugo <- SS_polyA_expr %>%
  convert_to_hugo()

SS_riboD_expression_hugo <- SS_riboD_expr %>%
  convert_to_hugo()

SS_gene_medians_hugo <- SS_gene_medians %>%
  convert_to_hugo()

# aRMS
aRMS_polyA_expression_hugo <- aRMS_polyA_expr %>%
  convert_to_hugo() 

aRMS_riboD_expression_hugo <- aRMS_riboD_expr %>%
  convert_to_hugo()

aRMS_gene_median_hugo <- aRMS_gene_medians %>%
  convert_to_hugo()

# WT
WT_polyA_expression_hugo <- WT_polyA_expr %>%
  convert_to_hugo()

WT_riboD_expression_hugo <- WT_riboD_expr %>%
  convert_to_hugo()

WT_gene_medians_hugo <- WT_gene_medians %>%
  convert_to_hugo()

# NB
NB_polyA_expression_hugo <- NB_polyA_expr %>%
  convert_to_hugo()

NB_riboD_expression_hugo <- NB_riboD_expr %>%
  convert_to_hugo()

NB_gene_medians_hugo <- NB_gene_medians %>%
  convert_to_hugo()

# ALL
ALL_polyA_expression_hugo <- ALL_polyA_expr %>%
  convert_to_hugo()

ALL_riboD_expression_hugo <- ALL_riboD_expr %>%
  convert_to_hugo()

ALL_gene_medians_hugo <- ALL_gene_medians %>%
  convert_to_hugo()

# AML
AML_polyA_expression_hugo <- AML_polyA_expr %>%
  convert_to_hugo()

AML_riboD_expression_hugo <- AML_riboD_expr %>%
  convert_to_hugo()

AML_gene_medians_hugo <- AML_gene_medians %>%
  convert_to_hugo()
```

Filtering for HIST1H1B, a non-polyadenylated gene

```{r HIST1H1B filter}
H1B_SS_polyA_expression <- SS_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_SS_riboD_expression <- SS_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_aRMS_polyA_expression <- aRMS_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_aRMS_riboD_expression <- aRMS_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_WT_polyA_expression <- WT_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_WT_riboD_expression <- WT_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_NB_polyA_expression <- NB_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_NB_riboD_expression <- NB_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_ALL_polyA_expression <- ALL_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_ALL_riboD_expression <- ALL_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 

H1B_AML_polyA_expression <- AML_polyA_expression_hugo %>%
   filter(Gene == "HIST1H1B")

H1B_AML_riboD_expression <- AML_riboD_expression_hugo %>%
   filter(Gene == "HIST1H1B") 
```

Filtering for most averagely expressed gene

```{r avg expression gene filter}
#SS Most Average Expressed gene
avg_SS_polyA_expression <- SS_polyA_expression_hugo %>%
   filter(Gene == SS_most_average_gene_hugo$Gene) 

avg_SS_riboD_expression <- SS_riboD_expression_hugo %>%
   filter(Gene == SS_most_average_gene_hugo$Gene)

#aRMS Most Average Expressed gene
avg_aRMS_riboD_expression <- aRMS_riboD_expression_hugo %>%
   filter(Gene == aRMS_most_average_gene_hugo$Gene)

avg_aRMS_polyA_expression <- aRMS_polyA_expression_hugo %>%
   filter(Gene == aRMS_most_average_gene_hugo$Gene)

#WT Most Average Expressed gene 
avg_WT_riboD_expression <- WT_riboD_expression_hugo %>%
   filter(Gene == WT_most_average_gene_hugo$Gene)

avg_WT_polyA_expression <- WT_polyA_expression_hugo %>%
   filter(Gene == WT_most_average_gene_hugo$Gene)

#NB Most Average Expressed gene
avg_NB_riboD_expression <- NB_riboD_expression_hugo %>%
   filter(Gene == NB_most_average_gene_hugo$Gene)

avg_NB_polyA_expression <- NB_polyA_expression_hugo %>%
   filter(Gene == NB_most_average_gene_hugo$Gene)

#ALL Most Average Expressed gene
avg_ALL_riboD_expression <- ALL_riboD_expression_hugo %>%
   filter(Gene == ALL_most_average_gene_hugo$Gene)

avg_ALL_polyA_expression <- ALL_polyA_expression_hugo %>%
   filter(Gene == ALL_most_average_gene_hugo$Gene)

#AML Most Average Expressed gene
avg_AML_riboD_expression <- AML_riboD_expression_hugo %>%
   filter(Gene == AML_most_average_gene_hugo$Gene)

avg_AML_polyA_expression <- AML_polyA_expression_hugo %>%
   filter(Gene == AML_most_average_gene_hugo$Gene)
```

Reformating HIST1H1B data

```{r reformat HIST1H1B}
H1B_SS_polyA_long <- H1B_SS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "polyA", Gene = "HIST1H1B")

H1B_SS_riboD_long <- H1B_SS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "riboD", Gene = "HIST1H1B")

H1B_aRMS_polyA_long <- H1B_aRMS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "polyA", Gene = "HIST1H1B")

H1B_aRMS_riboD_long <- H1B_aRMS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "riboD", Gene = "HIST1H1B")

H1B_WT_polyA_long <- H1B_WT_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "polyA", Gene = "HIST1H1B")

H1B_WT_riboD_long <- H1B_WT_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "riboD", Gene = "HIST1H1B")

H1B_NB_polyA_long <- H1B_NB_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "polyA", Gene = "HIST1H1B")

H1B_NB_riboD_long <- H1B_NB_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "riboD", Gene = "HIST1H1B")

H1B_ALL_polyA_long <- H1B_ALL_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "polyA", Gene = "HIST1H1B")

H1B_ALL_riboD_long <- H1B_ALL_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "riboD", Gene = "HIST1H1B")

H1B_AML_polyA_long <- H1B_AML_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "polyA", Gene = "HIST1H1B")

H1B_AML_riboD_long <- H1B_AML_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "riboD", Gene = "HIST1H1B")
```

Reformatting avg gene

```{r reformat avg gene}
# SS polyA 
avg_SS_polyA_long <- avg_SS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "polyA", Gene = Gene)

# SS riboD
avg_SS_riboD_long <- avg_SS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "SS", Compendia = "riboD", Gene = Gene)

# aRMS polyA
avg_aRMS_polyA_long <- avg_aRMS_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "polyA", Gene = Gene)

# aRMS riboD
avg_aRMS_riboD_long <- avg_aRMS_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "aRMS", Compendia = "riboD", Gene = Gene)

#NB polyA
avg_NB_polyA_long <- avg_NB_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "polyA", Gene = Gene)

#NB riboD
avg_NB_riboD_long <- avg_NB_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "NB", Compendia = "riboD", Gene = Gene)

#WT polyA
avg_WT_polyA_long <- avg_WT_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "polyA", Gene = Gene)

#WT riboD
avg_WT_riboD_long <- avg_WT_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "WT", Compendia = "riboD", Gene = Gene)

#AML polyA
avg_AML_polyA_long <- avg_AML_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "polyA", Gene = Gene)

#AML riboD
avg_AML_riboD_long <- avg_AML_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "AML", Compendia = "riboD", Gene = Gene)

#ALL polyA
avg_ALL_polyA_long <- avg_ALL_polyA_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "polyA", Gene = Gene)

#ALL riboD
avg_ALL_riboD_long <- avg_ALL_riboD_expression %>%
  pivot_longer(-Gene, names_to = "TH_ID", values_to = "Expression") %>%
  mutate(Disease = "ALL", Compendia = "riboD", Gene = Gene)
```

Combining dataframes

```{r combining}
combined_long <- bind_rows(

  # -------------------------
  # HIST1H1B — SS, aRMS, WT, NB, ALL, AML
  # -------------------------
  H1B_SS_polyA_long %>%
    mutate(Disease = "SS"),
  H1B_SS_riboD_long %>%
    mutate(Disease = "SS"),
  
  H1B_aRMS_polyA_long %>%
    mutate(Disease = "aRMS"),
  H1B_aRMS_riboD_long %>%
    mutate(Disease = "aRMS"),
  
  H1B_WT_polyA_long %>%
    mutate(Disease = "WT"),
  H1B_WT_riboD_long %>%
    mutate(Disease = "WT"),
  
  H1B_NB_polyA_long %>%
    mutate(Disease = "NB"),
  H1B_NB_riboD_long %>%
    mutate(Disease = "NB"),
  
  H1B_ALL_polyA_long %>%
    mutate(Disease = "ALL"),
  H1B_ALL_riboD_long %>%
    mutate(Disease = "ALL"),
  
  H1B_AML_polyA_long %>%
    mutate(Disease = "AML"),
  H1B_AML_riboD_long %>%
    mutate(Disease = "AML"),
   # -------------------------
  # Median-ratio genes (SS, aRMS, WT, NB, ALL, AML)
  # -------------------------
  avg_SS_polyA_long %>%
    mutate(Disease = "SS"),
  avg_SS_riboD_long %>%
    mutate(Disease = "SS"),

  avg_aRMS_polyA_long %>%
    mutate(Disease = "aRMS"),
  avg_aRMS_riboD_long %>%
    mutate(Disease = "aRMS"),

  avg_WT_polyA_long %>%
    mutate(Disease = "WT"),
  avg_WT_riboD_long %>%
    mutate(Disease = "WT"),

  avg_NB_polyA_long %>%
    mutate(Disease = "NB"),
  avg_NB_riboD_long %>%
    mutate(Disease = "NB"),
  
  avg_ALL_polyA_long %>%
    mutate(Disease = "ALL"),
  avg_ALL_riboD_long %>%
    mutate(Disease = "ALL"),
  
  avg_AML_polyA_long %>%
    mutate(Disease = "AML"),
  avg_AML_riboD_long %>%
    mutate(Disease = "AML"),
)
```

Custom color theme

```{r color theme}
color_theme <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "top",
      legend.title = element_blank(),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.4),
      axis.ticks = element_line(color = "black", linewidth = 0.4),
      strip.text = element_text(face = "bold", size = base_size * 0.9),
      plot.title = element_text(face = "bold", size = base_size * 1.1, hjust = 0.5),
      axis.title.y = element_text(angle = 90, hjust = 0.5, size = 16), 
      axis.title.x = element_text(angle = 0, hjust = 0.5, size = 17), 
      axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
      plot.margin = margin(10, 10, 10, 10)
    )
}

scale_fill_compendia <- function() {
  scale_fill_manual(values = c(
    "riboD" = "#E69F00",  # yellow
    "polyA" = "#0072B2"   # blue
  ))
}
scale_color_compendia <- function() {
  scale_color_manual(values = c( 
    "riboD" = "#E69F00", 
    "polyA" = "#0072B2" ))
  }
```

Merged plot

```{r}
# Map of median-ratio genes
median_gene_map <- list(
  SS      = SS_most_average_gene_hugo$Gene,
  aRMS    = aRMS_most_average_gene_hugo$Gene,
  WT      = WT_most_average_gene_hugo$Gene,
  NB      = NB_most_average_gene_hugo$Gene,
  ALL     = ALL_most_average_gene_hugo$Gene,
  AML     = AML_most_average_gene_hugo$Gene
)

diseases <- unique(combined_long$Disease)

```

Generate merged plot for each disease

```{r generate merged plot}
merged_plots <- map(diseases, function(d) {

  median_gene <- median_gene_map[[d]]

  # HIST1H1B expression
  df_h1b <- combined_long %>%
    filter(Disease == d, Gene == "HIST1H1B") %>%
    mutate(PlotType = "HIST1H1B (PolyA-)")

  # Median-ratio gene expression
  df_median <- combined_long %>%
    filter(Disease == d, Gene == median_gene) %>%
    mutate(PlotType = "Median Ratio Gene (PolyA+)")

  # Combine
  df_combined <- bind_rows(df_h1b, df_median)

  ggplot(df_combined, aes(x = Gene, y = Expression, fill = Compendia)) +
    geom_boxplot(
      alpha = 0.7,
      position = position_dodge(width = 0.8),
      outlier.shape = NA,
      linewidth = 0.4
    ) +
    geom_jitter(
      aes(color = Compendia),
      position = position_jitterdodge(jitter.width = 0.2,
                                      dodge.width = 0.8),
      size = 1.2,
      alpha = 0.6
    ) +
    scale_fill_compendia() +
    scale_color_compendia() +
    coord_cartesian(ylim = c(0, 17)) +
    facet_wrap(~ PlotType, ncol = 2, scales = "free_x") +
    labs(
      title = paste("Expression of Representative PolyA+ and PolyA- Gene in", d),
      x = "Gene",
      y = "Expression log2(TPM+1)"
    ) +
    color_theme()
})

names(merged_plots) <- diseases
merged_plots

```
