---
title: "aRMS_SS_DESeq2"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## aRMS v SS

This notebook will be analyzing the expression profiles of synovial sarcoma (SS) and alveolar rhabdomyosarcoma (aRMS) tumors from the v25 riboD and polyA compendia.

The goal is to determine the extent that genes get called as differentially expressed as a result of the transcript enrichment method of the library preparation process.

The comparisons we’ll be making (SS = synovial sarcoma, aRMS = alveolar rhabdomyosarcoma):

- aRMS polyA vs SS polyA

- aRMS riboD vs SS riboD

- aRMS polyA vs SS riboD

- aRMS riboD vs SS polyA

```{r install packages}
library(tidyverse)
library(DESeq2)
```

```{r import data}
# sample metadata
# This includes the top 15 samples from SS polyA, SS riboD, aRMS polyA, aRMS riboD
# The top 15 samples came from the correlation analysis, where this tsv file was generated
filtered_SS_aRMS_list <- read_tsv("../../input_data/filtered_SS_aRMS_list.tsv", show_col_types = FALSE)

expression_SS_aRMS <- read_tsv("../../input_data/expression_SS_aRMS.tsv.gz", show_col_types = FALSE)

gene_names <- read_tsv("../../input_data/EnsGeneID_Hugo_Observed_Conversions.txt", show_col_types = FALSE)

```

That expression file contains all expression information for all SS and aRMS samples across the compendia (170 samples). We want only the expression information for the samples we’re interested in, based on the correlation analysis we did previously (60 samples).

```{r filter expr data}
# selecting only the 15 most correlated samples for each category
interested_expression <- expression_SS_aRMS %>%
  select(all_of(c("Gene", filtered_SS_aRMS_list$term)))

```

Since filtering out genes with a count less than 10 across all samples will filter out relevant genes, we’ll filter out genes with a count less than 10 in 11 or more samples (i.e., we’re keeping genes that have a count greater than 10 in at least 10 samples).

For example, if Gene A is expressed in SS and not in aRMS, but it’s only picked up by polyA-selection, then only SS-polyA would have expression for it, and the gene would be filtered out in the pre-group filtering step below. However, we would want to capture that gene information, because it is both biologically differentially expressed (high expression in SS, low expression in aRMS) and represents an artifact of library prep (high expression in SS-polyA, low expression in SS-riboD).

Thus, we filter for genes with expression above 10 in at least 10 samples, which would capture the expression of Gene A in the example above.

```{r filtering out low genes}
# Explanation:
# across(where(is.numeric)) is selecting all the numeric columns
# i.e., all the sample columns. This deselects the Gene column, since that is a character column

# rowSums() counts how many samples meet the threshold for each gene,
# i.e., have a value of at least 10

# then keeps only the genes that meet that threshold in at least 10 samples
updated_expression <- interested_expression %>%
  filter(rowSums(across(where(is.numeric)) >= 10) >= 10)
```

### aRMS polyA vs SS polyA

```{r}
# metadata
aRMS_polyA_SS_polyA_meta <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep == "aRMS_polyA" | disease_and_prep == "SS_polyA")
# the pipe | means we're taking all samples called SS_riboD OR SS_polyA in column disease_and_prep

# expression
aRMS_polyA_SS_polyA_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", aRMS_polyA_SS_polyA_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_aRMS_polyA_SS_polyA_updated_expr <- aRMS_polyA_SS_polyA_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_aRMS_polyA_SS_polyA_meta <- aRMS_polyA_SS_polyA_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_aRMS_polyA_SS_polyA_meta) %in% colnames(fix_aRMS_polyA_SS_polyA_updated_expr))

# check in order is the same
all(rownames(fix_aRMS_polyA_SS_polyA_meta) == colnames(fix_aRMS_polyA_SS_polyA_updated_expr))
```

```{r}
# format data
round_aRMS_polyA_SS_polyA_updated_expr = round(fix_aRMS_polyA_SS_polyA_updated_expr)

aRMS_polyA_SS_polyA_updated_expr_matrix <- as.matrix(round_aRMS_polyA_SS_polyA_updated_expr)

fix_aRMS_polyA_SS_polyA_meta$disease_and_prep <- as.factor(fix_aRMS_polyA_SS_polyA_meta$disease_and_prep)
```

```{r}
# set reference levels
# we want to make sure the references are consistent across the comparisons
levels(fix_aRMS_polyA_SS_polyA_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_aRMS_polyA_SS_polyA_updated <- DESeqDataSetFromMatrix(countData = aRMS_polyA_SS_polyA_updated_expr_matrix,
                              colData = fix_aRMS_polyA_SS_polyA_meta,
                              design = ~disease_and_prep)
```

```{r}
dds_aRMS_polyA_SS_polyA_updated
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
aRMS_polyA_SS_polyA_updated_genes_to_keep <- rowSums(counts(dds_aRMS_polyA_SS_polyA_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_aRMS_polyA_SS_polyA_updated <- dds_aRMS_polyA_SS_polyA_updated[which(aRMS_polyA_SS_polyA_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_aRMS_polyA_SS_polyA_updated_object <- DESeq(dds_aRMS_polyA_SS_polyA_updated)
```

```{r}
# create results table
result_dds_aRMS_polyA_SS_polyA_updated <- results(dds_aRMS_polyA_SS_polyA_updated_object)
# summarize results table
summary(result_dds_aRMS_polyA_SS_polyA_updated, alpha = 0.05)
```

```{r}
# convert results to dataframe
result_dds_aRMS_polyA_SS_polyA_updated_df <- result_dds_aRMS_polyA_SS_polyA_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
aRMS_polyA_SS_polyA_LFC1great_padj <- result_dds_aRMS_polyA_SS_polyA_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
aRMS_polyA_SS_polyA_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
aRMS_polyA_SS_polyA_LFC1less_padj <- result_dds_aRMS_polyA_SS_polyA_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
aRMS_polyA_SS_polyA_LFC1less_padj
```

```{r}
# convert EnsemblID to HugoID for gene name readability
hugo_results_aRMS_polyA_SS_polyA_updated <- result_dds_aRMS_polyA_SS_polyA_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_aRMS_polyA_SS_polyA_updated, path = "../../output_data/aRMS_SS/hugo_results_aRMS_polyA_SS_polyA_updated.tsv.gz")
```

### aRMS riboD vs SS riboD

```{r}
# metadata
aRMS_riboD_SS_riboD_meta <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep == "aRMS_riboD" | disease_and_prep == "SS_riboD")

# expression
aRMS_riboD_SS_riboD_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", aRMS_riboD_SS_riboD_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_aRMS_riboD_SS_riboD_updated_expr <- aRMS_riboD_SS_riboD_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_aRMS_riboD_SS_riboD_meta <- aRMS_riboD_SS_riboD_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_aRMS_riboD_SS_riboD_meta) %in% colnames(fix_aRMS_riboD_SS_riboD_updated_expr))

# check in order is the same
all(rownames(fix_aRMS_riboD_SS_riboD_meta) == colnames(fix_aRMS_riboD_SS_riboD_updated_expr))
```

```{r}
# format data
round_aRMS_riboD_SS_riboD_updated_expr = round(fix_aRMS_riboD_SS_riboD_updated_expr)

aRMS_riboD_SS_riboD_updated_expr_matrix <- as.matrix(round_aRMS_riboD_SS_riboD_updated_expr)

fix_aRMS_riboD_SS_riboD_meta$disease_and_prep <- as.factor(fix_aRMS_riboD_SS_riboD_meta$disease_and_prep)
```

```{r}
# set reference levels
# we want to make sure the references are consistent across the comparisons
levels(fix_aRMS_riboD_SS_riboD_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_aRMS_riboD_SS_riboD_updated <- DESeqDataSetFromMatrix(countData = aRMS_riboD_SS_riboD_updated_expr_matrix,
                              colData = fix_aRMS_riboD_SS_riboD_meta,
                              design = ~disease_and_prep)
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
aRMS_riboD_SS_riboD_updated_genes_to_keep <- rowSums(counts(dds_aRMS_riboD_SS_riboD_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_aRMS_riboD_SS_riboD_updated <- dds_aRMS_riboD_SS_riboD_updated[which(aRMS_riboD_SS_riboD_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_aRMS_riboD_SS_riboD_updated_object <- DESeq(dds_aRMS_riboD_SS_riboD_updated)
```

```{r}
# create results table
res_aRMS_riboD_SS_riboD_updated <- results(dds_aRMS_riboD_SS_riboD_updated_object)

# summarize results table
summary(res_aRMS_riboD_SS_riboD_updated, alpha = 0.05)
```

```{r}
# first we convert to df
res_aRMS_riboD_SS_riboD_updated_df <- res_aRMS_riboD_SS_riboD_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
aRMS_riboD_SS_riboD_LFC1great_padj <- res_aRMS_riboD_SS_riboD_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
aRMS_riboD_SS_riboD_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
aRMS_riboD_SS_riboD_LFC1less_padj <- res_aRMS_riboD_SS_riboD_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
aRMS_riboD_SS_riboD_LFC1less_padj
```

```{r}
hugo_results_aRMS_riboD_SS_riboD_updated <- res_aRMS_riboD_SS_riboD_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_aRMS_riboD_SS_riboD_updated, path = "../../output_data/aRMS_SS/hugo_results_aRMS_riboD_SS_riboD_updated.tsv.gz")
```

### **aRMS polyA vs SS riboD**

```{r}
# metadata
aRMS_polyA_SS_riboD_meta <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep == "aRMS_polyA" | disease_and_prep == "SS_riboD")
# the pipe | means we're taking all samples called SS_riboD OR SS_polyA in column disease_and_prep

# expression
aRMS_polyA_SS_riboD_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", aRMS_polyA_SS_riboD_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_aRMS_polyA_SS_riboD_updated_expr <- aRMS_polyA_SS_riboD_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_aRMS_polyA_SS_riboD_meta <- aRMS_polyA_SS_riboD_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_aRMS_polyA_SS_riboD_meta) %in% colnames(fix_aRMS_polyA_SS_riboD_updated_expr))

# check in order is the same
all(rownames(fix_aRMS_polyA_SS_riboD_meta) == colnames(fix_aRMS_polyA_SS_riboD_updated_expr))
```

```{r}
# format data
round_aRMS_polyA_SS_riboD_updated_expr = round(fix_aRMS_polyA_SS_riboD_updated_expr)

aRMS_polyA_SS_riboD_updated_expr_matrix <- as.matrix(round_aRMS_polyA_SS_riboD_updated_expr)

fix_aRMS_polyA_SS_riboD_meta$disease_and_prep <- as.factor(fix_aRMS_polyA_SS_riboD_meta$disease_and_prep)
```

```{r}
# check reference levels
levels(fix_aRMS_polyA_SS_riboD_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_aRMS_polyA_SS_riboD_updated <- DESeqDataSetFromMatrix(countData = aRMS_polyA_SS_riboD_updated_expr_matrix,
                              colData = fix_aRMS_polyA_SS_riboD_meta,
                              design = ~disease_and_prep)
dds_aRMS_polyA_SS_riboD_updated
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
aRMS_polyA_SS_riboD_updated_genes_to_keep <- rowSums(counts(dds_aRMS_polyA_SS_riboD_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_aRMS_polyA_SS_riboD_updated <- dds_aRMS_polyA_SS_riboD_updated[which(aRMS_polyA_SS_riboD_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_aRMS_polyA_SS_riboD_updated_object <- DESeq(dds_aRMS_polyA_SS_riboD_updated)
```

```{r}
# create results table
result_dds_aRMS_polyA_SS_riboD_updated <- results(dds_aRMS_polyA_SS_riboD_updated_object)
# summarize results table
summary(result_dds_aRMS_polyA_SS_riboD_updated, alpha = 0.05)
```

```{r}
# convert results to dataframe
result_dds_aRMS_polyA_SS_riboD_updated_df <- result_dds_aRMS_polyA_SS_riboD_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
aRMS_polyA_SS_riboD_LFC1great_padj <- result_dds_aRMS_polyA_SS_riboD_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
aRMS_polyA_SS_riboD_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
aRMS_polyA_SS_riboD_LFC1less_padj <- result_dds_aRMS_polyA_SS_riboD_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
aRMS_polyA_SS_riboD_LFC1less_padj
```

```{r}
hugo_results_aRMS_polyA_SS_riboD_updated <- result_dds_aRMS_polyA_SS_riboD_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_aRMS_polyA_SS_riboD_updated, path = "../../output_data/aRMS_SS/hugo_results_aRMS_polyA_SS_riboD_updated.tsv.gz")
```

### **aRMS riboD vs SS polyA**

```{r}
# metadata
aRMS_riboD_SS_polyA_meta <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep == "aRMS_riboD" | disease_and_prep == "SS_polyA")

# expression
aRMS_riboD_SS_polyA_updated_expr <- updated_expression %>%
  select(all_of(c("Gene", aRMS_riboD_SS_polyA_meta$term)))
```

```{r}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_aRMS_riboD_SS_polyA_updated_expr <- aRMS_riboD_SS_polyA_updated_expr %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names for clin_merged_STS
fix_aRMS_riboD_SS_polyA_meta <- aRMS_riboD_SS_polyA_meta %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

```{r}
# check if row names are same as column names
all(rownames(fix_aRMS_riboD_SS_polyA_meta) %in% colnames(fix_aRMS_riboD_SS_polyA_updated_expr))
# check in order is the same
all(rownames(fix_aRMS_riboD_SS_polyA_meta) == colnames(fix_aRMS_riboD_SS_polyA_updated_expr))
```

```{r}
# format data
round_aRMS_riboD_SS_polyA_updated_expr = round(fix_aRMS_riboD_SS_polyA_updated_expr)

aRMS_riboD_SS_polyA_updated_expr_matrix <- as.matrix(round_aRMS_riboD_SS_polyA_updated_expr)

fix_aRMS_riboD_SS_polyA_meta$disease_and_prep <- as.factor(fix_aRMS_riboD_SS_polyA_meta$disease_and_prep)
```

```{r}
# check reference levels
levels(fix_aRMS_riboD_SS_polyA_meta$disease_and_prep)
```

```{r}
# set up DESeq function
dds_aRMS_riboD_SS_polyA_updated <- DESeqDataSetFromMatrix(countData = aRMS_riboD_SS_polyA_updated_expr_matrix,
                              colData = fix_aRMS_riboD_SS_polyA_meta,
                              design = ~disease_and_prep)
```

```{r}
# create a vector of TRUE and FALSE values where
# TRUE corresponds to genes with counts of at least 10 
aRMS_riboD_SS_polyA_updated_genes_to_keep <- rowSums(counts(dds_aRMS_riboD_SS_polyA_updated)) >= 10

# use which() to prevent any NAs sneaking through
dds_aRMS_riboD_SS_polyA_updated <- dds_aRMS_riboD_SS_polyA_updated[which(aRMS_riboD_SS_polyA_updated_genes_to_keep), ]
```

```{r}
# Run DESeq2
dds_aRMS_riboD_SS_polyA_updated_object <- DESeq(dds_aRMS_riboD_SS_polyA_updated)
```

```{r}
# create results table
res_aRMS_riboD_SS_polyA_updated <- results(dds_aRMS_riboD_SS_polyA_updated_object)

# summarize results table
summary(res_aRMS_riboD_SS_polyA_updated, alpha = 0.05)
```

```{r}
# first we convert to df
res_aRMS_riboD_SS_polyA_updated_df <- res_aRMS_riboD_SS_polyA_updated %>%
  #convert to a data frame
  as.data.frame() %>%
  #the gene ids were stored as row names -- let's them a column
  tibble::rownames_to_column(var = "Gene")
```

```{r}
# how many genes have a Log2 fold change greater or equal to 1 and have an adjusted p-value less than 0.05?
aRMS_riboD_SS_polyA_LFC1great_padj <- res_aRMS_riboD_SS_polyA_updated_df %>%
  filter(log2FoldChange >= 1 & padj < 0.05) %>%
  nrow()
aRMS_riboD_SS_polyA_LFC1great_padj
```

```{r}
# how many genes have a Log2 fold change less than or equal to -1 and have an adjusted p-value less than 0.05?
aRMS_riboD_SS_polyA_LFC1less_padj <- res_aRMS_riboD_SS_polyA_updated_df %>%
  filter(log2FoldChange <= -1 & padj < 0.05) %>%
  nrow()
aRMS_riboD_SS_polyA_LFC1less_padj
```

```{r}
hugo_results_aRMS_riboD_SS_polyA_updated <- res_aRMS_riboD_SS_polyA_updated_df %>%
  left_join(gene_names, by=c("Gene"="EnsGeneID")) %>%
  relocate(HugoID) # moving HugoID column to the front

write_tsv(hugo_results_aRMS_riboD_SS_polyA_updated, path = "../../output_data/aRMS_SS/hugo_results_aRMS_riboD_SS_polyA_updated.tsv.gz")
```

### Session Info

```{r}
sessioninfo::session_info()
```
