---
title: "Fig1G"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## Fig 1G

Scatterplot showing median expression per gene.

```{r load libraries}
library(tidyverse)
library(ggrepel)
library(patchwork)
```

```{r load in files}
# sample ID files
# polyA
SS_polyA_list <- read_tsv("../../input_data/sample_selection/SS_polyA_list.tsv")
aRMS_polyA_list <- read_tsv("../../input_data/sample_selection/aRMS_polyA_list.tsv")
WT_polyA_list <- read_tsv("../../input_data/sample_selection/WT_polyA_list.tsv")
NB_polyA_list <- read_tsv("../../input_data/sample_selection/NB_polyA_list.tsv")
ALL_polyA_list <- read_tsv("../../input_data/sample_selection/ALL_polyA_list.tsv")
AML_polyA_list <- read_tsv("../../input_data/sample_selection/AML_polyA_list.tsv")

# riboD
SS_riboD_list <- read_tsv("../../input_data/sample_selection/SS_riboD_list.tsv")
aRMS_riboD_list <- read_tsv("../../input_data/sample_selection/aRMS_riboD_list.tsv")
WT_riboD_list <- read_tsv("../../input_data/sample_selection/WT_riboD_list.tsv")
NB_riboD_list <- read_tsv("../../input_data/sample_selection/NB_riboD_list.tsv")
ALL_riboD_list <- read_tsv("../../input_data/sample_selection/ALL_riboD_list.tsv")
AML_riboD_list <- read_tsv("../../input_data/sample_selection/AML_riboD_list.tsv")

# log2(TPM+1) expression files
SS_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/SS_polyA_log2tpm1.tsv")
SS_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/SS_riboD_log2tpm1.tsv")

aRMS_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/aRMS_polyA_log2tpm1.tsv")
aRMS_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/aRMS_riboD_log2tpm1.tsv")

WT_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/WT_polyA_log2tpm1.tsv")
WT_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/WT_riboD_log2tpm1.tsv")

NB_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/NB_polyA_log2tpm1.tsv")
NB_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/NB_riboD_log2tpm1.tsv")

ALL_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/ALL_polyA_log2tpm1.tsv")
ALL_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/ALL_riboD_log2tpm1.tsv")

AML_polyA_log2tpm1 <- read_tsv("../../input_data/sample_selection/AML_polyA_log2tpm1.tsv")
AML_riboD_log2tpm1 <- read_tsv("../../input_data/sample_selection/AML_riboD_log2tpm1.tsv")

# to convert EnsemblIDs to HugoIDs
gene_names <- read.table("../../input_data/EnsGeneID_Hugo_Observed_Conversions.txt",
header = TRUE, sep = "\t", stringsAsFactors = FALSE )
```

Compute gene medians

```{r function for gene medians}
compute_gene_medians <- function(polyA_df, riboD_df, disease_name) {
  polyA_Medians <- polyA_df %>%
    rowwise() %>%
    mutate(polyA_medians = median(c_across(-Gene), na.rm = TRUE)) %>%
    ungroup() %>%
    select(Gene, polyA_medians)

  riboD_Medians <- riboD_df %>%
    rowwise() %>%
    mutate(riboD_medians = median(c_across(-Gene), na.rm = TRUE)) %>%
    ungroup() %>%
    select(Gene, riboD_medians)

  merged <- polyA_Medians %>%
    inner_join(riboD_Medians, by = "Gene") %>%
    mutate(Disease = disease_name)

  return(merged)
}
```

Applying function to each disease/library prep type

```{r applying gene medians function}
aRMS_medians <- compute_gene_medians(aRMS_polyA_log2tpm1, aRMS_riboD_log2tpm1, "aRMS")
SS_medians   <- compute_gene_medians(SS_polyA_log2tpm1,   SS_riboD_log2tpm1,   "SS")
WT_medians   <- compute_gene_medians(WT_polyA_log2tpm1,   WT_riboD_log2tpm1,   "WT")
NB_medians   <- compute_gene_medians(NB_polyA_log2tpm1,   NB_riboD_log2tpm1,   "NB")
ALL_medians   <- compute_gene_medians(ALL_polyA_log2tpm1,   ALL_riboD_log2tpm1,   "ALL")
AML_medians   <- compute_gene_medians(AML_polyA_log2tpm1,   AML_riboD_log2tpm1,   "AML")

combined_medians <- bind_rows(aRMS_medians, SS_medians, WT_medians, NB_medians, ALL_medians, AML_medians)
```

```{r custom plot theme}
plot_theme <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "top",
      legend.title = element_blank(),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.4),
      axis.ticks = element_line(color = "black", linewidth = 0.4),
      strip.text = element_text(face = "bold", size = base_size * 0.9),
      plot.title = element_text(face = "bold", size = base_size * 1, hjust = 0.5),
      plot.margin = margin(10, 10, 10, 10)
    )
}

scale_fill_compendia <- function() {
  scale_fill_manual(values = c(
    "riboD" = "#E69F00",  # yellow
    "polyA" = "#0072B2"   # blue
  ))
}
scale_color_compendia <- function() {
  scale_color_manual(values = c( 
    "riboD" = "#E69F00", 
    "polyA" = "#0072B2" ))
  }
```

convert EnsemblIDs to Hugo IDs for readability

```{r convert to hugo}
convert_to_hugo <- function(df) {
  df %>%
    left_join(gene_names, by = c("Gene" = "EnsGeneID")) %>%
    mutate(Gene = ifelse(!is.na(HugoID), HugoID, Gene)) %>%
    select(-HugoID)
}

combined_medians_hugo <- combined_medians %>%
  convert_to_hugo()
```

Compute residual for each disease

```{r compute residual}
# for each gene in each disease, taking the average riboD expression and subtracting from the average polyA expression
combined_medians_hugo_res <- combined_medians_hugo %>%
  mutate(
    residual = riboD_medians - polyA_medians
  )
```

Define representative polyadenylated and non-polyadenylated genes

```{r representative polyA+ and polyA- genes}
# representative polyadenylated genes were found in Fig1E_F, calculated as the median expression level in PolyA samples divided by the median expression level in RiboD samples


poly_genes <- c(
  SS  = "MTMR2",
  aRMS = "TM2D1",
  WT   = "SYPL1",
  NB   = "CCDC64",
  ALL     = "XPO5",
  AML     = "EML3"
    
)
nonpoly_gene <- "HIST1H1B"
```

Color labeled genes

```{r label genes function}
combined_medians_hugo_res <- combined_medians_hugo_res %>%
  mutate(
    label_color = case_when(
      Gene %in% poly_genes[Disease] ~ "#BB5566",                     # pinkish red
      Gene %in% nonpoly_gene ~ "#BB5566",                         # pinkish red
      TRUE ~ "grey70"
    )
  )
```

Create labeled dateframe with representative genes

```{r representative genes}
combined_medians_hugo_res <- combined_medians_hugo_res %>%
  mutate(
    label = case_when(
      Gene == poly_genes[Disease] ~ Gene,
      Gene == nonpoly_gene ~ Gene,            # HIST1H1B + median ratio gene
      TRUE ~ NA_character_
    )
  )
# this creates a dataframe that shows the polyA median expression and riboD median expression with the calculated residual for each disease. The representative polyadenylated genes are colored for their respective disease. HIST1H1B is colored for all diseases as the representative nonpolyadenylated gene.
write_tsv(combined_medians_hugo_res, "../../output_data/Fig1G/combined_medians_hugo_res.tsv.gz")
```

Fig1G

Plot average expression across library prep methods for each disease

```{r Fig1G}
diseases <- unique(combined_medians_hugo_res$Disease)

Fig1G <- map(diseases, function(d) {

  df <- combined_medians_hugo_res %>% filter(Disease == d)

 ggplot(df, aes(x = polyA_medians, y = riboD_medians)) +
    
   # background points
    geom_point(data = df %>% filter(is.na(label)),
               color = "grey70", alpha = 0.5, size = 1.2) +

    # highlighted labeled points
    geom_point(data = df %>% filter(!is.na(label)),
               aes(color = label_color),
               size = 2.5, alpha = 0.9) +

    
       # Labels for representative genes and HIST1H1B
    geom_text_repel(
      aes(label = label),
      size = 4.5,
      max.overlaps = Inf,
      min.segment.length = 0,
      box.padding = 0.4,
      point.padding = 0.2
    ) +

    scale_color_identity() +   # use colors exactly as defined

    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "gray50") +

    labs(
      title = paste(d),
      x = "Average PolyA Expression (log2(TPM+1))",
      y = "Average RiboD Expression (log2(TPM+1))",
    ) +
    theme_minimal(base_size = 14) +
    plot_theme()
})
names(Fig1G) <- diseases

Fig1G
```

```{r save Fig1G}
ggsave("../../Figures/Fig1G.png", plot = Fig1G$SS)
ggsave("../../Figures/Fig1G.tif", plot = Fig1G$SS)
```

## Fig S6

```{r Fig1G_tocombine}
diseases <- unique(combined_medians_hugo_res$Disease)

Fig1G_all <- map(diseases, function(d) {

  df <- combined_medians_hugo_res %>% filter(Disease == d)

 ggplot(df, aes(x = polyA_medians, y = riboD_medians)) +
    
   # background points
    geom_point(data = df %>% filter(is.na(label)),
               color = "grey70", alpha = 0.5, size = 4) +

    # highlighted labeled points
    geom_point(data = df %>% filter(!is.na(label)),
               aes(color = label_color),
               size = 8, alpha = 0.9) +

    
       # Labels for representative genes and HIST1H1B
    geom_text_repel(
      aes(label = label),
      size = 12,
      max.overlaps = Inf,
      min.segment.length = 0,
      box.padding = 0.4,
      point.padding = 0.2
    ) +

    scale_color_identity() +   # use colors exactly as defined

    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "gray50") +

    labs(
      title = paste(d),
      x = "Average PolyA Expression (log2(TPM+1))",
      y = "Average RiboD Expression (log2(TPM+1))",
    ) +
    theme_minimal(base_size = 14) +
    plot_theme() +
    coord_cartesian(xlim = c(0, 15), ylim = c(0, 20)) +
    theme(
    axis.title.y = element_text(size = 32),
    axis.title.x = element_text(size = 32),
    axis.text.y = element_text(size = 28),
    axis.text.x = element_text(size = 28),
    strip.text = element_blank(),
    plot.title = element_text(vjust = -2),
    legend.text = element_text(size = 26),       # <-- label text size
    legend.title = element_text(size = 26)      # <-- title text size
    )
})
names(Fig1G_all) <- diseases

Fig1G_all
```

```{r FigS6}
FigS6 <- wrap_plots(Fig1G_all, ncol = 1) +
  plot_layout(guides = "collect", axis_titles = "collect", axes = "collect") +
  plot_annotation(
    theme = theme(legend.position = "bottom")
  ) &
  theme(
    plot.margin = margin(t = 2, b = 2, l = 4, r = 4),
    plot.title = element_text(face = "bold", size = 36)
  )
FigS6
```

```{r save FigS6}
ggsave("../../Figures/FigS6.png", FigS6, width = 20, height = 36, dpi = 500)
ggsave("../../Figures/FigS6.tif", FigS6, width = 20, height = 36, dpi = 500)
```

Session Info

```{r session info}
sessioninfo::session_info()
```
