---
title: "Fig_1C"
Date: "`r Sys.Date()`"
format: 
  gfm:
    df-print: kable
editor: visual
---

## Fig 1C

Stacked bar chart showing the proportion of unique genes expressed at different log2(TPM+1) levels.

```{r install package}
library(tidyverse)
```

```{r load in files}
# expression files
SS_aRMS_log2tpm1 <- read_tsv("../input_data/SS_aRMS_log2TPM1_ensembl_TEBA.tsv.gz")
WT_NB_log2tpm1 <- read_tsv("../input_data/WT_NB_log2TPM1_ensembl_TEBA.tsv.gz")
ALL_AML_log2tpm1 <- read_tsv("../input_data/ALL_AML_log2TPM1_ensembl_TEBA.tsv.gz")

# sample ID files, from correlation_analysis
filtered_SS_aRMS_list <- read_tsv("../input_data/filtered_SS_aRMS_list.tsv")
filtered_WT_NB_list <- read_tsv("../input_data/filtered_WT_NB_list.tsv")
filtered_ALL_AML_list <- read_tsv("../input_data/filtered_ALL_AML_list.tsv")
```

```{r reformat files}
# Since "Gene" is a column name, it won't match the row names
# use values from the first column ("Gene") as row names
fix_SS_aRMS_log2tpm1 <- SS_aRMS_log2tpm1 %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

fix_WT_NB_log2tpm1 <- WT_NB_log2tpm1 %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

fix_ALL_AML_log2tpm1 <- ALL_AML_log2tpm1 %>%
  remove_rownames %>%
  column_to_rownames(var = "Gene")

# use values from first column ("th_dataset_id") as row names
fix_filtered_SS_aRMS_list <- filtered_SS_aRMS_list %>%
  remove_rownames %>%
  column_to_rownames(var = "term")

fix_filtered_WT_NB_list <- filtered_WT_NB_list %>%
  remove_rownames %>%
  column_to_rownames(var = "term")

fix_filtered_ALL_AML_list <- filtered_ALL_AML_list %>%
  remove_rownames %>%
  column_to_rownames(var = "term")
```

Separating THIDs into disease and lib prep

```{r separating THIDs into disease + prep}
# SS polyA
SS_polyA_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("SS_polyA")) %>% 
  select(term, disease_and_prep)

SS_polyA_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(SS_polyA_list$term))

# SS riboD
SS_riboD_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("SS_riboD")) %>% 
  select(term, disease_and_prep)

SS_riboD_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(SS_riboD_list$term))

# aRMS polyA
aRMS_polyA_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("aRMS_polyA")) %>% 
  select(term, disease_and_prep)

aRMS_polyA_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(aRMS_polyA_list$term))

# aRMS riboD
aRMS_riboD_list <- filtered_SS_aRMS_list %>%
  filter(disease_and_prep %in% c("aRMS_riboD")) %>% 
  select(term, disease_and_prep)

aRMS_riboD_expr <- SS_aRMS_log2tpm1 %>%
  select(Gene, all_of(aRMS_riboD_list$term))

# WT polyA
WT_polyA_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("WT_polyA")) %>% 
  select(term, disease_and_prep)

WT_polyA_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(WT_polyA_list$term))

# WT riboD 
WT_riboD_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("WT_riboD")) %>% 
  select(term, disease_and_prep)

WT_riboD_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(WT_riboD_list$term))

# NB polyA
NB_polyA_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("NB_polyA")) %>% 
  select(term, disease_and_prep)

NB_polyA_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(NB_polyA_list$term))

# NB riboD
NB_riboD_list <- filtered_WT_NB_list %>%
  filter(disease_and_prep %in% c("NB_riboD")) %>% 
  select(term, disease_and_prep)

NB_riboD_expr <- WT_NB_log2tpm1 %>%
  select(Gene, all_of(NB_riboD_list$term))

# ALL polyA
ALL_polyA_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("ALL_polyA")) %>% 
  select(term, disease_and_prep)

ALL_polyA_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(ALL_polyA_list$term))

# ALL riboD
ALL_riboD_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("ALL_riboD")) %>% 
  select(term, disease_and_prep)

ALL_riboD_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(ALL_riboD_list$term))


# AML polyA
AML_polyA_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("AML_polyA")) %>% 
  select(term, disease_and_prep)

AML_polyA_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(AML_polyA_list$term))

# AML riboD
AML_riboD_list <- filtered_ALL_AML_list %>%
  filter(disease_and_prep %in% c("AML_riboD")) %>% 
  select(term, disease_and_prep)

AML_riboD_expr <- ALL_AML_log2tpm1 %>%
  select(Gene, all_of(AML_riboD_list$term))
```

Defining log2(TPM+1) bins

```{r define expr bins}
expr_bins <- c(0, 1e-10, 0.09, 0.99, 2.99, 4.99, Inf)

expr_labels <- c("0","0-0.09", "0.1-0.99", "1-2.99", "3-4.99", ">5")
```

define gene medians function

```{r gene medians function}
compute_gene_medians <- function(df, disease, compendia) {
  df %>% 
    rowwise() %>% 
    mutate(median_expr = median(c_across(starts_with("T")), na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(
      Disease = disease,
      Compendia = compendia
    ) %>%
    select(Gene, median_expr, Disease, Compendia)
}

```

Compute gene medians

```{r compute gene medians}
SS_polyA_med   <- compute_gene_medians(SS_polyA_expr,   "SS",   "polyA")
SS_riboD_med   <- compute_gene_medians(SS_riboD_expr,   "SS",   "riboD")
aRMS_polyA_med <- compute_gene_medians(aRMS_polyA_expr, "aRMS", "polyA")
aRMS_riboD_med <- compute_gene_medians(aRMS_riboD_expr, "aRMS", "riboD")
WT_polyA_med   <- compute_gene_medians(WT_polyA_expr,   "WT",   "polyA")
WT_riboD_med   <- compute_gene_medians(WT_riboD_expr,   "WT",   "riboD")
NB_polyA_med   <- compute_gene_medians(NB_polyA_expr,   "NB",   "polyA")
NB_riboD_med   <- compute_gene_medians(NB_riboD_expr,   "NB",   "riboD")
AML_polyA_med   <- compute_gene_medians(AML_polyA_expr,   "AML",   "polyA")
AML_riboD_med   <- compute_gene_medians(AML_riboD_expr,   "AML",   "riboD")
ALL_polyA_med   <- compute_gene_medians(ALL_polyA_expr,   "ALL",   "polyA")
ALL_riboD_med  <- compute_gene_medians(ALL_riboD_expr,   "ALL",   "riboD")

all_medians <- bind_rows(
  SS_polyA_med, SS_riboD_med,
  aRMS_polyA_med, aRMS_riboD_med,
  WT_polyA_med, WT_riboD_med,
  NB_polyA_med, NB_riboD_med,  
  ALL_polyA_med, ALL_riboD_med,
  AML_riboD_med, AML_polyA_med
  )
```

Bin genes into expression thresholds

```{r bin gene medians}
all_medians_bins <- all_medians %>%
  mutate(
    ExprBin = cut(
      median_expr,
      breaks = expr_bins,
      labels = expr_labels,
      include.lowest = TRUE
    )
  )
```

Compute proportion of genes in each bin per lib prep

```{r compute proportions}
bin_prop <- all_medians_bins %>% 
  group_by(Compendia, ExprBin, Disease) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Disease, Compendia) %>%
  mutate(Proportion = n / sum(n))

```

Custom color theme

```{r custom theme}
theme_2D <- function(base_size = 14) {
  theme_minimal(base_size = base_size) +
    theme(
      legend.position = "right",
      legend.title = element_blank(),
      panel.grid.major = element_line(color = "grey85", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      axis.line = element_line(color = "black", linewidth = 0.4),
      axis.ticks = element_line(color = "black", linewidth = 0.4),
      strip.text = element_text(face = "bold", size = base_size * 0.9),
      plot.title = element_text(face = "bold", size = base_size * 1.1, hjust = 0.5),
      axis.title.y = element_text(angle = 90, hjust = 0.5, size = 16), 
      axis.title.x = element_text(angle = 0, hjust = 0.5, size = 17), 
      axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
      plot.margin = margin(10, 10, 10, 10)
    )
}

scale_fill_compendia <- function() {
  scale_fill_manual(values = c(
    "polyA" = "#E69F00",  # yellow
    "riboD" = "#0072B2"   # blue
  ))
}
scale_color_compendia <- function() {
  scale_color_manual(values = c( 
    "polyA" = "#E69F00", 
    "riboD" = "#0072B2" ))
  }

```

Generate stacked bar plot

```{r}
# Get unique diseases
diseasesC <- unique(bin_prop$Disease)

# Generate one plot per disease
plots_by_diseaseC <- map(diseasesC, function(d) {
  
  dfC <- bin_prop %>% filter(Disease == d)
  
  ggplot(dfC, aes(x = Compendia, y = Proportion, fill = ExprBin)) +
    geom_bar(stat = "identity", position = "stack") +
    scale_fill_viridis_d(option = "D") +
    labs(
      title = paste("Proportion of Genes in log2(TPM+1) Thresholds -", d),
      x = "Library Prep",
      y = "Proportion of Genes",
      fill = "Expression Bin"
    ) +
    theme_2D()
})

# Name the plots by disease
names(plots_by_diseaseC) <- diseasesC
plots_by_diseaseC
```

Save plot

```{r}
# ggsave("/figures/fig1C.png", plot = plots_by_diseaseC, width = 6, height = 4, units = "in", dpi = 300)
```

Session Info

```{r}
sessioninfo::session_info()
```
